---
title: "HAMMS - network & clustering analysis"
author: "William Louie"
date: "`r Sys.Date()`"
output: html_document
---

### Load Libraries

```{r load libraries, warning=FALSE, message=FALSE}

library(moire)
library(dcifer)
library(dplyr)
library(magrittr)
library(ggplot2)
library(igraph) #v0.5.1 supports "add_shape" function
library(ggnewscale)
library(descriptr)
library(readr)
library(plotly)
library(stringr)
library(ggpubr)
library(tidyr)

```

Import dcifer results: "dcifer_results.rds"

```{r import, exercise=TRUE, echo=F}

# setwd("/Users/williamlouie/Dropbox/My Mac (Williams-MacBook-Pro.local)/Downloads/Ethiopia/HAMMS/hamms_dcifer_NEW")

long_ibd <- read_csv("C:/Users/Yisihak/Desktop/tes_multiple_run_result/dcifer/long_ibd_updated.csv")
long_ibd_rp5 <- read_csv("C:/Users/Yisihak/Desktop/tes_multiple_run_result/dcifer/long_ibd_125_updated.csv") %>% 
  filter(grepl("D0", sample.x)) %>% 
  filter(grepl("D0", sample.y)) %>% 
  mutate(site.x = case_when(
    grepl("AA", sample.x) ~ "Asayita",
    grepl("AC", sample.x) ~ "Abobo",
    grepl("BA", sample.x) ~ "Asosa",
    grepl("GM", sample.x) ~ "Maksegnit",
    grepl("MM", sample.x) ~ "Mizan",
  )) %>% 
  mutate(site.y = case_when(
    grepl("AA", sample.y) ~ "Asayita",
    grepl("AC", sample.y) ~ "Abobo",
    grepl("BA", sample.y) ~ "Asosa",
    grepl("GM", sample.y) ~ "Maksegnit",
    grepl("MM", sample.y) ~ "Mizan",
  ))


check_BH <- long_ibd %>% filter(p_value <= 0.05) %>% 
  select(p_value, p_BH) %>% 
  filter(p_BH <= 0.05) %>% 
  summarize(p_value = max(p_value))
# threshold
p_val_BH <- check_BH$p_value


check_BH <- long_ibd_rp5 %>% filter(p_value <= 0.05) %>%
  select(p_value, p_BH) %>%
  filter(p_BH <= 0.05) %>%
  summarize(p_value = max(p_value))
# threshold
p_val_BH_0.125 <- check_BH$p_value






threshold <- 0.5

# List of all samples
all_samples <- unique(c(long_ibd_rp5$sample.x, long_ibd_rp5$sample.y))

# Samples that are related to at least one sample from another site
cross_site_samples <- long_ibd_rp5 %>%
  filter(same_site == "Different site", ibd >= threshold) %>%
  select(sample.x, sample.y) %>%
  pivot_longer(cols = everything(), values_to = "sample") %>%
  distinct(sample) %>%
  pull(sample)

# Proportion
proportion_cross_site <- length(cross_site_samples) / length(all_samples)


length(cross_site_samples)

length(all_samples)

proportion_cross_site





threshold <- 0.5

# All samples and their sites
sample_sites <- long_ibd_rp5 %>%
  select(sample = sample.x, site = site.x) %>%
  bind_rows(long_ibd_rp5 %>% select(sample = sample.y, site = site.y)) %>%
  distinct()

# Samples that have cross-site IBD >= threshold
cross_site_samples <- long_ibd_rp5 %>%
  filter(ibd >= threshold, site.x != site.y) %>%
  pivot_longer(cols = c(sample.x, sample.y), values_to = "sample", names_to = NULL) %>%
  distinct(sample)

# Summarize per site
result <- sample_sites %>%
  mutate(cross = sample %in% cross_site_samples$sample) %>%
  group_by(site) %>%
  summarise(
    total_samples = n(),
    cross_site_related = sum(cross),
    proportion_cross_site_related = cross_site_related / total_samples
  )

result





threshold <- 0.5

site_pair_counts <- long_ibd_rp5 %>%
  filter(ibd >= threshold) %>%
  group_by(site.x, site.y) %>%
  summarise(n_pairs = n(), .groups = "drop") %>%
  # ensure symmetric matrix
  mutate(row_site = pmin(site.x, site.y),
         col_site = pmax(site.x, site.y)) %>%
  group_by(row_site, col_site) %>%
  summarise(n_pairs = sum(n_pairs), .groups = "drop") %>%
  pivot_wider(names_from = col_site, values_from = n_pairs, values_fill = 0)

site_pair_counts



site_pair_prop <- long_ibd_rp5 %>%
  group_by(site.x, site.y) %>%
  summarise(
    total_pairs = n(),
    high_ibd_pairs = sum(ibd >= threshold),
    prop_high_ibd = high_ibd_pairs / total_pairs,
    .groups = "drop"
  ) %>%
  mutate(row_site = pmin(site.x, site.y),
         col_site = pmax(site.x, site.y)) %>%
  group_by(row_site, col_site) %>%
  summarise(
    prop_high_ibd = sum(high_ibd_pairs) / sum(total_pairs),
    .groups = "drop"
  ) %>%
  pivot_wider(names_from = col_site, values_from = prop_high_ibd, values_fill = 0)

site_pair_prop








threshold <- 0.5

# All samples with their sites
sample_sites <- long_ibd_rp5 %>%
  select(sample = sample.x, site = site.x) %>%
  bind_rows(long_ibd_rp5 %>% select(sample = sample.y, site = site.y)) %>%
  distinct()

# For each sample, identify all sites it is related to (via IBD≥threshold)
cross_links <- long_ibd_rp5 %>%
  filter(ibd >= threshold, site.x != site.y) %>%
  select(sample = sample.x, site = site.x, related_site = site.y) %>%
  bind_rows(
    long_ibd_rp5 %>%
      filter(ibd >= threshold, site.x != site.y) %>%
      select(sample = sample.y, site = site.y, related_site = site.x)
  ) %>%
  distinct(sample, site, related_site)

# Count for each site how many samples link to each other site
site_to_site_counts <- cross_links %>%
  group_by(site, related_site) %>%
  summarise(count = n_distinct(sample), .groups = "drop")

# Total samples per site
site_totals <- sample_sites %>%
  count(site, name = "total_samples")

# Merge & format
formatted <- site_to_site_counts %>%
  left_join(site_totals, by = "site") %>%
  mutate(
    formatted = sprintf("%d/%d (%.1f%%)", count, total_samples, 100 * count / total_samples)
  ) %>%
  select(site, related_site, formatted) %>%
  pivot_wider(names_from = related_site, values_from = formatted, values_fill = "0/total (0%)")

formatted



```

### COI paired samples

```{r coi_paired, exercise=TRUE, echo=F}

tes_pairs <- read_csv("C:/Users/Yisihak/Desktop/tes_multiple_run_result/TES_pairs.csv")

coi_summary <- read_csv("C:/Users/Yisihak/Desktop/tes_multiple_run_result/moire/coi_summary.csv") %>%
  select(-site)

aster_results <- read_csv("C:/Users/Yisihak/Desktop/tes_multiple_run_result/aster_result.csv") 

coi_tes <- coi_summary %>%
            inner_join(tes_pairs, by = "sample_id") %>%
            mutate(day = case_when(grepl("D0", sample_id) ~ "D0",
                                   !grepl("D0", sample_id) ~ "Df")) %>%
          group_by(person_id) %>%
          filter(n() >= 2) %>% 
          ungroup() %>%
          inner_join(aster_results, by = "person_id") %>%
          select(sample_id, person_id, day, day_sp, site,
                 post_coi_mean, post_effective_coi_mean, Aster) 


# filter long_ibd for sample pairs

d0 <- coi_tes %>%
  filter(day == "D0")
d0_samples <- d0$sample_id

df <- coi_tes %>%
  filter(day == "Df")
df_samples <- df$sample_id

tes_samples <- c(d0_samples, df_samples)

# Create lines connecting each pair, COI
ibd_tes <- long_ibd_rp5 %>%
  filter(
    (sample.x %in% d0_samples & sample.y %in% df_samples) | 
    (sample.x %in% df_samples & sample.y %in% d0_samples)) %>%
  left_join(coi_tes, by = c("sample.x" = "sample_id")) %>%
  rename(x1 = day_sp, y1 = post_coi_mean, site1 = site) %>%
  left_join(coi_tes, by = c("sample.y" = "sample_id")) %>%
  rename(x2 = day_sp, y2 = post_coi_mean, site2 = site) %>%
  mutate(aster = case_when(Aster.x == "1" | Aster.x == "1" ~ "yes",
                           !(Aster.x == "1" | Aster.x == "1") ~ "no")) %>%
  filter(person_id.x == person_id.y) %>%
  select(-c(site1, person_id.x, Aster.x, Aster.y)) %>%
  rename(site = site2, person_id = person_id.y) 
# Create lines connecting each pair, eCOI
ibd_tes2 <- long_ibd_rp5 %>%
  filter(
    (sample.x %in% d0_samples & sample.y %in% df_samples) | 
    (sample.x %in% df_samples & sample.y %in% d0_samples)) %>%
  left_join(coi_tes, by = c("sample.x" = "sample_id")) %>%
  rename(x1 = day_sp, y1 = post_effective_coi_mean, site1 = site) %>%
  select(-post_coi_mean) %>%
  left_join(coi_tes, by = c("sample.y" = "sample_id")) %>%
  rename(x2 = day_sp, y2 = post_effective_coi_mean, site2 = site) %>%
  mutate(aster = case_when(Aster.x == "1" | Aster.x == "1" ~ "yes",
                           !(Aster.x == "1" | Aster.x == "1") ~ "no")) %>%
  filter(person_id.x == person_id.y) %>%
  select(-c(site1, person_id.x, Aster.x, Aster.y)) %>%
  rename(site = site2, person_id = person_id.y) 


# Initial COI plot for each D0 Df comparison
# added line segments connecting the paired person_id
g_tes_coi <- ggplot(data = coi_tes, aes(x = day_sp, y = post_coi_mean)) +
  geom_point(aes(size = factor(Aster),
                 shape = factor(Aster))) +
  scale_shape_manual(values = c(
          "0" = "circle", "1" = "triangle"), guide = "none") +
  scale_size_manual(
    values = c("0" = 2, "1" = 5), guide = "none") +
  geom_segment(
    data = ibd_tes,
    aes(x = x1, y = y1, xend = x2, yend = y2, color = ibd),
    inherit.aes = FALSE, linewidth = 0.8,
    linetype = ifelse(ibd_tes$aster == "yes", "solid", "dotted")) +
  theme_bw() +
  theme(panel.grid.major = element_blank()) +
  scale_color_gradient(
    low = "grey", high = "blue", name = "IBD") +
  facet_wrap(~site) +
  ylab("Posterior mean COI") +
  xlab("Day")
g_tes_coi
g_tes_ecoi <- ggplot(data = coi_tes, aes(x = day_sp, y = post_effective_coi_mean)) +
  geom_point(aes(size = factor(Aster),
                 shape = factor(Aster))) +
  scale_shape_manual(values = c(
          "0" = "circle", "1" = "triangle"), guide = "none") +
  scale_size_manual(
    values = c("0" = 2, "1" = 5), guide = "none") +
  geom_segment(
    data = ibd_tes2,
    aes(x = x1, y = y1, xend = x2, yend = y2, color = ibd),
    inherit.aes = FALSE, linewidth = 0.8,
    linetype = ifelse(ibd_tes$aster == "yes", "solid", "dotted")) +
  theme_bw() +
  theme(panel.grid.major = element_blank()) +
  scale_color_gradient(
    low = "grey", high = "blue", name = "IBD") +
  facet_wrap(~site) +
  ylab("Posterior mean eCOI") +
  xlab("Day")
g_tes_ecoi

g_coi_pair <- ggarrange(g_tes_coi, g_tes_ecoi, nrow = 2, common.legend = TRUE)
ggsave("C:/Users/Yisihak/Desktop/tes_multiple_run_result/moire/Figure_coi_pairs.jpeg", plot = g_coi_pair, height = 10, width = 8)


```

### Global clustering
Dcifer cluster plot, using BH correction

```{r network, exercise=TRUE, echo=F}

# Should only be looking at D0s
long_ibd_d0 <- long_ibd %>%
  filter(grepl("D0", sample.x) & grepl("D0", sample.y))
long_ibd_rp5_d0 <- long_ibd_rp5 %>%
  filter(grepl("D0", sample.x) & grepl("D0", sample.y))

# Define the function for plotting the network
plot_network <- function(long_ibd_rp5, relatedness_threshold = 0.5, pvalue_threshold = 0.05, 
                         scaling_factor = 1.0, jitter_amount = 0.1, 
                         show_isolates = TRUE, edge_color_by = same_site, vertex_size = 5, seed = 100) { 
  
  # Set seed for reproducibility
  set.seed(seed)
  
  # Filter edges based on relatedness (ibd) and p-value thresholds
  filtered_edges <- long_ibd %>%
    filter(ibd >= relatedness_threshold, p_BH < pvalue_threshold)

  # Create a data frame of unique sample names for vertices
  unique_samples <- unique(c(filtered_edges$sample.x, filtered_edges$sample.y))
  
  # Create an igraph object from the filtered edges
  g <- graph_from_data_frame(filtered_edges, directed = FALSE, vertices = data.frame(name = unique_samples))

  # Identify isolates (samples that are not in the edges)
  samples_in_edges <- unique(c(filtered_edges$sample.x, filtered_edges$sample.y))
  all_samples <- unique(c(long_ibd$sample.x, long_ibd$sample.y))
  isolates <- setdiff(all_samples, samples_in_edges)

  # Add isolates as vertices in the graph if show_isolates is TRUE
  if (show_isolates) {
    g_full <- g + vertices(isolates)
  } else {
    g_full <- g  # Only use the main graph without isolates
  }

  # # Merge metadata to get site information for coloring vertices
  vertex_data <- data.frame(name = V(g_full)$name) %>%
    mutate(site = case_when(
      grepl(pattern = "AA", name) ~ "Asayita",
      grepl(pattern = "AC", name) ~ "Gambella",
      grepl(pattern = "GM", name) ~ "Gondar",
      grepl(pattern = "BA", name) ~ "Assossa",
      grepl(pattern = "ARC", name) ~ "Djibouti",
      grepl(pattern = "MM", name) ~ "Mizan"))
    
  #   left_join(metadata_cluster, by = "name")  # Updated to use metadata_cluster

  # Check for NA in site and set default color if necessary
  if ("site" %in% colnames(vertex_data)) {
    # Set vertex colors based on site
    unique_populations <- unique(vertex_data$site)
    colors <- c("Asayita" = "darkred", "Gambella" = "yellow3",
                "Assossa" = "darkorange", "Gondar" = "blue", "Mizan" = "darkviolet")

    # Map colors to sites
    color_map <- setNames(colors, unique_populations)

    # Assign colors to vertices based on site
    V(g_full)$color <- color_map[as.character(vertex_data$site)]
    
    # Assign default color for any samples without a site
    V(g_full)$color[is.na(V(g_full)$color)] <- "lightgray"  # Default color for unmatched samples
  } else {
    V(g_full)$color <- "gray"  # Default color if site is not found
  }

  # Set vertex size
  V(g_full)$size <- vertex_size

  # Color edges based on the specified category (now "cat")
  if (edge_color_by %in% colnames(filtered_edges)) {
    edge_categories <- filtered_edges %>%
      select(sample.x, sample.y, !!sym(edge_color_by)) %>%
      distinct()

    # Ensure there are no NA values in the edge categories
    edge_categories <- edge_categories[!is.na(edge_categories[[edge_color_by]]), ]

    if (nrow(edge_categories) > 0) {
      # Manually set edge colors based on "same_site"
      E(g_full)$color <- ifelse(edge_categories[[edge_color_by]] == "Same site", "darkgoldenrod", 
                                ifelse(edge_categories[[edge_color_by]] == "Different site", "grey", NA))
      
      # Set a default color for any unmatched edges
      E(g_full)$color[is.na(E(g_full)$color)] <- "black"  # Use black for NA values
    } else {
      E(g_full)$color <- "black"  # If no edge categories, default to black
    }
  } else {
    E(g_full)$color <- "black"  # Default edge color if category not found
  }

  # Set edge widths to ensure visibility
  edge_widths <- pmax(E(g_full)$ibd * 2, 0.2)

  # Use layout_nicely() to automatically select an appropriate layout
  layout <- layout_nicely(g_full)

  # Jitter the layout coordinates to avoid overlap
  jitter_matrix <- matrix(rnorm(nrow(layout) * 2, mean = 0, sd = jitter_amount), ncol = 2)
  layout <- layout + jitter_matrix

  # Plot the graph
  plot(g_full, 
       layout = layout, 
       vertex.label = NA,  # Remove labels
       vertex.size = V(g_full)$size,  # Use modifiable vertex sizes
       edge.width = edge_widths,  # Use adjusted edge widths
       edge.color = E(g_full)$color,  # Use colors for edges
       main = paste("Relatedness Threshold:", 
                    relatedness_threshold))

  # Add a legend for the vertex colors (site)
  legend("bottomright", legend = names(color_map), col = color_map, pch = 19, pt.cex = 1.5, title = "site")
  # Add a legend for the edge colors (Category)
  legend("topright", legend = c("Same site", "Different site"), col = c("darkgoldenrod", "grey"), lty = 1, title = edge_color_by)
  
}

# pdf("C:/Users/Yisihak/Desktop/tes_multiple_run_result/dcifer/global_relatedness_1.pdf", width = 12, height = 10)

# Plot clusters
plot_network(long_ibd_rp5_d0, relatedness_threshold = 1, pvalue_threshold = p_val_BH, scaling_factor = 1, jitter_amount = 0.7, vertex_size = 5, show_isolates = FALSE, edge_color_by = "same_site", seed = 300)

# dev.off()

pdf("C:/Users/Yisihak/Desktop/tes_multiple_run_result/dcifer/global_relatedness_0.5_jitter.pdf", width = 12, height = 10)
# Plot clusters
plot_network(long_ibd_rp5_d0, relatedness_threshold = 0.5, pvalue_threshold = p_val_BH, scaling_factor = 1, jitter_amount = 0, vertex_size = 5, show_isolates = FALSE, edge_color_by = "same_site", seed = 300)

dev.off()

# pdf("C:/Users/Yisihak/Desktop/tes_multiple_run_result/dcifer/global_relatedness_1_0.5_updated.pdf", width = 12, height = 10)
# Save
# jpeg("hamms_dcifer_NEW/Figure_network_plots_1and05.jpg", width = 1500, height = 1200, res = 150)
par(mfrow = c(1, 2)) 
plot_network(long_ibd_rp5_d0, relatedness_threshold = 1, pvalue_threshold = p_val_BH, scaling_factor = 1, jitter_amount = 0.7, vertex_size = 4, show_isolates = FALSE, edge_color_by = "same_site", seed = 300)
plot_network(long_ibd_rp5_d0, relatedness_threshold = 0.5, pvalue_threshold = p_val_BH, scaling_factor = 1, jitter_amount = 0.7, vertex_size = 4, show_isolates = FALSE, edge_color_by = "same_site", seed = 300)
# dev.off()
# Reset the plotting area to default
# par(mfrow = c(1, 1))
# Close the JPEG device
# dev.off()


```

### Clustering by HRP deletions
Determine whether HRP-deleted samples are also connected by IBD, for each site

```{r network_hrp, exercise=TRUE, echo=F}

# Import CNV report
cnv_metadata <- read_csv("C:/Users/Yisihak/Desktop/tes_multiple_run_result/dcifer/k13_for_relatedness.csv")

# filter samples
cnv_samples <- cnv_metadata$SampleID
long_ibd_cnv <- long_ibd_rp5  %>%
  filter(sample.x %in% cnv_samples & sample.y %in% cnv_samples) %>% 
  filter(grepl("D0", sample.x) & grepl("D0", sample.y))

plot_network_cnv_site <- function(long_ibd_cnv, cnv_metadata, relatedness_threshold = 0.5, pvalue_threshold = 0.05, 
                                  show_isolates = TRUE, seed = 100, site_filter) {  

  # Set seed for reproducibility
  set.seed(seed)

  # Mapping of site abbreviations to full names
  site_name_mapping <- c("AA" = "Asayita", 
                         "AC" = "Abobo", 
                         "GM" = "Gondar", 
                         "BA" = "Assossa",
                         "MM" = "Mizan")

  # Full site name based on site_filter (abbreviation)
  full_site_name <- site_name_mapping[site_filter]
  
  # Filter the long_ibd_cnv data frame for relevant samples starting with the site filter
  filtered_samples <- long_ibd_cnv %>%
    filter(grepl(paste("^", site_filter, sep = ""), sample.x) & grepl(paste("^", site_filter, sep = ""), sample.y))

  # Filter edges based on relatedness (ibd) and p-value thresholds
  filtered_edges <- filtered_samples %>%
    filter(ibd >= relatedness_threshold, p_BH < pvalue_threshold)

  # Create a data frame of unique sample names for vertices
  unique_samples <- unique(c(filtered_edges$sample.x, filtered_edges$sample.y))

  # Create an igraph object from the filtered edges
  g <- graph_from_data_frame(filtered_edges, directed = FALSE, vertices = data.frame(name = unique_samples))

  # Identify isolates (samples that are not in the edges)
  samples_in_edges <- unique(c(filtered_edges$sample.x, filtered_edges$sample.y))
  all_samples <- unique(c(long_ibd_cnv$sample.x, long_ibd_cnv$sample.y))
  isolates <- setdiff(all_samples, samples_in_edges)

  # Add isolates as vertices in the graph if show_isolates is TRUE
  if (show_isolates) {
    g_full <- g + vertices(isolates)
  } else {
    g_full <- g  # Only use the main graph without isolates
  }

  # Merge site information and HRP deletion status into vertex_data
  vertex_data <- data.frame(name = V(g_full)$name) %>%
    mutate(site = case_when(
      grepl(pattern = "^AA", name) ~ "Asayita",
      grepl(pattern = "^AC", name) ~ "Gambella",
      grepl(pattern = "^GM", name) ~ "Gondar",
      grepl(pattern = "^BA", name) ~ "Assossa",
      grepl(pattern = "^MM", name) ~ "Mizan")) %>%
    left_join(cnv_metadata, by = c("name" = "SampleID"))  # Join cnv_metadata for HRP deletion status

  # # Define site colors for only relevant sites
  # site_colors <- c("Asayita" = "#8B0000", "Gambella" = "#CCCC00", "Gondar" = "#006400",
  #                  "Assossa" = "#FF8C00", "Mizan" = "#0000FF", "Djibouti" = "#9400D3")

  # Apply site color
  # V(g_full)$color <- ifelse(vertex_data$HRP_stat == "HRP2_only_deleted",
  #                          "red", ifelse(vertex_data$HRP_stat == "HRP3_only_deleted", "blue", ifelse(vertex_data$HRP_stat == "HRP2/3_deleted", "purple", ifelse(vertex_data$HRP_stat == "no_deletion", "grey", "black"))))
# 
  V(g_full)$color <- ifelse(vertex_data$R622_k13 == "I",
                           "#0077b6", ifelse(vertex_data$R622_k13 == "R", "#03045e", ifelse(vertex_data$R622_k13 == "R/I", "#90e0ef", ifelse(vertex_data$R622_k13 == "-", "grey", "black"))))

  # V(g_full)$color <- ifelse(vertex_data$Pfcrt == "CVIET",
  #                          "#dabfff", ifelse(vertex_data$Pfcrt == "CVMNK", "#2c2a4a", ifelse(vertex_data$Pfcrt == "CVMNK|CVIET", "#d67ad1", ifelse(vertex_data$Pfcrt == "-", "grey", "black"))))
#   
#   # Set vertex size: 7 if any deletion is present, 4 if neither
#   V(g_full)$size <- ifelse(vertex_data$Hrp2_del == "Deleted" | vertex_data$Hrp3_del == "Deleted", 
#                            7, 4)
# 
# V(g_full)$shape <- ifelse(vertex_data$Hrp2_del == "Deleted" & vertex_data$Hrp3_del == "Deleted", 
#                             "square", "circle")

  # # Apply transparency adjustments based on deletions:
  # V(g_full)$color <- ifelse(vertex_data$Hrp2_del == "Deleted" & vertex_data$Hrp3_del == "Deleted", 
  #                           adjustcolor(V(g_full)$color, alpha.f = 1),   # Both deletions = no transparency
  #                           ifelse(vertex_data$Hrp2_del == "Deleted" | vertex_data$Hrp3_del == "Deleted", 
  #                                  adjustcolor(V(g_full)$color, alpha.f = 0.9),   # One deletion = slight transparency
  #                                  adjustcolor(V(g_full)$color, alpha.f = 0.2)))  # No deletions = high transparency

  # Set all edge colors to black
  E(g_full)$color <- "black"

  # Set edge widths to ensure visibility
  edge_widths <- pmax(E(g_full)$ibd * 2, 0.2)

  # Use layout_nicely() to automatically select an appropriate layout without jitter
  layout <- layout_nicely(g_full)

  # Create the plot
  plot(g_full, 
       layout = layout, 
       vertex.label = NA,  # Remove labels
       vertex.size = V(g_full)$size,  # Use conditional vertex sizes
       edge.width = edge_widths,  # Use adjusted edge widths
       edge.color = E(g_full)$color,  # All edges set to black
       vertex.shape = V(g_full)$shape,  # Use conditional vertex shapes
       main = full_site_name,
       cex.main = 8) 

  # Add legends (No opacity legend)
  # legend("bottomright",
  #        col = c("#03045e","#0077b6","#90e0ef","grey"),
  #       legend = c( "wildtype", "mutant", "mixed", "NA"),
  #        pch = c(16, 16, 16),  # 16 is circle, 15 is square
  #        pt.cex = c(1, 1, 1),
  #        title = "HRP2/3 deletion")
}


# Define the sites
site_list <- c("AA", "AC", "GM", "BA", "MM")  # Abbreviations for the sites

# Open JPEG device for saving the plots
# jpeg("hamms_dcifer_NEW/Figure_network_HRP_deletions.jpeg", width = 1800, height = 1200, res = 150)
pdf("C:/Users/Yisihak/Desktop/tes_multiple_run_result/dcifer/k13_relatedness_0.9D0.pdf", width = 10, height = 10)

# Set up the plotting grid
par(mfrow = c(2, 3))  # Set 2 rows and 3 columns for grid layout

# Loop over each site and call the function
for (site in site_list) { 
  # Print current site for debugging
  print(paste("Plotting for site:", site))
  
  # Run the function for each site, using the site abbreviation as the filter
  plot_network_cnv_site(long_ibd_cnv = long_ibd_cnv,  # Ensure you're using the correct dataset
                        cnv_metadata = cnv_metadata, 
                        relatedness_threshold = 0.9, 
                        pvalue_threshold = p_val_BH_0.125,
                        show_isolates = FALSE, 
                        seed = 200, 
                        site_filter = site)  # Pass the abbreviation to filter
}

# Close the JPEG device after plotting
dev.off()


```

### (old) Cluster plot: Asayita

```{r cluster_aa, exercise=TRUE, echo=F}

# Make dcifer_list for cluster plot
dres0 <- dres_aa[, , "p_value"]

# Identify rows and columns with "D0"
# Filter dres0
row_filter <- grepl("D0", rownames(dres0))
col_filter <- grepl("D0", colnames(dres0))
dres0 <- dres0[row_filter, col_filter, drop = FALSE]
dres0[upper.tri(dres0, diag = F)] <- 0
dres1 <- dres0 + t(dres0) 

# Filter dmat
dmat <- dres_aa[, , "estimate"]
dmat <- dmat[row_filter, col_filter, drop = FALSE]
dmat[upper.tri(dmat)] <- t(dmat)[upper.tri(t(dmat))]

isSymmetric.matrix(dres1) #TRUE
isSymmetric.matrix(dmat) #TRUE

dcifer_list = list(id=colnames(dmat), relatedness=dmat, pvalue=dres1)

# Relatedness distribution
hist_relatedness <- hist(dmat[upper.tri(dmat, diag = F)], xlab = "Estimated pairwise relatedness", 
     main = "Histogram of Dcifer estimated pairwise relatedness", breaks = 50) 
r_value_df <- data.frame(r_value = dmat[upper.tri(dmat, diag = F)])
r_value_freq_table <- ds_freq_table(r_value_df, r_value, 10)
print(r_value_freq_table) 

# P-value distribution
hist_p_value <- hist(dres1[upper.tri(dmat, diag = F)], xlab = "Estimated p-value", 
     main = "Histogram of Dcifer estimated significance of pairwise relatedness", breaks = 50) #p-value
p_value_df <- data.frame(p_value = dres1[upper.tri(dmat, diag = F)])
p_value_freq_table <- ds_freq_table(p_value_df, p_value, 10)
print(p_value_freq_table)

# Significant relatedness distribution
isig <- which(dres_aa <= 0.05, arr.ind = TRUE)[, 2:1] 
isig <- isig[row_filter]

hist(dmat[isig], breaks = 50,  xlab = "Significant relatedness", 
     main = "Histogram of significant relatedness")
sig_r_df <- data.frame(sig_r = dmat[isig])
sig_r_freq_table <- ds_freq_table(sig_r_df, sig_r, 10)
print(sig_r_freq_table)



# Define the function for plotting the network
plot_network_per_site <- function(long_ibd_cnv, relatedness_threshold = 0.5, pvalue_threshold = 0.05, 
                         scaling_factor = 1.0, jitter_amount = 0.1, 
                         show_isolates = TRUE, edge_color_by = same_site, vertex_size = 5) {
  
  # Filter edges based on relatedness (ibd) and p-value thresholds
  filtered_edges <- long_ibd %>%
    filter(ibd >= relatedness_threshold, p_BH < pvalue_threshold)
  # Create a data frame of unique sample names for vertices
  unique_samples <- unique(c(filtered_edges$sample.x, filtered_edges$sample.y))
  
  # Create an igraph object from the filtered edges
  g <- graph_from_data_frame(filtered_edges, directed = FALSE, vertices = data.frame(name = unique_samples))

  # Identify isolates (samples that are not in the edges)
  samples_in_edges <- unique(c(filtered_edges$sample.x, filtered_edges$sample.y))
  all_samples <- unique(c(long_ibd$sample.x, long_ibd$sample.y))
  isolates <- setdiff(all_samples, samples_in_edges)

  # Add isolates as vertices in the graph if show_isolates is TRUE
  if (show_isolates) {
    g_full <- g + vertices(isolates)
  } else {
    g_full <- g  # Only use the main graph without isolates
  }

  # # Merge metadata to get site information for coloring vertices
  vertex_data <- data.frame(name = V(g_full)$name) %>%
    mutate(site = case_when(
      grepl(pattern = "AA", name) ~ "Asaiyta",
      grepl(pattern = "AC", name) ~ "Gambella",
      grepl(pattern = "GM", name) ~ "Gondar",
      grepl(pattern = "BA", name) ~ "Assossa",
      grepl(pattern = "ARC", name) ~ "Djibouti",
      grepl(pattern = "MM", name) ~ "Mizan"))
  #   left_join(metadata_cluster, by = "name")  # Updated to use metadata_cluster

  # Check for NA in site and set default color if necessary
  if ("site" %in% colnames(vertex_data)) {
    # Set vertex colors based on site
    unique_populations <- unique(vertex_data$site)
    colors <- rainbow(length(unique_populations))  # Create a color palette

    # Map colors to sites
    color_map <- setNames(colors, unique_populations)

    # Assign colors to vertices based on site
    V(g_full)$color <- color_map[as.character(vertex_data$site)]
    
    # Assign default color for any samples without a site
    V(g_full)$color[is.na(V(g_full)$color)] <- "lightgray"  # Default color for unmatched samples
  } else {
    V(g_full)$color <- "gray"  # Default color if site is not found
  }

  # Set vertex size
  V(g_full)$size <- vertex_size

  # Color edges based on the specified category (now "cat")
  if (edge_color_by %in% colnames(filtered_edges)) {
    edge_categories <- filtered_edges %>%
      select(sample.x, sample.y, !!sym(edge_color_by)) %>%
      distinct()

    # Ensure there are no NA values in the edge categories
    edge_categories <- edge_categories[!is.na(edge_categories[[edge_color_by]]), ]

    if (nrow(edge_categories) > 0) {
      # Manually set edge colors based on "same_site"
      E(g_full)$color <- ifelse(edge_categories[[edge_color_by]] == "Same site", "darkgoldenrod", 
                                ifelse(edge_categories[[edge_color_by]] == "Different site", "grey", NA))
      
      # Set a default color for any unmatched edges
      E(g_full)$color[is.na(E(g_full)$color)] <- "black"  # Use black for NA values
    } else {
      E(g_full)$color <- "black"  # If no edge categories, default to black
    }
  } else {
    E(g_full)$color <- "black"  # Default edge color if category not found
  }

  # Set edge widths to ensure visibility
  edge_widths <- pmax(E(g_full)$ibd * 5, 1)

  # Use layout_nicely() to automatically select an appropriate layout
  layout <- layout_nicely(g_full)

  # Jitter the layout coordinates to avoid overlap
  jitter_matrix <- matrix(rnorm(nrow(layout) * 2, mean = 0, sd = jitter_amount), ncol = 2)
  layout <- layout + jitter_matrix

  # Plot the graph with the new layout
  plot(g_full, 
       layout = layout, 
       vertex.label = NA,  # Remove labels
       vertex.size = V(g_full)$size,  # Use modifiable vertex sizes
       edge.width = edge_widths,  # Use adjusted edge widths
       edge.color = E(g_full)$color,  # Use colors for edges
       main = paste("Relatedness Threshold:", 
                    relatedness_threshold, ", p-value Threshold:", pvalue_threshold))

  # Add a legend for the vertex colors (site)
  legend("bottomright", legend = names(color_map), col = color_map, pch = 19, pt.cex = 1.5, title = "site")
  # Add a legend for the edge colors (Category)
  legend("topright", legend = c("Same site", "Different site"), col = c("darkgoldenrod", "grey"), lty = 1, title = edge_color_by)
  
}

clus_p_aa1 <- plotClusters(dcifer_list, level=0.9, alpha=0.05, vSize=5, vColor='darkred', thick=3, vFontSize=.5, labelOffset=.5, showLabel = F) # level = relatedness cutoff, alpha = p-value cutoff

clus_p_aa2 <- plotClusters(dcifer_list, level=0.75, alpha=0.05, vSize=5, vColor='darkred', thick=3, vFontSize=.5, labelOffset=.5, showLabel = F) 

clus_p_aa3 <- plotClusters(dcifer_list, level=0.5, alpha=0.05, vSize=5, vColor='darkred', thick=3, vFontSize=.5, labelOffset=.5, showLabel = F) 



```

### (old) Cluster plot: Assossa

```{r cluster_ba, exercise=TRUE, echo=F}

# Make dcifer_list for cluster plot
dres0 <- dres_ba[, , "p_value"]

# Identify rows and columns with "D0"
# Filter dres0
row_filter <- grepl("D0", rownames(dres0))
col_filter <- grepl("D0", colnames(dres0))
dres0 <- dres0[row_filter, col_filter, drop = FALSE]
dres0[upper.tri(dres0, diag = F)] <- 0
dres1 <- dres0 + t(dres0) 

# Filter dmat
dmat <- dres_ba[, , "estimate"]
dmat <- dmat[row_filter, col_filter, drop = FALSE]
dmat[upper.tri(dmat)] <- t(dmat)[upper.tri(t(dmat))]

isSymmetric.matrix(dres1) #TRUE
isSymmetric.matrix(dmat) #TRUE

dcifer_list = list(id=colnames(dmat), relatedness=dmat, pvalue=dres1)

# Relatedness distribution
hist_relatedness <- hist(dmat[upper.tri(dmat, diag = F)], xlab = "Estimated pairwise relatedness", 
     main = "Histogram of Dcifer estimated pairwise relatedness", breaks = 50) 
r_value_df <- data.frame(r_value = dmat[upper.tri(dmat, diag = F)])
r_value_freq_table <- ds_freq_table(r_value_df, r_value, 10)
print(r_value_freq_table) 

# P-value distribution
hist_p_value <- hist(dres1[upper.tri(dmat, diag = F)], xlab = "Estimated p-value", 
     main = "Histogram of Dcifer estimated significance of pairwise relatedness", breaks = 50) #p-value
p_value_df <- data.frame(p_value = dres1[upper.tri(dmat, diag = F)])
p_value_freq_table <- ds_freq_table(p_value_df, p_value, 10)
print(p_value_freq_table)

# Significant relatedness distribution
isig <- which(dres_ba <= 0.05, arr.ind = TRUE)[, 2:1] 
isig <- isig[row_filter]

hist(dmat[isig], breaks = 50,  xlab = "Significant relatedness", 
     main = "Histogram of significant relatedness")
sig_r_df <- data.frame(sig_r = dmat[isig])
sig_r_freq_table <- ds_freq_table(sig_r_df, sig_r, 10)
print(sig_r_freq_table)

# Use same plotClusters() function

clus_p_ba1 <- plotClusters(dcifer_list, level=0.9, alpha=0.05, vSize=5, vColor='darkorange', thick=3, vFontSize=.5, labelOffset=.5, showLabel = F) # level = relatedness cutoff, alpha = p-value cutoff

clus_p_ba2 <- plotClusters(dcifer_list, level=0.75, alpha=0.05, vSize=5, vColor='darkorange', thick=3, vFontSize=.5, labelOffset=.5, showLabel = F) 

clus_p_ba3 <- plotClusters(dcifer_list, level=0.5, alpha=0.05, vSize=5, vColor='darkorange', thick=3, vFontSize=.5, labelOffset=.5, showLabel = F) 


```

### (old) Cluster plot: Gambella

```{r cluster_ac, exercise=TRUE, echo=F}

# Make dcifer_list for cluster plot
dres0 <- dres_ac[, , "p_value"]

# Identify rows and columns with "D0"
# Filter dres0
row_filter <- grepl("D0", rownames(dres0))
col_filter <- grepl("D0", colnames(dres0))
dres0 <- dres0[row_filter, col_filter, drop = FALSE]
dres0[upper.tri(dres0, diag = F)] <- 0
dres1 <- dres0 + t(dres0) 

# Filter dmat
dmat <- dres_ac[, , "estimate"]
dmat <- dmat[row_filter, col_filter, drop = FALSE]
dmat[upper.tri(dmat)] <- t(dmat)[upper.tri(t(dmat))]

isSymmetric.matrix(dres1) #TRUE
isSymmetric.matrix(dmat) #TRUE

dcifer_list = list(id=colnames(dmat), relatedness=dmat, pvalue=dres1)

# Relatedness distribution
hist_relatedness <- hist(dmat[upper.tri(dmat, diag = F)], xlab = "Estimated pairwise relatedness", 
     main = "Histogram of Dcifer estimated pairwise relatedness", breaks = 50) 
r_value_df <- data.frame(r_value = dmat[upper.tri(dmat, diag = F)])
r_value_freq_table <- ds_freq_table(r_value_df, r_value, 10)
print(r_value_freq_table) 

# P-value distribution
hist_p_value <- hist(dres1[upper.tri(dmat, diag = F)], xlab = "Estimated p-value", 
     main = "Histogram of Dcifer estimated significance of pairwise relatedness", breaks = 50) #p-value
p_value_df <- data.frame(p_value = dres1[upper.tri(dmat, diag = F)])
p_value_freq_table <- ds_freq_table(p_value_df, p_value, 10)
print(p_value_freq_table)

# Significant relatedness distribution
isig <- which(dres_ac <= 0.05, arr.ind = TRUE)[, 2:1] 
isig <- isig[row_filter]

hist(dmat[isig], breaks = 50,  xlab = "Significant relatedness", 
     main = "Histogram of significant relatedness")
sig_r_df <- data.frame(sig_r = dmat[isig])
sig_r_freq_table <- ds_freq_table(sig_r_df, sig_r, 10)
print(sig_r_freq_table)

# Use same plotClusters() function

clus_p_ac1 <- plotClusters(dcifer_list, level=0.9, alpha=0.05, vSize=5, vColor='yellow3', thick=3, vFontSize=.5, labelOffset=.5, showLabel = F) # level = relatedness cutoff, alpha = p-value cutoff

clus_p_ac2 <- plotClusters(dcifer_list, level=0.75, alpha=0.05, vSize=5, vColor='yellow3', thick=3, vFontSize=.5, labelOffset=.5, showLabel = F) 

clus_p_ac3 <- plotClusters(dcifer_list, level=0.5, alpha=0.05, vSize=5, vColor='yellow3', thick=3, vFontSize=.5, labelOffset=.5, showLabel = F) 


```

### (old) Cluster plot: Gondar

```{r cluster_gm, exercise=TRUE, echo=F}

# Make dcifer_list for cluster plot
dres0 <- dres_gm[, , "p_value"]

# Identify rows and columns with "D0"
# Filter dres0
row_filter <- grepl("D0", rownames(dres0))
col_filter <- grepl("D0", colnames(dres0))
dres0 <- dres0[row_filter, col_filter, drop = FALSE]
dres0[upper.tri(dres0, diag = F)] <- 0
dres1 <- dres0 + t(dres0) 

# Filter dmat
dmat <- dres_gm[, , "estimate"]
dmat <- dmat[row_filter, col_filter, drop = FALSE]
dmat[upper.tri(dmat)] <- t(dmat)[upper.tri(t(dmat))]

isSymmetric.matrix(dres1) #TRUE
isSymmetric.matrix(dmat) #TRUE

dcifer_list = list(id=colnames(dmat), relatedness=dmat, pvalue=dres1)

# Relatedness distribution
hist_relatedness <- hist(dmat[upper.tri(dmat, diag = F)], xlab = "Estimated pairwise relatedness", 
     main = "Histogram of Dcifer estimated pairwise relatedness", breaks = 50) 
r_value_df <- data.frame(r_value = dmat[upper.tri(dmat, diag = F)])
r_value_freq_table <- ds_freq_table(r_value_df, r_value, 10)
print(r_value_freq_table) 

# P-value distribution
hist_p_value <- hist(dres1[upper.tri(dmat, diag = F)], xlab = "Estimated p-value", 
     main = "Histogram of Dcifer estimated significance of pairwise relatedness", breaks = 50) #p-value
p_value_df <- data.frame(p_value = dres1[upper.tri(dmat, diag = F)])
p_value_freq_table <- ds_freq_table(p_value_df, p_value, 10)
print(p_value_freq_table)

# Significant relatedness distribution
isig <- which(dres_gm <= 0.05, arr.ind = TRUE)[, 2:1] 
isig <- isig[row_filter]

hist(dmat[isig], breaks = 50,  xlab = "Significant relatedness", 
     main = "Histogram of significant relatedness")
sig_r_df <- data.frame(sig_r = dmat[isig])
sig_r_freq_table <- ds_freq_table(sig_r_df, sig_r, 10)
print(sig_r_freq_table)

# Use same plotClusters() function

clus_p_gm1 <- plotClusters(dcifer_list, level=0.9, alpha=0.05, vSize=5, vColor='darkgreen', thick=3, vFontSize=.5, labelOffset=.5, showLabel = F) # level = relatedness cutoff, alpha = p-value cutoff

clus_p_gm2 <- plotClusters(dcifer_list, level=0.75, alpha=0.05, vSize=5, vColor='darkgreen', thick=3, vFontSize=.5, labelOffset=.5, showLabel = F) 

clus_p_gm3 <- plotClusters(dcifer_list, level=0.5, alpha=0.05, vSize=5, vColor='darkgreen', thick=3, vFontSize=.5, labelOffset=.5, showLabel = F) 


```

### (old) Cluster plot: Mizan

```{r cluster_mm, exercise=TRUE, echo=F}

# Make dcifer_list for cluster plot
dres0 <- dres_mm[, , "p_value"]

# Identify rows and columns with "D0"
# Filter dres0
row_filter <- grepl("D0", rownames(dres0))
col_filter <- grepl("D0", colnames(dres0))
dres0 <- dres0[row_filter, col_filter, drop = FALSE]
dres0[upper.tri(dres0, diag = F)] <- 0
dres1 <- dres0 + t(dres0) 

# Filter dmat
dmat <- dres_mm[, , "estimate"]
dmat <- dmat[row_filter, col_filter, drop = FALSE]
dmat[upper.tri(dmat)] <- t(dmat)[upper.tri(t(dmat))]

isSymmetric.matrix(dres1) #TRUE
isSymmetric.matrix(dmat) #TRUE

dcifer_list = list(id=colnames(dmat), relatedness=dmat, pvalue=dres1)

# Relatedness distribution
hist_relatedness <- hist(dmat[upper.tri(dmat, diag = F)], xlab = "Estimated pairwise relatedness", 
     main = "Histogram of Dcifer estimated pairwise relatedness", breaks = 50) 
r_value_df <- data.frame(r_value = dmat[upper.tri(dmat, diag = F)])
r_value_freq_table <- ds_freq_table(r_value_df, r_value, 10)
print(r_value_freq_table) 

# P-value distribution
hist_p_value <- hist(dres1[upper.tri(dmat, diag = F)], xlab = "Estimated p-value", 
     main = "Histogram of Dcifer estimated significance of pairwise relatedness", breaks = 50) #p-value
p_value_df <- data.frame(p_value = dres1[upper.tri(dmat, diag = F)])
p_value_freq_table <- ds_freq_table(p_value_df, p_value, 10)
print(p_value_freq_table)

# Significant relatedness distribution
isig <- which(dres_mm <= 0.05, arr.ind = TRUE)[, 2:1] 
isig <- isig[row_filter]

hist(dmat[isig], breaks = 50,  xlab = "Significant relatedness", 
     main = "Histogram of significant relatedness")
sig_r_df <- data.frame(sig_r = dmat[isig])
sig_r_freq_table <- ds_freq_table(sig_r_df, sig_r, 10)
print(sig_r_freq_table)

# Use same plotClusters() function

clus_p_mm1 <- plotClusters(dcifer_list, level=0.9, alpha=0.05, vSize=5, vColor='blue', thick=3, vFontSize=.5, labelOffset=.5, showLabel = F) # level = relatedness cutoff, alpha = p-value cutoff

clus_p_mm2 <- plotClusters(dcifer_list, level=0.75, alpha=0.05, vSize=5, vColor='blue', thick=3, vFontSize=.5, labelOffset=.5, showLabel = F) 

clus_p_mm3 <- plotClusters(dcifer_list, level=0.5, alpha=0.05, vSize=5, vColor='blue', thick=3, vFontSize=.5, labelOffset=.5, showLabel = F) 

```

### (old) Cluster plot: Djibouti

```{r cluster_arc, exercise=TRUE, echo=F}

# Make dcifer_list for cluster plot
dres0 <- dres_arc[, , "p_value"]

# Identify rows and columns with "D0"
# Filter dres0
row_filter <- grepl("D0", rownames(dres0))
col_filter <- grepl("D0", colnames(dres0))
dres0 <- dres0[row_filter, col_filter, drop = FALSE]
dres0[upper.tri(dres0, diag = F)] <- 0
dres1 <- dres0 + t(dres0) 

# Filter dmat
dmat <- dres_arc[, , "estimate"]
dmat <- dmat[row_filter, col_filter, drop = FALSE]
dmat[upper.tri(dmat)] <- t(dmat)[upper.tri(t(dmat))]

isSymmetric.matrix(dres1) #TRUE
isSymmetric.matrix(dmat) #TRUE

dcifer_list = list(id=colnames(dmat), relatedness=dmat, pvalue=dres1)

# Relatedness distribution
hist_relatedness <- hist(dmat[upper.tri(dmat, diag = F)], xlab = "Estimated pairwise relatedness", 
     main = "Histogram of Dcifer estimated pairwise relatedness", breaks = 50) 
r_value_df <- data.frame(r_value = dmat[upper.tri(dmat, diag = F)])
r_value_freq_table <- ds_freq_table(r_value_df, r_value, 10)
print(r_value_freq_table) 

# P-value distribution
hist_p_value <- hist(dres1[upper.tri(dmat, diag = F)], xlab = "Estimated p-value", 
     main = "Histogram of Dcifer estimated significance of pairwise relatedness", breaks = 50) #p-value
p_value_df <- data.frame(p_value = dres1[upper.tri(dmat, diag = F)])
p_value_freq_table <- ds_freq_table(p_value_df, p_value, 10)
print(p_value_freq_table)

# Significant relatedness distribution
isig <- which(dres_arc <= 0.05, arr.ind = TRUE)[, 2:1] 
isig <- isig[row_filter]

hist(dmat[isig], breaks = 50,  xlab = "Significant relatedness", 
     main = "Histogram of significant relatedness")
sig_r_df <- data.frame(sig_r = dmat[isig])
sig_r_freq_table <- ds_freq_table(sig_r_df, sig_r, 10)
print(sig_r_freq_table)

# Use same plotClusters() function

clus_p_arc1 <- plotClusters(dcifer_list, level=0.9, alpha=0.05, vSize=5, vColor='darkviolet', thick=3, vFontSize=.5, labelOffset=.5, showLabel = F) # level = relatedness cutoff, alpha = p-value cutoff

clus_p_arc2 <- plotClusters(dcifer_list, level=0.75, alpha=0.05, vSize=5, vColor='darkviolet', thick=3, vFontSize=.5, labelOffset=.5, showLabel = F) 

clus_p_arc3 <- plotClusters(dcifer_list, level=0.5, alpha=0.05, vSize=5, vColor='darkviolet', thick=3, vFontSize=.5, labelOffset=.5, showLabel = F) 

```

### (old) Extract clustered sampleIDs

```{r extract clusters, eval=FALSE, exercise=TRUE, include=FALSE}

extract_clusters <- function(cluster_plot) {
  
g <- igraph::max_cliques(cluster_plot) #This cluster_list would link to the level you set for cluster_plot (in this case would report samples with R>=0.9)
cluster_list <- lapply(g, names)

result_list <- list()
for (sublist in cluster_list) {
  if (length(sublist) > 1) {
    result_list <- append(result_list, list(sublist))
  }
}

### Merge sublists with overlapping elements
merge_overlap <- function(lst) {
  # Create an empty list to store merged sublists
  merged_list <- list()
  
  for (sublist in lst) {
    # Check if the sublist has any overlapping elements with existing merged sublists
    overlapping_index <- sapply(merged_list, function(merged_sublist) any(sublist %in% merged_sublist))
    
    if (any(overlapping_index)) {
      # Merge the sublist with overlapping sublists
      merged_sublist <- unique(unlist(c(sublist, merged_list[overlapping_index])))
      
      # Remove the overlapping sublists
      merged_list <- merged_list[!overlapping_index]
      
      # Add the merged sublist to the merged_list
      merged_list <- append(merged_list, list(merged_sublist))
    } else {
      # Add the original sublist to the merged_list
      merged_list <- append(merged_list, list(sublist))
    }
  }
  
  return(merged_list)
}

# Merge sublists with overlapping elements
merged_result_list <- merge_overlap(result_list)

### Turn the list into dataframe
# Initialize empty vectors to store values and sublist numbers
all_values <- c()
sublist_number <- c()

# Iterate through the merged_result_list
for (i in seq_along(merged_result_list)) {
  values <- merged_result_list[[i]]
  
  # Extend the all_values vector with the values from the current sublist
  all_values <- c(all_values, values)
  
  # Extend the sublist_number vector with the corresponding sublist number for each value
  sublist_number <- c(sublist_number, rep(i, length(values)))
}

# Create a dataframe from the vectors
cluster_df <- data.frame(Values = all_values, SublistNumber = sublist_number)
cluster_df %<>% rename(sampleID = Values, cluster = SublistNumber)
cluster_df %<>% arrange(cluster, sampleID)

return(cluster_df)

}

clusterID <- extract_clusters(cluster_plot)
write_csv(clusterID, "cluster_id.csv")

```

