---
title: "Moire diagnostics, Dcifer & P-adjustments"
author: "William Louie"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

## Load Libraries

```{r load libraries, warning=FALSE, message=FALSE}

library(moire)
library(dcifer)
library(dplyr)
library(magrittr)
library(ggplot2)
library(igraph)
library(ggnewscale)
library(descriptr)
library(readr)
library(plotly)
library(stringr)
library(ggpubr)
library(purrr)
library(ggbeeswarm)
library(gtsummary)
library(gt)
library(patchwork)
library(grid)
library(tidyverse)
```

## Import data
Import & manage pre- and post-moire data

```{r import, echo=FALSE, message=FALSE, warning=FALSE}

# setwd("/Users/williamlouie/Dropbox/My Mac (Williams-MacBook-Pro.local)/Downloads/Ethiopia/HAMMS/hamms_moire_NEW")

moire_input <- read_csv("C:/Users/Yisihak/Desktop/tes_multiple_run_result/moire/TES_moire_input_filtered_updated.csv")

# Split into sitess
moire_input_aa <- moire_input %>%
  filter(site == "Asayita")
moire_input_ac <- moire_input %>%
  filter(site == "Gambella")
# moire_input_arc <- moire_input %>%
#   filter(site == "Djibouti")
moire_input_ba <- moire_input %>%
  filter(site == "Benshangul")
moire_input_gm <- moire_input %>%
  filter(site == "Gondar")
moire_input_mm <- moire_input %>%
  filter(site == "Mizan")

# Load long form
dat <- moire::load_long_form_data(moire_input)
dat_aa <- moire::load_long_form_data(moire_input_aa)
dat_ac <- moire::load_long_form_data(moire_input_ac)
# dat_arc <- moire::load_long_form_data(moire_input_arc)
dat_ba <- moire::load_long_form_data(moire_input_ba)
dat_gm <- moire::load_long_form_data(moire_input_gm)
dat_mm <- moire::load_long_form_data(moire_input_mm)

# Import moire results
moire_results_aa <- readRDS("C:/Users/Yisihak/Desktop/tes_multiple_run_result/moire/result/TES_moire_result_updated_aa.rds")
moire_results_ac <- readRDS("C:/Users/Yisihak/Desktop/tes_multiple_run_result/moire/result/TES_moire_result_updated_ac.rds")
# moire_results_arc <- readRDS("./filt/filt_moire_results_arc.rds")
moire_results_ba <- readRDS("C:/Users/Yisihak/Desktop/tes_multiple_run_result/moire/result/TES_moire_result_updated_ba.rds")
moire_results_gm <- readRDS("C:/Users/Yisihak/Desktop/tes_multiple_run_result/moire/result/TES_moire_result_updated_gm.rds")
moire_results_mm <- readRDS("C:/Users/Yisihak/Desktop/tes_multiple_run_result/moire/result/TES_moire_result_updated_mm.rds")

# Filter uninformative loci
filtered_dat <- moire_input %>% filter(! locus %in% dat$uninformative_loci)

filtered_dat_aa <- moire_input_aa %>% filter(! locus %in% dat_aa$uninformative_loci)
setdiff(moire_results_aa[["args"]][["data"]][["loci"]], unique(filtered_dat_aa$locus))

filtered_dat_ac <- moire_input_ac %>% filter(! locus %in% dat_ac$uninformative_loci)
setdiff(moire_results_ac[["args"]][["data"]][["loci"]], unique(filtered_dat_ac$locus))

# filtered_dat_arc <- moire_input_arc %>% filter(! locus %in% dat_arc$uninformative_loci)
# setdiff(moire_results_arc[["args"]][["data"]][["loci"]], unique(filtered_dat_arc$locus))

filtered_dat_ba <- moire_input_ba %>% filter(! locus %in% dat_ba$uninformative_loci)
setdiff(moire_results_ba[["args"]][["data"]][["loci"]], unique(filtered_dat_ba$locus))

filtered_dat_gm <- moire_input_gm %>% filter(! locus %in% dat_gm$uninformative_loci)
setdiff(moire_results_gm[["args"]][["data"]][["loci"]], unique(filtered_dat_gm$locus))

filtered_dat_mm <- moire_input_mm %>% filter(! locus %in% dat_mm$uninformative_loci)
setdiff(moire_results_mm[["args"]][["data"]][["loci"]], unique(filtered_dat_mm$locus))

```



```{r}

coi_sum <- read_csv("C:/Users/Yisihak/Desktop/tes_multiple_run_result/moire/coi_summary_upated.csv")

result <- coi_sum %>%
  group_by(site) %>%
  summarise(IQR = IQR(post_coi_mean, na.rm = TRUE))

print(result)

```




# Post-moire exploration

**MCMC diagnostics**
```{r moire chain diagnostics, echo=FALSE, warning = FALSE, message= FALSE}

# Import files
moire_results <- readr::read_rds("C:/Users/Yisihak/Desktop/tes_multiple_run_result/moire/result/TES_moire_input_filtered_updated_acPfPHAST.rds")
# qplot(1:1e4,moire_results$chains[[1]]$llik_sample,geom = "line")
moire_data <- read.csv("C:/Users/Yisihak/Desktop/tes_multiple_run_result/moire/TES_moire_input_filtered_updated_acPfPHAST.csv", header = T)

dat <- moire::load_long_form_data(moire_data)

# Convergence
plot_posterior_llik <- function(moire_results) {
  # does not currently support multiple chains
  posterior_burnin <- moire_results$chains[[1]]$posterior_burnin
  posterior_sample <- moire_results$chains[[1]]$posterior_sample
  prior_burnin <- moire_results$chains[[1]]$prior_burnin
  prior_sample <- moire_results$chains[[1]]$prior_sample
  likelihood_burnin <- moire_results$chains[[1]]$llik_burnin
  likelihood_sample <- moire_results$chains[[1]]$llik_sample
  total = length(posterior_burnin) + length(posterior_sample)
  dat <- rbind(
    data.frame(llik = "posterior", value = c(posterior_burnin, posterior_sample), sample = 1:total),
    data.frame(llik = "prior", value = c(prior_burnin, prior_sample), sample = 1:total),
    data.frame(llik = "likelihood", value = c(likelihood_burnin, likelihood_sample), sample = 1:total)
  )
  g <- ggplot(dat, aes(x = sample, y = value, color = llik, group = llik)) +
    geom_line() +
    geom_vline(xintercept = length(posterior_burnin), linetype = "dashed", alpha = .5)
  g
}

ggplotly(plot_posterior_llik(moire_results))

# Chain swap
plot_chain_swap_dist <- function(moire_results) {
  swap_dist <- moire_results$chains[[1]]$swap_acceptances / moire_results$args$samples_per_chain
  temps <- moire_results$chains[[1]]$temp_gradient
  swap_idx <- (temps[1:length(temps)-1] + temps[2:length(temps)]) / 2
  dat <- data.frame(swap_rate = swap_dist, temp = swap_idx)
  g <- ggplot(dat, aes(x = temp, y = swap_rate)) +
    geom_point() +
    geom_vline(data = data.frame(x = temps), aes(xintercept=x), linetype = "dashed", alpha = 0.25) +
    coord_cartesian(ylim=c(0,1))
  g
}
ggplotly(plot_chain_swap_dist(moire_results))

#Check eps_pos/eps_neg distributions
plot_eps_pos <- function(moire_results) {
  eps_pos <- moire::summarize_epsilon_pos(moire_results)

  g <- ggplot(eps_pos, aes(x = sample_id, y = post_eps_pos_med,
                           ymin = post_eps_pos_lower,
                           ymax = post_eps_pos_upper)) +
    geom_errorbar() +
    geom_point() +
    theme(axis.text.x = element_blank())

  g
}

plot_eps_neg <- function(moire_results) {
  eps_neg <- moire::summarize_epsilon_neg(moire_results)

  g <- ggplot(eps_neg, aes(x = sample_id, y = post_eps_neg_med,
                           ymin = post_eps_neg_lower,
                           ymax = post_eps_neg_upper)) +
    geom_errorbar() +
    geom_point() +
    theme(axis.text.x = element_blank())

  g
}

ggplotly(plot_eps_pos(moire_results))
ggplotly(plot_eps_neg(moire_results))

```

## Summarize COI

```{r moire COI_eCOI, echo=FALSE, warning = FALSE, message= FALSE} 

# Summarize COI for each site
coi_aa <- moire::summarize_coi(moire_results_aa)
coi_ac <- moire::summarize_coi(moire_results_ac)
# coi_arc <- moire::summarize_coi(moire_results_arc)
coi_ba <- moire::summarize_coi(moire_results_ba)
coi_gm <- moire::summarize_coi(moire_results_gm)
coi_mm <- moire::summarize_coi(moire_results_mm)
coi_summary <- rbind(coi_aa, coi_ac, coi_ba, coi_gm, coi_mm)

coi_summary <- coi_summary %<>%
    mutate(site = case_when(
    grepl(pattern = "AA", sample_id) ~ "Asayita",
    grepl(pattern = "AC", sample_id) ~ "Gambella",
    grepl(pattern = "GM", sample_id) ~ "Gondar",
    grepl(pattern = "BA", sample_id) ~ "Assossa",
    grepl(pattern = "MM", sample_id) ~ "Mizan")) 


# Summarize eCOI for each site
ecoi_aa <- moire::summarize_effective_coi(moire_results_aa)
ecoi_ac <- moire::summarize_effective_coi(moire_results_ac)
# ecoi_arc <- moire::summarize_effective_coi(moire_results_arc)
ecoi_ba <- moire::summarize_effective_coi(moire_results_ba)
ecoi_gm <- moire::summarize_effective_coi(moire_results_gm)
ecoi_mm <- moire::summarize_effective_coi(moire_results_mm)
ecoi_summary <- rbind(ecoi_aa, ecoi_ac, ecoi_ba, ecoi_gm, ecoi_mm)
rm(ecoi_aa, ecoi_ac, ecoi_arc, ecoi_ba, ecoi_gm, ecoi_mm)

coi_summary <- coi_summary %>%
  full_join(ecoi_summary, by = "sample_id")

ecoi_summary %<>%
  mutate(site = case_when(
    grepl(pattern = "AA", sample_id) ~ "Asayita",
    grepl(pattern = "AC", sample_id) ~ "Gambella",
    grepl(pattern = "GM", sample_id) ~ "Gondar",
    grepl(pattern = "BA", sample_id) ~ "Assossa",
    grepl(pattern = "MM", sample_id) ~ "Mizan")) 

# write_csv(coi_summary, "C:/Users/Yisihak/Desktop/tes_multiple_run_result/moire/coi_summary_upated.csv")

print("What is the distribution of median COI?")
# Plot individual coi median
ggplot(coi_summary, aes(x = sample_id, y = post_coi_med,
                           ymin = post_coi_lower,
                           ymax = post_coi_upper)) +
    geom_errorbar() +
    geom_point() +
    theme_classic(base_size = 10) +
    theme(axis.text.x = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank()) +
    facet_wrap(~site, scales = "free") +
    ggtitle("Median COI, by site")

print("What is the distribution of mean COI?")
# Plot individual coi mean
ggplot(coi_summary, aes(x=post_coi_mean,
                            ymin = post_coi_lower,
                            ymax = post_coi_upper)) + 
  geom_histogram(bins = 20) +
  scale_x_continuous(breaks=seq(1,30,1)) +
  geom_vline(aes(xintercept = mean(post_coi_mean)), color = "gray", linetype ='dashed') +
  xlab("Mean COI") +
  ylab ("Count") +
  ggtitle("Distribution of Mean COI") +
  theme_classic() +
  facet_wrap(~site, scales = "free") +
  ggtitle("Mean COI, by site")


print("Plot distribution of median effective COI")
# Plot individual effective coi median
ggplot(coi_summary, aes(x = sample_id, y = post_effective_coi_med,
                           ymin = post_effective_coi_lower,
                           ymax = post_effective_coi_upper)) +
    geom_errorbar() +
    geom_point() +
    theme_classic(base_size = 10) +
    theme(axis.text.x = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank()) +
    facet_wrap(~site, scales = "free") +
    ggtitle("Median effective COI, by site")

print("Plot distribution of mean effective COI")
# Plot individual effective coi mean
ggplot(ecoi_summary, aes(x=post_effective_coi_mean)) + 
  geom_histogram(bins = 20) +
  scale_x_continuous(breaks=seq(1,15,1)) +
  geom_vline(aes(xintercept = mean(post_effective_coi_mean)), color = "gray", linetype = "dashed") +
  xlab("Mean Effective COI") + 
  ylab("Count") +
  ggtitle("Distribution of Mean Effective COI") +
  theme_classic() +
  facet_wrap(~site, scales = "free") +
  ggtitle("Mean effective COI, by site")


```


### COI dot plot

```{r}

## COI plot of Day 0s only
coi_summary_d0 <- coi_summary %>%
  filter(grepl("D0", sample_id))

coi_arrange <- coi_summary_d0 %>%
  mutate(mean_COI = as.numeric(post_coi_mean)) %>%
  mutate(median_COI = as.numeric(post_coi_med)) %>%
  mutate(mean_eCOI = as.numeric(post_effective_coi_mean)) %>%
  mutate(median_eCOI = as.numeric(post_effective_coi_med)) %>%
  select(sample_id, mean_COI, median_COI, mean_eCOI, median_eCOI, site) %>%
  mutate(median_COI = paste0(median_COI, ".00")) %>%
  mutate(median_COI = as.numeric(median_COI)) 



# COI table summary
table_coi <- coi_arrange |>
  tbl_summary(by = site, 
    include = c(mean_COI, mean_eCOI),
    statistic = list(
      mean_COI ~ "{mean} ({sd})",
      mean_eCOI ~ "{mean} ({sd})")) %>%
  add_p()
table_coi



# Save as gt
## We will combine this with the plot later
gt_table <- table_coi %>%
  as_gt()


# Site colors
site_color <- c("Gondar" = "blue","Asayita" = "darkred","Assossa" = "darkorange", "Mizan" = "darkviolet","Gambella" = "yellow3" )

# Site order
site_order <- c("Asayita", "Assossa", "Gambella", "Gondar", "Mizan")
coi_summary_d0$site <- factor(coi_summary_d0$site, levels = site_order)

# Calc the standard deviations
coi_summary_d0_sd <- coi_summary_d0 %>%
  group_by(site) %>%
  dplyr::summarize(
    mean = mean(post_coi_mean),
    sd = sd(post_coi_mean),
    ymin = mean - sd,
    ymax = mean + sd,
    .groups = 'drop') %>%
  mutate(site = factor(site, levels = site_order))

# Use coi_summary
g_coi <- ggplot(coi_summary_d0, 
  aes(x = site, y = post_coi_mean)) +
  geom_point(aes(color = site), position = "jitter") +
  geom_boxplot(aes(color = site, fill = site), alpha = 0.1, coef = 0, outlier.shape = NA) +
  geom_errorbar(data = coi_summary_d0_sd, aes(y = mean, ymin = ymin, ymax = ymax, x = site), width = 0.5,
                color = "black") +
  geom_segment(data = coi_summary_d0_sd, 
               aes(x = as.numeric(site) - 0.25, 
                   xend = as.numeric(site) + 0.25, 
                   y = mean, yend = mean), 
               inherit.aes = FALSE, color = "black", size = 1) +
  theme_minimal() +
  scale_y_continuous(breaks = seq(1, 10, by = 1)) +
  ylab("posterior mean COI") +
  scale_color_manual(values = site_color) +
  scale_fill_manual(values = site_color) +
  theme(axis.title.x = element_blank(),
        axis.text = element_text(size = 14),
        legend.text = element_text(size = 16),
      panel.grid = element_blank(), 
      axis.line = element_line(size = 1.1))
g_coi

ggplotly(g_coi)

library(plotly)

# Do the stats
comparisons <- compare_means(post_coi_mean ~ site, data = coi_summary_d0, method = "wilcox.test")
# Add a y.position column to set where the significance bars should appear
y_max <- max(coi_summary_d0$post_coi_mean)
spacing <- (y_max * 0.02)  # Space lines at 2% of the maximum y value
comparisons <- comparisons %>%
  mutate(y.position = y_max + (seq(0.5, nrow(comparisons)) - 1) * spacing)

g_coi_stat <- g_coi + stat_pvalue_manual(
  comparisons,
  method = "wilcox.test",
  label = "p.signif", hide.ns = TRUE,
  size = 6)
g_coi_stat

# Save the gt table as an image
# gtsave(gt_table, filename = "Figure_d0_moi_distribution_101925.png", path = "./")
# # Read the saved table image as a grob (graphical object)
# table_grob <- grid::rasterGrob(png::readPNG(file.path("./Figure_d0_moi_distribution_101925.png")), interpolate = TRUE)
# # Combine the plot and table using patchwork
# combined_plot <- g_coi_stat + table_grob + plot_layout(ncol = 1, heights = c(2, 1))
# # Final save
# ggsave("./Figure_d0_moi_distribution_101925.pdf", plot = combined_plot, width = 9, height = 11, dpi = 300, device = "pdf")

# Sys.setenv(
#   CHROMOTE_CHROME = "C:/Program Files/Google/Chrome/Application/chrome.exe"
# )



```



### Make COI plot

```{r coi_plot, echo=FALSE, warning = FALSE, message= FALSE}

## COI plot of Day 0s only
coi_summary_d0 <- coi_summary %>%
  filter(grepl("D0", sample_id))

coi_arrange <- coi_summary_d0 %>%
  mutate(mean_COI = as.numeric(post_coi_mean)) %>%
  mutate(median_COI = as.numeric(post_coi_med)) %>%
  mutate(mean_eCOI = as.numeric(post_effective_coi_mean)) %>%
  mutate(median_eCOI = as.numeric(post_effective_coi_med)) %>%
  select(sample_id, mean_COI, median_COI, mean_eCOI, median_eCOI, naive_coi, site) %>%
  mutate(median_COI = paste0(median_COI, ".00")) %>%
  mutate(median_COI = as.numeric(median_COI)) 

# pdf("C:/Users/Yisihak/Pictures/TES/coi_all_site_table.pdf", width = 12, height = 10)
# COI table summary
table_coi <- coi_arrange |>
  tbl_summary(by = site, 
    include = c(mean_COI, mean_eCOI),
    statistic = list(
      mean_COI ~ "{mean} ({sd})",
      mean_eCOI ~ "{mean} ({sd})")) %>%
  add_p()
table_coi
# dev.off()

# Save as gt
## We will combine this with the plot later
gt_table <- table_coi %>%
  as_gt()


# Site colors
site_color <- c("Asayita" = "darkred", "Gambella" = "yellow3", "Gondar" = "darkgreen",
                "Benshangul" = "darkorange", "Mizan" = "blue")


# Site order
site_order <- c("Asayita", "Benshangul", "Gambella", "Gondar", "Mizan")
coi_summary_d0$site <- factor(coi_summary_d0$site, levels = site_order)

# Calc the standard deviations
coi_summary_d0_sd <- coi_summary_d0 %>%
  group_by(site) %>%
  summarize(
    mean = mean(post_coi_mean),
    sd = sd(post_coi_mean),
    ymin = mean - sd,
    ymax = mean + sd,
    .groups = 'drop') %>%
  mutate(site = factor(site, levels = site_order))

# # Use coi_summary
# g_coi <- ggplot(coi_summary_d0, 
#   aes(x = site, y = post_coi_mean)) +
#   geom_point(aes(color = site), position = "jitter") +
#   geom_boxplot(aes(color = site, fill = site), alpha = 0.1, coef = 0, outlier.shape = NA) +
#   geom_errorbar(data = coi_summary_d0_sd, aes(y = mean, ymin = ymin, ymax = ymax, x = site), width = 0.5,
#                 color = "black") +
#   geom_segment(data = coi_summary_d0_sd, 
#                aes(x = as.numeric(site) - 0.25, 
#                    xend = as.numeric(site) + 0.25, 
#                    y = mean, yend = mean), 
#                inherit.aes = FALSE, color = "black", size = 1) +
#   theme_minimal() +
#   scale_y_continuous(breaks = seq(1, 10, by = 1)) +
#   ylab("posterior mean COI") +
#   scale_color_manual(values = site_color) +
#   scale_fill_manual(values = site_color) +
#   theme(axis.title.x = element_blank(),
#         axis.text = element_text(size = 14),
#         legend.text = element_text(size = 16),
#       panel.grid = element_blank(), 
#       axis.line = element_line(size = 1.1))
# g_coi

# pdf("C:/Users/Yisihak/Desktop/tes_multiple_run_result/moire/coi.pdf", width = 12, height = 10)

g_coi_hist <- ggplot(coi_summary_d0) +
  geom_histogram(
    aes(x = post_coi_mean, fill = site),
    binwidth = 1,
    color = "black",
    alpha = 1) +
  geom_histogram(
    aes(x = post_effective_coi_mean),
    binwidth = 0.5,
    color = "darkgrey",
    fill = "grey",       # Set color to grey
    alpha = 0.5) +
  theme_classic() +
  theme(
    axis.text.x = element_text(vjust = 0, hjust = 1),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14)) +
  geom_vline(
    data = coi_summary_d0_sd,
    aes(xintercept = mean),
    color = "black",
    linetype = "dashed",
    size = 0.7) +
  scale_fill_manual(values = site_color, guide = "none") +
  ylab("Sample count") +
  xlab("COI / eCOI") +
  facet_wrap(~site) +
  geom_text(
    data = coi_summary_d0_sd,
    aes(x = mean + 0.4, y = 50, label = round(mean, 2)),
    color = "black",
    size = 4, vjust = -1)
g_coi_hist


# dev.off()

# Do the stats
comparisons <- compare_means(post_coi_mean ~ site, data = coi_summary_d0, method = "wilcox.test")
# Add a y.position column to set where the significance bars should appear
y_max <- max(coi_summary_d0$post_coi_mean)
spacing <- (y_max * 0.02)  # Space lines at 2% of the maximum y value
comparisons <- comparisons %>%
  mutate(y.position = y_max + (seq(0.5, nrow(comparisons)) - 1) * spacing)

# g_coi_stat <- g_coi + stat_pvalue_manual(
#   comparisons,
#   method = "wilcox.test",
#   label = "p.signif", hide.ns = TRUE,
#   size = 6)
# g_coi_stat

# # Save the gt table as an image
# gtsave(gt_table, filename = "C:/Users/Yisihak/Pictures/TES/coi_all_site_table.pdf")
# # Read the saved table image as a grob (graphical object)
# table_grob <- grid::rasterGrob(png::readPNG(file.path(tempdir(), "Figure_d0_moi_distribution_120224.png")), interpolate = TRUE)
# # Combine the plot and table using patchwork
# combined_plot <- g_coi_hist + table_grob + plot_layout(ncol = 1, heights = c(2, 1))
# # Final save
# ggsave("hamms_moire_NEW/Figure_d0_moi_distribution_120224.jpeg", plot = combined_plot, width = 9, height = 11, dpi = 300, device = "jpeg")

```

### other plots

```{r other_plot, echo=FALSE, warning = FALSE, message= FALSE}

# Import metadata
metadata <- read_csv("/Users/williamlouie/Dropbox/My Mac (Williams-MacBook-Pro.local)/Downloads/Ethiopia/HAMMS/hamms_metadata/processed_sample_metadata_WL.csv") %>%
  rename("sample_id" = "study_id_crf")

coi_combined <- coi_summary %>%
  left_join(metadata, by = "sample_id") %>%
  mutate(parasitemia_log = log10(pf_par_ul)) %>%
  select(sample_id, post_coi_mean, post_effective_coi_mean, parasitemia_log)

# Plot correlation between coi, eff_coi, and parasitemias
max_coi <- round(max(coi_combined$post_coi_mean))
g_corr <- ggplot(coi_combined, 
  aes(x = post_coi_mean, y = post_effective_coi_mean, label = sample_id)) +
  geom_point(aes(color = parasitemia_log)) +
  geom_abline(slope = 1, intercept = 0, linetype = "solid", color = "gray") +
  scale_x_continuous(expand = c(0,0), limits = c(0, max_coi + 1), breaks = seq(from = 1, to = max_coi + 1, 1)) +
  scale_y_continuous(expand = c(0,0), limits = c(0, max_coi + 1), breaks = seq(from = 1, to = max_coi + 1, 1)) +
  theme_classic() +
  scale_colour_gradient(low = "blue", high = "red", na.value = NA, name = " log10 p/ul") +
  xlab("COI") +
  ylab("eCOI") +
  theme_classic()
g_corr
ggsave("Figure_coi_correlation.jpeg", plot = g_corr, height = 6, width = 7)

print("Correlation between Naive COI and Moire COI, for polyclonals only")

# Filter out monoclonals
coi_pol <- coi_summary %>%
  filter(naive_coi > 1) %>%
  arrange(post_coi_med)

max_coi <- round(max(coi_pol$post_coi_mean))

# Plot correlation between moire COI and naive COI
ggplot(coi_pol, 
  aes(x = naive_coi, y = post_coi_mean, label = sample_id)) +
  geom_beeswarm(aes(color = site)) +
  geom_abline(slope = 1, intercept = 0, linetype = "solid", color = "gray") +
  scale_x_continuous(expand = c(0,0), limits = c(0, 8), breaks = seq(from = 1, to = max_coi + 1, 1)) +
  scale_y_continuous(expand = c(0,0), limits = c(0, max_coi + 1), breaks = seq(from = 1, to = max_coi + 1, 1)) +
  theme_classic() +
  xlab("Naive COI") +
  ylab("Moire COI") +
  theme_classic() +
  facet_wrap(~site)



print("Correlation between COI and effective COI, for polyclonals only")

# Extract polyclonal samples - defined by eCOI_cutoffs
sample_data_select <- coi_summary %>% 
  filter(post_effective_coi_mean >= 1.1)

# Plot correlation between coi, eff_coi, and relatedness
g <- ggplot(sample_data_select, 
  aes(x = post_coi_mean, y = post_effective_coi_mean, label = sample_id)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, linetype = "solid", color = "gray") +
  scale_x_continuous(expand = c(0,0), limits = c(0, max_coi + 1), breaks = seq(from = 1, to = max_coi + 1, 1)) +
  scale_y_continuous(expand = c(0,0), limits = c(0, max_coi + 1), breaks = seq(from = 1, to = max_coi + 1, 1)) +
  theme_classic() +
  scale_colour_gradient(low = "blue", high = "red", na.value = NA, name = "rw") +
  xlab("COI") +
  ylab("eCOI") +
  theme_classic()
g

ibs_d0 <- sample_data_select %>%
  filter(grepl("D0", sample_id))

```

## Allele frequencies
Set up allele frequencies: we need to average out the allele frequencies of all sites in order to run dcifer

```{r allele_freq, echo=FALSE}

# We need to average out allele frequencies for the dcifer input

# Export allele frequencies too
af_AA <- moire::summarize_allele_freqs(moire_results_aa)
af_AC <- moire::summarize_allele_freqs(moire_results_ac)
# af_ARC <- moire::summarize_allele_freqs(moire_results_arc)
af_BA <- moire::summarize_allele_freqs(moire_results_ba)
af_GM <- moire::summarize_allele_freqs(moire_results_gm)
af_MM <- moire::summarize_allele_freqs(moire_results_mm)


# Sample sizes
AA_size <- 91
AC_size <- 96
# ARC_size <- 85
BA_size <- 87
GM_size <- 97
MM_size <- 108


# sample_size <- AA_size + AC_size + BA_size + GM_size + MM_size

# Combine all af
af_list <- list(
  af_AA %>% mutate(site = "AA"),
  af_AC %>% mutate(site = "AC"),
  # af_ARC %>% mutate(site = "ARC"),
  af_BA %>% mutate(site = "BA"),
  af_GM %>% mutate(site = "GM"),
  af_MM %>% mutate(site = "MM")
)
# Combine all allele frequency data frames
af_combined <- bind_rows(af_list)
# Coalesce the allele frequencies from both dataframes, treating NAs as 0


# Calculate weighted averages
# Normalize allele frequencies per locus
af_combined <- af_combined %>%
  group_by(locus) %>%
  mutate(
    # Normalize allele frequencies within each locus (so they sum to 1 for each locus)
    total_freq = sum(coalesce(post_allele_freqs_med, 0), na.rm = TRUE),
    normalized_allele_freq = coalesce(post_allele_freqs_med, 0) / total_freq
  )

# Now calculate the weighted average of the normalized allele frequencies across all sites
af_weighted <- af_combined %>%
  group_by(allele, locus) %>%
  summarise(
    # Weighted average frequency calculation
    weighted_avg_freq = sum(normalized_allele_freq * sample_size, na.rm = TRUE) / sum(sample_size, na.rm = TRUE),
    .groups = "drop"
  )

# Check the sum of frequencies by locus (should sum to 1 for each locus)
af_check <- af_weighted %>%
  group_by(locus) %>%
  summarise(sum_af = sum(weighted_avg_freq))


# Check that all alleles are accounted for
## Get unique strings from both data frames
unique_filtered <- unique(filtered_dat$allele)
unique_table_merge <- unique(af_weighted$allele)
missing_from_dat <- setdiff(unique_table_merge, unique_filtered)
missing_from_afreq <- setdiff(unique_filtered, unique_table_merge)

# write_csv(af_weighted, "hamms_dcifer_NEW/hamms_avg_allele_freq.csv")

# Format allele frequencies (avg) for Dcifer
afreq <- formatAfreq(af_weighted, lvar = "locus", avar = "allele", fvar = "weighted_avg_freq")
# write_rds(afreq, "hamms_dcifer_NEW/hamms_avg_allele_freq.rds")

```

Heterozygosity and FST/ Jost's D (popultion differentiation). The diversity within the population. With COI, moire can summarize the posterior distribution to provide quantile estimates of He.

```{r heterozygosity, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

# Calculate Heterozygosity (He)

# Calc He for AA
he_aa <-summarize_he(moire_results_aa)
he_aa$site <- "Asayita"
hist(he_aa$post_stat_mean)
he_aa_mean <- mean(he_aa$post_stat_mean)

# Calc He for AC
he_ac <-summarize_he(moire_results_ac)
he_ac$site <- "Gambella"
hist(he_ac$post_stat_mean)
he_ac_mean <- mean(he_ac$post_stat_mean)

# Calc He for ARC
# he_arc <-summarize_he(moire_results_arc)
# he_arc$site <- "Djibouti"
# hist(he_arc$post_stat_mean)
# he_arc_mean <- mean(he_arc$post_stat_mean)

# Calc He for BA
he_ba <-summarize_he(moire_results_ba)
he_ba$site <- "Assossa"
hist(he_ba$post_stat_mean)
he_ba_mean <- mean(he_ba$post_stat_mean)

# Calc He for GM
he_gm <- moire::summarize_he(moire_results_gm)
he_gm$site <- "Gondar"
hist(he_gm$post_stat_mean)
he_gm_mean <- mean(he_gm$post_stat_mean)

# Calc He for MM
he_mm <- moire::summarize_he(moire_results_mm)
he_mm$site <- "Mizan"
hist(he_mm$post_stat_mean)
he_mm_mean <- mean(he_mm$post_stat_mean)

he_merge <- rbind(he_aa, he_ac, he_ba, he_gm, he_mm)
# Do stats. Kruskal-Wallis test across all sites
# We will add this to the histogram
he_merge_summmary <- he_merge %>%
  group_by(site) %>%
  summarise(mean_he = mean(post_stat_med),
            median_he = median(post_stat_med))
library(dunn.test)
he_stats <- he_merge %>%
  summarise(kruskal_pvalue = kruskal.test(post_stat_med ~ site, data = cur_data())$p.value, dunn_test_results = list(
      dunn.test(
        x = post_stat_med, 
        g = site, 
        method = "bh")), .groups = "drop")
he_stats_pairwise <- he_stats  %>%
  filter(!map_lgl(dunn_test_results, is.null)) %>%  # Filter out null results
  mutate(
    pairwise_comparisons = map(dunn_test_results, ~ {
      # Extract 'comparisons' and 'P.adjusted' from each test result
      data.frame(
        comparison = .x$comparisons, 
        p_adj = .x$P.adjusted
      )
    })
  )

# Flatten the list of pairwise comparisons and ensure each result is in a separate row
he_stats_pairwise <- bind_rows(
  map(he_stats_pairwise$pairwise_comparisons, as.data.frame),
  .id = "site")
he_stats_pairwise %<>%
  mutate(significant = case_when(p_adj < 0.05 ~ "yes",
                                 p_adj > 0.05 ~ "no"))

# write_csv(he_merge, "hamms_moire_NEW/heterozygosity_all_loci.csv")
# write_csv(he_merge_summmary, "hamms_moire_NEW/heterozygosity_summary.csv")
# write_csv(he_stats_pairwise, "hamms_moire_NEW/heterozygosity_stats.csv")


# He plot
# Add median and stats
g_he <- ggplot(he_merge) +
  geom_histogram(aes(y = post_stat_med, fill = site)) +
  theme_classic() +
  theme(axis.text.x = element_text(vjust = 0, hjust = 1),
        axis.text = element_text(size = 12),
        legend.position = "none") +
  geom_hline(data = he_merge_summmary, aes(yintercept = median_he), color = "black", linetype = "dashed", size = 0.8) +
  geom_text(data = he_merge_summmary, 
            aes(x = 18, y = median_he, 
                label = ifelse(p_value < 0.05, paste0(round(median_he, 2), "*")), 
            vjust = -0.5, size = 3)) +
  scale_fill_manual(values = site_color) +
  xlab("Locus count") +
  ylab("Locus Heterozygosity") +
  facet_wrap(~site)
g_he
# ggsave("hamms_moire_NEW/Figure_heterozygosity.jpeg", plot = g_he, height = 8, width = 10)


#####---------- NOT READY --------------########
# # Calculating and plotting FST
# library(polysat)
# 
# # formatting allele frequency data to match structure required for package
# myfreq_aa <- AF_table_aa %>% select(mean_AF, allele) %>% pivot_wider(names_from = allele, values_from = mean_AF)
# myfreq_aa$Genomes<-length(dat_aa$sample_ids)
# myfreq_aa$Pop<-"Asayita"
# 
# myfreq_ac <- AF_table_ac %>% select(mean_AF, allele) %>% pivot_wider(names_from = allele, values_from = mean_AF)
# myfreq_ac$Genomes<-length(dat_ac$sample_ids)
# myfreq_ac$Pop<-"Gambella"
# 
# myfreq_merge<-rbind(myfreq_aa, myfreq_ac)
#                                                              
# # calculate pairwise differentiation statistics)
# myfst <- calcPopDiff(myfreq_merge, metric = "Fst")
# myD <- calcPopDiff(myfreq_merge, metric = "Jost's D")
# 
# # examine the results
# myfst
# myD


```

## Dcifer - Asayita
*Heatmap shows relatedness between D0s only

```{r}

# === FORMAT DATA ===
mydsmp <- dcifer::formatDat(filtered_dat_aa, "sample_id", "locus", "allele")
dsmp <- dcifer::matchAfreq(mydsmp, afreq)

# Combined COI, sorted correctly
coi <- coi_summary$post_coi_med
ecoi <- coi_summary$post_effective_coi_med

# === Run Dcifer ===
dres_rp <- ibdDat(dsmp, coi, afreq, pval = TRUE, confint = TRUE, rnull = 0.125, alpha = 0.01, nr = 1e3)

# === Relatedness matrix ===
dmat <- dres_rp[, , "estimate"]
dmat[upper.tri(dmat)] <- t(dmat)[upper.tri(t(dmat))]

# === Adjust p-values ===
dres0 <- dres_rp[, , "p_value"]
pval_vec <- dres0[lower.tri(dres0)]                           # raw p-values
pval_adj <- p.adjust(pval_vec, method = "BH")                 # [MODIFIED]: BH-adjustment
adjusted_pmat <- matrix(NA, nrow = nrow(dres0), ncol = ncol(dres0),
                        dimnames = dimnames(dres0))           # [MODIFIED]
adjusted_pmat[lower.tri(adjusted_pmat)] <- pval_adj          # [MODIFIED]
adjusted_pmat[upper.tri(adjusted_pmat)] <- t(adjusted_pmat)[upper.tri(adjusted_pmat)]  # symmetric

# === Categorize adjusted p-values ===
dres_low <- -log10(adjusted_pmat)                            # [MODIFIED]
dres_low[dres_low > -log10(0.0001)] <- "P << 0.001"
dres_low[dres_low <= -log10(0.0001) & dres_low > -log10(0.001)] <- "P < 0.001"
dres_low[dres_low <= -log10(0.001) & dres_low > -log10(0.01)] <- "P < 0.01"
dres_low[dres_low <= -log10(0.01)] <- "Not Sig"
dres_low[upper.tri(dres_low)] <- t(dres_low)[upper.tri(dres_low)]

# === Matrix to long format ===
matrix_to_longformat <- function(mat) {
  as.data.frame(mat) |>
    rownames_to_column("sample.x") |>
    pivot_longer(-sample.x, names_to = "sample.y", values_to = "value")
}

to_plot1 <- matrix_to_longformat(dmat)

# === Clustering for ordering ===
hc <- hclust(as.dist(1 - dmat))
order <- hc$order
clustered_levels <- colnames(dmat)[order]

to_plot1 <- to_plot1 |>
  mutate(sample.x = fct_relevel(sample.x, clustered_levels),
         sample.y = fct_relevel(sample.y, clustered_levels))

to_plot2 <- matrix_to_longformat(dres_low)
to_plot2 <- to_plot2 |>
  mutate(sample.x = fct_relevel(sample.x, clustered_levels),
         sample.y = fct_relevel(sample.y, clustered_levels))

# === Define colors ===
colours_upper <- colorRampPalette(c("#FFFFFF", "#1520A6"))(100)
colours_lower <- c("P << 0.001" = "#C70039", "P < 0.001" = "#FF5733", "P < 0.01" = "#FFC300", "Not Sig" = "white")

# === Filter for D0 samples ===
to_plot1_0 <- to_plot1[grepl("D0", to_plot1$sample.x) & grepl("D0", to_plot1$sample.y), ]
to_plot2_0 <- to_plot2[grepl("D0", to_plot2$sample.x) & grepl("D0", to_plot2$sample.y), ]

sample_d0 <- unique(to_plot1_0$sample.x)
cluster_0 <- clustered_levels[clustered_levels %in% sample_d0]

to_plot1_0 <- to_plot1_0 |>
  mutate(sample.x = fct_relevel(sample.x, cluster_0),
         sample.y = fct_relevel(sample.y, cluster_0))

to_plot2_0 <- to_plot2_0 |>
  mutate(sample.x = fct_relevel(sample.x, cluster_0),
         sample.y = fct_relevel(sample.y, cluster_0))

# pdf("C:/Users/Yisihak/Pictures/Djbouti figures/Djibouti_figure_relatedness_pvalue_corrected.pdf", width = 10, height = 10)

# === Heatmap ===
g0_aa <- ggplot(to_plot1_0, aes(x = sample.x, y = sample.y)) +
  geom_tile(data = subset(to_plot1_0, as.numeric(sample.x) > as.numeric(sample.y)), aes(fill = value)) +
  scale_fill_gradientn(colors = colours_upper, name = "Relatedness") +
  theme(axis.text.x = element_blank()) +
  theme(axis.text.y = element_blank()) +
  new_scale_fill() +
  geom_tile(data = subset(to_plot2_0, as.numeric(sample.x) < as.numeric(sample.y)),
            aes(fill = value)) +
  scale_fill_manual(values = colours_lower, name = "Significance", 
                    breaks = c("P << 0.001", "P < 0.001", "P < 0.01")) +
  scale_x_discrete(limits = cluster_0) +
  scale_y_discrete(limits = cluster_0) +
  ggtitle("Djibouti (BH adjusted pvalue)")

g0_aa

dev.off()

####################################################################################################################################


# === Categorize unadjusted p-values ===
dres_unadj <- -log10(dres0)
dres_unadj[dres_unadj > -log10(0.0001)] <- "P << 0.0001"
dres_unadj[dres_unadj <= -log10(0.0001) & dres_unadj > -log10(0.001)] <- "P < 0.001"
dres_unadj[dres_unadj <= -log10(0.001) & dres_unadj > -log10(0.01)] <- "P < 0.01"
dres_unadj[dres_unadj <= -log10(0.01)] <- "Not Sig"
dres_unadj[upper.tri(dres_unadj)] <- t(dres_unadj)[upper.tri(dres_unadj)]

# === Convert to long format ===
to_plot2_unadj <- matrix_to_longformat(dres_unadj)

# === Reorder by clustering ===
to_plot2_unadj <- to_plot2_unadj |>
  dplyr::mutate(sample.x = forcats::fct_relevel(sample.x, clustered_levels),
                sample.y = forcats::fct_relevel(sample.y, clustered_levels))

# === Filter D0 samples ===
to_plot2_unadj_0 <- to_plot2_unadj[grepl("D0", to_plot2_unadj$sample.x) & grepl("D0", to_plot2_unadj$sample.y), ]
to_plot2_unadj_0 <- to_plot2_unadj_0 |>
  dplyr::mutate(sample.x = forcats::fct_relevel(sample.x, cluster_0),
                sample.y = forcats::fct_relevel(sample.y, cluster_0))

# === Unadjusted p-value plot ===
g0_unadj <- ggplot(to_plot1_0, aes(x = sample.x, y = sample.y)) +
  geom_tile(data = subset(to_plot1_0, as.numeric(sample.x) > as.numeric(sample.y)), aes(fill = value)) +
  scale_fill_gradientn(colors = colours_upper, name = "Relatedness") +
  theme(axis.text.x = element_blank()) +
  theme(axis.text.y = element_blank()) +
  new_scale_fill() +
  geom_tile(data = subset(to_plot2_unadj_0, as.numeric(sample.x) < as.numeric(sample.y)),
            aes(fill = value)) +
  scale_fill_manual(values = colours_lower, name = "Significance", breaks = c("P << 0.001", "P < 0.001", "P < 0.01")) +
  scale_x_discrete(limits = cluster_0) +
  scale_y_discrete(limits = cluster_0) +
  ggtitle("Djibouti (Unadjusted P-values)")

# === Show or save
g0_unadj

#########################################################################################################################################

# Extract IBD estimates from rnull > 0.125 run
ibd_est_rp5 <- dres_rp[, , "estimate"]
ibd_lower_rp5 <- dres_rp[, , "CI_lower"]
ibd_upper_rp5 <- dres_rp[, , "CI_upper"]

ibd_df_rp5_df  <- as.data.frame(ibd_est_rp5)
ibd_df_rp5_df  <- tibble::rownames_to_column(ibd_df_rp5_df, var = "sample.x")

ibd_df_rp5_lower_df  <- as.data.frame(ibd_lower_rp5)
ibd_df_rp5_lower_df  <- tibble::rownames_to_column(ibd_df_rp5_lower_df, var = "sample.x")

ibd_df_rp5_upper_df  <- as.data.frame(ibd_upper_rp5)
ibd_df_rp5_upper_df  <- tibble::rownames_to_column(ibd_df_rp5_upper_df, var = "sample.x")

#this is a DF with all ibd data in longform between non-matching pairs of samples
ibd_df_long_rp5 <-  ibd_df_rp5_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd"
  ) %>% filter(!is.na(ibd))
hist(ibd_df_long_rp5$ibd)
# CIs for rnull >0.125
CI_lower_df_long_r5 <- ibd_df_rp5_lower_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_lower")
CI_upper_df_long_r5 <- ibd_df_rp5_upper_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_upper")
# bind CIs into long df for rnull > 0.125
ibd_df_long_rp5 <- ibd_df_long_rp5 %>%
  left_join(CI_lower_df_long_r5) %>%
  left_join(CI_upper_df_long_r5)
pvals_r5 <- dres_rp[, , "p_value"]
pvals_r5 <- as.data.frame(pvals_r5)
pvals_r5 <- tibble::rownames_to_column(pvals_r5, var = "sample.x")


# pval 0.05
pvals_df_long_r5 <- pvals_r5 %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "p_value"
  ) %>% filter(!is.na(p_value)) %>%
  mutate(p_value = p_value/2) # divide by 2 for 1 sided test
# bind pvals into long df for rnull > 0.125
ibd_df_long_rp5 <- ibd_df_long_rp5 %>%
  left_join(pvals_df_long_r5)
# make p-value very large if r <0.125, set = 1
ibd_df_long_rp5 <- ibd_df_long_rp5 %>% mutate(p_value = ifelse(ibd <0.125, 1, p_value))
#here is where i do corrections for rnull > 0.5 dataframe
long_ibd_rp5 <- ibd_df_long_rp5 %>%
  mutate(p_BH = p.adjust(p_value, method = "BH"),
         p_BF = p.adjust(p_value, method = "bonferroni")) %>%
  mutate(sig_0.01 = ifelse(p_value < 0.01, 1, 0), sig_0.01_BH = ifelse(p_BH < 0.01, 1, 0),
         sig_0.01_BF = ifelse(p_BF < 0.01, 1, 0)) %>%
  mutate(sig_0.05 = ifelse(p_value < 0.05, 1, 0), sig_0.05_BH = ifelse(p_BH < 0.05, 1, 0),
         sig_0.05_BF = ifelse(p_BF < 0.05, 1, 0))

long_ibd_rp5 %<>% 
  filter(grepl("D0", sample.x)) %>% 
  filter(grepl("D0", sample.y))



# Step 1: Get all related pairs (IBD > 0)
related_pairs <- long_ibd_rp5 %>% filter(ibd > 0.5)

# Step 2: Get all unique samples involved in related pairs
related_samples <- unique(c(related_pairs$sample.x, related_pairs$sample.y))

# Step 3: Get all unique samples in the full dataset
all_samples <- unique(c(long_ibd_rp5$sample.x, long_ibd_rp5$sample.y))

# Step 4: Compute proportion
proportion_related <- length(related_samples) / length(all_samples)

# Show result
proportion_related

# write_csv(long_ibd_rp5, "C:/Users/Yisihak/Desktop/tes_djibouti/filtered data/long_ibd_125_djibouti.csv")


```

```{r dcifer_Asayita, echo=FALSE}

mydsmp <- dcifer::formatDat(filtered_dat_aa, "sample_id", "locus", "allele")
coi   <- coi_aa$post_coi_med
afreq_aa <- dcifer::calcAfreq(mydsmp, coi)

# Match alleles
dsmp <- dcifer::matchAfreq(mydsmp, afreq_aa)

# Run ibdDat
dres_aa <- ibdDat(dsmp, coi, afreq_aa,
                  pval = TRUE, confint = TRUE,
                  rnull = 0.125, alpha = 0.01, nr = 1e3)

## Output dcifer results
# readr::write_rds(dres_aa, "C:/Users/Yisihak/Desktop/tes_multiple_run_result/dcifer/dcifer_result_aa.rds")

# === Relatedness matrix ===
dmat <- dres_aa[, , "estimate"]
dmat[upper.tri(dmat)] <- t(dmat)[upper.tri(t(dmat))]

# === Adjust p-values ===
dres0 <- dres_aa[, , "p_value"]
pval_vec <- dres0[lower.tri(dres0)]                           # raw p-values
pval_adj <- p.adjust(pval_vec, method = "BH")                 # [MODIFIED]: BH-adjustment
adjusted_pmat <- matrix(NA, nrow = nrow(dres0), ncol = ncol(dres0),
                        dimnames = dimnames(dres0))           # [MODIFIED]
adjusted_pmat[lower.tri(adjusted_pmat)] <- pval_adj          # [MODIFIED]
adjusted_pmat[upper.tri(adjusted_pmat)] <- t(adjusted_pmat)[upper.tri(adjusted_pmat)]  # symmetric

# === Categorize adjusted p-values ===
dres_low <- -log10(adjusted_pmat)                            # [MODIFIED]
dres_low[dres_low > -log10(0.0001)] <- "P << 0.001"
dres_low[dres_low <= -log10(0.0001) & dres_low > -log10(0.001)] <- "P < 0.001"
dres_low[dres_low <= -log10(0.001) & dres_low > -log10(0.01)] <- "P < 0.01"
dres_low[dres_low <= -log10(0.01)] <- "Not Sig"
dres_low[upper.tri(dres_low)] <- t(dres_low)[upper.tri(dres_low)]

# === Matrix to long format ===
matrix_to_longformat <- function(mat) {
  as.data.frame(mat) |>
    rownames_to_column("sample.x") |>
    pivot_longer(-sample.x, names_to = "sample.y", values_to = "value")
}

to_plot1 <- matrix_to_longformat(dmat)

# === Clustering for ordering ===
hc <- hclust(as.dist(1 - dmat))
order <- hc$order
clustered_levels <- colnames(dmat)[order]

to_plot1 <- to_plot1 |>
  mutate(sample.x = fct_relevel(sample.x, clustered_levels),
         sample.y = fct_relevel(sample.y, clustered_levels))

to_plot2 <- matrix_to_longformat(dres_low)
to_plot2 <- to_plot2 |>
  mutate(sample.x = fct_relevel(sample.x, clustered_levels),
         sample.y = fct_relevel(sample.y, clustered_levels))

# === Define colors ===
colours_upper <- colorRampPalette(c("#FFFFFF", "#1520A6"))(100)
colours_lower <- c("P << 0.001" = "#C70039", "P < 0.001" = "#FF5733", "P < 0.01" = "#FFC300", "Not Sig" = "white")

# === Filter for D0 samples ===
to_plot1_0 <- to_plot1[grepl("D0", to_plot1$sample.x) & grepl("D0", to_plot1$sample.y), ]
to_plot2_0 <- to_plot2[grepl("D0", to_plot2$sample.x) & grepl("D0", to_plot2$sample.y), ]

sample_d0 <- unique(to_plot1_0$sample.x)
cluster_0 <- clustered_levels[clustered_levels %in% sample_d0]

to_plot1_0 <- to_plot1_0 |>
  mutate(sample.x = fct_relevel(sample.x, cluster_0),
         sample.y = fct_relevel(sample.y, cluster_0))

to_plot2_0 <- to_plot2_0 |>
  mutate(sample.x = fct_relevel(sample.x, cluster_0),
         sample.y = fct_relevel(sample.y, cluster_0))

# pdf("C:/Users/Yisihak/Pictures/Djbouti figures/Djibouti_figure_relatedness_pvalue_corrected.pdf", width = 10, height = 10)

# === Heatmap ===
g0_aa <- ggplot(to_plot1_0, aes(x = sample.x, y = sample.y)) +
  geom_tile(data = subset(to_plot1_0, as.numeric(sample.x) > as.numeric(sample.y)), aes(fill = value)) +
  scale_fill_gradientn(colors = colours_upper, name = "Relatedness") +
  theme(axis.text.x = element_blank()) +
  theme(axis.text.y = element_blank()) +
  new_scale_fill() +
  geom_tile(data = subset(to_plot2_0, as.numeric(sample.x) < as.numeric(sample.y)),
            aes(fill = value)) +
  scale_fill_manual(values = colours_lower, name = "Significance", 
                    breaks = c("P << 0.001", "P < 0.001", "P < 0.01")) +
  scale_x_discrete(limits = cluster_0) +
  scale_y_discrete(limits = cluster_0) +
  ggtitle("Asayita (BH adjusted pvalue)")

g0_aa






# Step 1: Get all related pairs (IBD > 0)
related_pairs <- to_plot1_0 %>% filter(value > 0.5)

# Step 2: Get all unique samples involved in related pairs
related_samples <- unique(c(related_pairs$sample.x, related_pairs$sample.y))

# Step 3: Get all unique samples in the full dataset
all_samples <- unique(c(to_plot1_0$sample.x, to_plot1_0$sample.y))

# Step 4: Compute proportion
proportion_related <- length(related_samples) / length(all_samples)

print("proportion_related")
# Show result
proportion_related





# Extract IBD estimates from rnull > 0.125 run
ibd_est_rp5 <- dres_aa[, , "estimate"]
ibd_lower_rp5 <- dres_aa[, , "CI_lower"]
ibd_upper_rp5 <- dres_aa[, , "CI_upper"]

ibd_df_rp5_df  <- as.data.frame(ibd_est_rp5)
ibd_df_rp5_df  <- tibble::rownames_to_column(ibd_df_rp5_df, var = "sample.x")

ibd_df_rp5_lower_df  <- as.data.frame(ibd_lower_rp5)
ibd_df_rp5_lower_df  <- tibble::rownames_to_column(ibd_df_rp5_lower_df, var = "sample.x")

ibd_df_rp5_upper_df  <- as.data.frame(ibd_upper_rp5)
ibd_df_rp5_upper_df  <- tibble::rownames_to_column(ibd_df_rp5_upper_df, var = "sample.x")

#this is a DF with all ibd data in longform between non-matching pairs of samples
ibd_df_long_rp5 <-  ibd_df_rp5_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd"
  ) %>% filter(!is.na(ibd))
hist(ibd_df_long_rp5$ibd)
# CIs for rnull >0.125
CI_lower_df_long_r5 <- ibd_df_rp5_lower_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_lower")
CI_upper_df_long_r5 <- ibd_df_rp5_upper_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_upper")
# bind CIs into long df for rnull > 0.125
ibd_df_long_rp5 <- ibd_df_long_rp5 %>%
  left_join(CI_lower_df_long_r5) %>%
  left_join(CI_upper_df_long_r5)
pvals_r5 <- dres_aa[, , "p_value"]
pvals_r5 <- as.data.frame(pvals_r5)
pvals_r5 <- tibble::rownames_to_column(pvals_r5, var = "sample.x")


# pval 0.05
pvals_df_long_r5 <- pvals_r5 %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "p_value"
  ) %>% filter(!is.na(p_value)) %>%
  mutate(p_value = p_value/2) # divide by 2 for 1 sided test
# bind pvals into long df for rnull > 0.125
ibd_df_long_rp5 <- ibd_df_long_rp5 %>%
  left_join(pvals_df_long_r5)
# make p-value very large if r <0.125, set = 1
ibd_df_long_rp5 <- ibd_df_long_rp5 %>% mutate(p_value = ifelse(ibd <0.125, 1, p_value))
#here is where i do corrections for rnull > 0.5 dataframe
long_ibd_rp5 <- ibd_df_long_rp5 %>%
  mutate(p_BH = p.adjust(p_value, method = "BH"),
         p_BF = p.adjust(p_value, method = "bonferroni")) %>%
  mutate(sig_0.01 = ifelse(p_value < 0.01, 1, 0), sig_0.01_BH = ifelse(p_BH < 0.01, 1, 0),
         sig_0.01_BF = ifelse(p_BF < 0.01, 1, 0)) %>%
  mutate(sig_0.05 = ifelse(p_value < 0.05, 1, 0), sig_0.05_BH = ifelse(p_BH < 0.05, 1, 0),
         sig_0.05_BF = ifelse(p_BF < 0.05, 1, 0))

long_ibd_rp5 %<>% 
  filter(grepl("D0", sample.x)) %>% 
  filter(grepl("D0", sample.y)) %>% 
  filter(p_BH <= 0.05)

# write.csv(long_ibd_rp5,"C:/Users/Yisihak/Desktop/tes_multiple_run_result/dcifer/long_ibd_125_Asayita.csv", row.names = F)


threshold <- 0.5

# Get all sample names in the dataset
all_samples <- unique(c(long_ibd_rp5$sample.x, long_ibd_rp5$sample.y))

# Identify samples that have IBD >= threshold with at least one other sample
samples_with_ibd <- long_ibd_rp5 %>%
  filter(ibd >= threshold) %>%
  select(sample.x, sample.y) %>%
  tidyr::pivot_longer(cols = everything(), values_to = "sample") %>%
  distinct(sample) %>%
  pull(sample)

# Proportion of samples involved in such pairs
proportion <- length(samples_with_ibd) / length(all_samples)



print(length(all_samples))


print(length(samples_with_ibd))

proportion



```

## Dcifer - Gambella

```{r dcifer_Gambella, echo=FALSE}

mydsmp <- dcifer::formatDat(filtered_dat_ac, "sample_id", "locus", "allele")
coi   <- coi_ac$post_coi_med
afreq_ac <- dcifer::calcAfreq(mydsmp, coi)

dsmp <- dcifer::matchAfreq(mydsmp, afreq_ac)
dres_ac <- ibdDat(dsmp, coi, afreq_ac, pval = TRUE, confint = TRUE, rnull = 0.125, alpha = 0.01, nr = 1e3)

## Output dcifer results
# readr::write_rds(dres_aa, "C:/Users/Yisihak/Desktop/tes_multiple_run_result/dcifer/dcifer_result_ac.rds")

# === Relatedness matrix ===
dmat <- dres_ac[, , "estimate"]
dmat[upper.tri(dmat)] <- t(dmat)[upper.tri(t(dmat))]

# === Adjust p-values ===
dres0 <- dres_ac[, , "p_value"]
pval_vec <- dres0[lower.tri(dres0)]                           # raw p-values
pval_adj <- p.adjust(pval_vec, method = "BH")                 # [MODIFIED]: BH-adjustment
adjusted_pmat <- matrix(NA, nrow = nrow(dres0), ncol = ncol(dres0),
                        dimnames = dimnames(dres0))           # [MODIFIED]
adjusted_pmat[lower.tri(adjusted_pmat)] <- pval_adj          # [MODIFIED]
adjusted_pmat[upper.tri(adjusted_pmat)] <- t(adjusted_pmat)[upper.tri(adjusted_pmat)]  # symmetric

# === Categorize adjusted p-values ===
dres_low <- -log10(adjusted_pmat)                            # [MODIFIED]
dres_low[dres_low > -log10(0.0001)] <- "P << 0.001"
dres_low[dres_low <= -log10(0.0001) & dres_low > -log10(0.001)] <- "P < 0.001"
dres_low[dres_low <= -log10(0.001) & dres_low > -log10(0.01)] <- "P < 0.01"
dres_low[dres_low <= -log10(0.01)] <- "Not Sig"
dres_low[upper.tri(dres_low)] <- t(dres_low)[upper.tri(dres_low)]

# === Matrix to long format ===
matrix_to_longformat <- function(mat) {
  as.data.frame(mat) |>
    rownames_to_column("sample.x") |>
    pivot_longer(-sample.x, names_to = "sample.y", values_to = "value")
}

to_plot1 <- matrix_to_longformat(dmat)

# === Clustering for ordering ===
hc <- hclust(as.dist(1 - dmat))
order <- hc$order
clustered_levels <- colnames(dmat)[order]

to_plot1 <- to_plot1 |>
  mutate(sample.x = fct_relevel(sample.x, clustered_levels),
         sample.y = fct_relevel(sample.y, clustered_levels))

to_plot2 <- matrix_to_longformat(dres_low)
to_plot2 <- to_plot2 |>
  mutate(sample.x = fct_relevel(sample.x, clustered_levels),
         sample.y = fct_relevel(sample.y, clustered_levels))

# === Define colors ===
colours_upper <- colorRampPalette(c("#FFFFFF", "#1520A6"))(100)
colours_lower <- c("P << 0.001" = "#C70039", "P < 0.001" = "#FF5733", "P < 0.01" = "#FFC300", "Not Sig" = "white")

# === Filter for D0 samples ===
to_plot1_0 <- to_plot1[grepl("D0", to_plot1$sample.x) & grepl("D0", to_plot1$sample.y), ]
to_plot2_0 <- to_plot2[grepl("D0", to_plot2$sample.x) & grepl("D0", to_plot2$sample.y), ]

sample_d0 <- unique(to_plot1_0$sample.x)
cluster_0 <- clustered_levels[clustered_levels %in% sample_d0]

to_plot1_0 <- to_plot1_0 |>
  mutate(sample.x = fct_relevel(sample.x, cluster_0),
         sample.y = fct_relevel(sample.y, cluster_0))

to_plot2_0 <- to_plot2_0 |>
  mutate(sample.x = fct_relevel(sample.x, cluster_0),
         sample.y = fct_relevel(sample.y, cluster_0))

# pdf("C:/Users/Yisihak/Pictures/Djbouti figures/Djibouti_figure_relatedness_pvalue_corrected.pdf", width = 10, height = 10)

# === Heatmap ===
g0_ac <- ggplot(to_plot1_0, aes(x = sample.x, y = sample.y)) +
  geom_tile(data = subset(to_plot1_0, as.numeric(sample.x) > as.numeric(sample.y)), aes(fill = value)) +
  scale_fill_gradientn(colors = colours_upper, name = "Relatedness") +
  theme(axis.text.x = element_blank()) +
  theme(axis.text.y = element_blank()) +
  new_scale_fill() +
  geom_tile(data = subset(to_plot2_0, as.numeric(sample.x) < as.numeric(sample.y)),
            aes(fill = value)) +
  scale_fill_manual(values = colours_lower, name = "Significance", 
                    breaks = c("P << 0.001", "P < 0.001", "P < 0.01")) +
  scale_x_discrete(limits = cluster_0) +
  scale_y_discrete(limits = cluster_0) +
  ggtitle("Gamvella (BH adjusted pvalue)")

g0_ac


# Step 1: Get all related pairs (IBD > 0)
related_pairs <- to_plot1_0 %>% filter(value > 0)

# Step 2: Get all unique samples involved in related pairs
related_samples <- unique(c(related_pairs$sample.x, related_pairs$sample.y))

# Step 3: Get all unique samples in the full dataset
all_samples <- unique(c(to_plot1_0$sample.x, to_plot1_0$sample.y))

# Step 4: Compute proportion
proportion_related <- length(related_samples) / length(all_samples)

# Show result
proportion_related





# Extract IBD estimates from rnull > 0.125 run
ibd_est_rp5 <- dres_ac[, , "estimate"]
ibd_lower_rp5 <- dres_ac[, , "CI_lower"]
ibd_upper_rp5 <- dres_ac[, , "CI_upper"]

ibd_df_rp5_df  <- as.data.frame(ibd_est_rp5)
ibd_df_rp5_df  <- tibble::rownames_to_column(ibd_df_rp5_df, var = "sample.x")

ibd_df_rp5_lower_df  <- as.data.frame(ibd_lower_rp5)
ibd_df_rp5_lower_df  <- tibble::rownames_to_column(ibd_df_rp5_lower_df, var = "sample.x")

ibd_df_rp5_upper_df  <- as.data.frame(ibd_upper_rp5)
ibd_df_rp5_upper_df  <- tibble::rownames_to_column(ibd_df_rp5_upper_df, var = "sample.x")

#this is a DF with all ibd data in longform between non-matching pairs of samples
ibd_df_long_rp5 <-  ibd_df_rp5_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd"
  ) %>% filter(!is.na(ibd))
hist(ibd_df_long_rp5$ibd)
# CIs for rnull >0.125
CI_lower_df_long_r5 <- ibd_df_rp5_lower_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_lower")
CI_upper_df_long_r5 <- ibd_df_rp5_upper_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_upper")
# bind CIs into long df for rnull > 0.125
ibd_df_long_rp5 <- ibd_df_long_rp5 %>%
  left_join(CI_lower_df_long_r5) %>%
  left_join(CI_upper_df_long_r5)
pvals_r5 <- dres_ac[, , "p_value"]
pvals_r5 <- as.data.frame(pvals_r5)
pvals_r5 <- tibble::rownames_to_column(pvals_r5, var = "sample.x")


# pval 0.05
pvals_df_long_r5 <- pvals_r5 %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "p_value"
  ) %>% filter(!is.na(p_value)) %>%
  mutate(p_value = p_value/2) # divide by 2 for 1 sided test
# bind pvals into long df for rnull > 0.125
ibd_df_long_rp5 <- ibd_df_long_rp5 %>%
  left_join(pvals_df_long_r5)
# make p-value very large if r <0.125, set = 1
ibd_df_long_rp5 <- ibd_df_long_rp5 %>% mutate(p_value = ifelse(ibd <0.125, 1, p_value))
#here is where i do corrections for rnull > 0.5 dataframe
long_ibd_rp5 <- ibd_df_long_rp5 %>%
  mutate(p_BH = p.adjust(p_value, method = "BH"),
         p_BF = p.adjust(p_value, method = "bonferroni")) %>%
  mutate(sig_0.01 = ifelse(p_value < 0.01, 1, 0), sig_0.01_BH = ifelse(p_BH < 0.01, 1, 0),
         sig_0.01_BF = ifelse(p_BF < 0.01, 1, 0)) %>%
  mutate(sig_0.05 = ifelse(p_value < 0.05, 1, 0), sig_0.05_BH = ifelse(p_BH < 0.05, 1, 0),
         sig_0.05_BF = ifelse(p_BF < 0.05, 1, 0))

long_ibd_rp5 %<>% 
  filter(grepl("D0", sample.x)) %>% 
  filter(grepl("D0", sample.y))


# write.csv(long_ibd_rp5,"C:/Users/Yisihak/Desktop/tes_multiple_run_result/dcifer/long_ibd_125_Abobo.csv", row.names = F)



threshold <- 0.5

# Get all sample names in the dataset
all_samples <- unique(c(long_ibd_rp5$sample.x, long_ibd_rp5$sample.y))

# Identify samples that have IBD >= threshold with at least one other sample
samples_with_ibd <- long_ibd_rp5 %>%
  filter(ibd >= threshold) %>%
  filter(p_BH <= 0.05) %>% 
  select(sample.x, sample.y) %>%
  tidyr::pivot_longer(cols = everything(), values_to = "sample") %>%
  distinct(sample) %>%
  pull(sample)

# Proportion of samples involved in such pairs
proportion <- length(samples_with_ibd) / length(all_samples)


print(length(all_samples))


print(length(samples_with_ibd))


proportion




```

## Dcifer - Djibouti

```{r dcifer_Djibouti, echo=FALSE}

mydsmp <- dcifer::formatDat(filtered_dat_arc, "sample_id", "locus", "allele")
coi   <- coi_arc$post_coi_med
afreq_arc <- calculate_median_afs(moire_results_arc)

## Name allele frequencies
for (i in 1:length(afreq_arc)) {
  names(afreq_arc[[i]]) <- names(moire_results_arc$data$data[[i]][[1]])
  afreq_arc[[i]] = abs(afreq_arc[[i]]) / sum(abs(afreq_arc[[i]]))
}

dsmp <- dcifer::matchAfreq(mydsmp, afreq_arc)
dres_arc <- ibdDat(dsmp, coi, afreq_arc, pval = TRUE, confint = TRUE, rnull = 0.125, alpha = 0.01, nr = 1e3)

# ## Output dcifer results
readr::write_rds(dres_arc, "hamms_dcifer_NEW/dcifer_results_arc.rds")

# relatedness
dmat <- dres_arc[, , "estimate"]
dmat[upper.tri(dmat)] <- t(dmat)[upper.tri(t(dmat))]
# p-value groups
dres0 <- dres_arc[, , "p_value"]
dres_low <- -log10(dres0)

dres_low[dres_low > -log10(0.0001)] <- "P << 0.001"
dres_low[dres_low <= -log10(0.0001) & dres_low > -log10(0.001)] <- "P < 0.001"
dres_low[dres_low <= -log10(0.001) & dres_low > -log10(0.01)] <- "P < 0.01"
dres_low[dres_low <= -log10(0.01)] <- "Not Sig"
dres_low[upper.tri(dres_low)] <- t(dres_low)[upper.tri(dres_low)]



to_plot1_arc <- matrix_to_longformat(dmat)

# dmat, Relatedness    
hc <- hclust(as.dist(1 - dmat))
order = hc$order
clustered_levels = colnames(dmat)[order]

# dres_low, p-value    
to_plot1_arc <- to_plot1_arc |>
  dplyr::mutate(sample.x = forcats::fct_relevel(sample.x, clustered_levels),
                sample.y = forcats::fct_relevel(sample.y, clustered_levels))

to_plot2_arc <- matrix_to_longformat(dres_low)
to_plot2_arc <- to_plot2_arc |>
  dplyr::mutate(sample.x = forcats::fct_relevel(sample.x, clustered_levels),
                sample.y = forcats::fct_relevel(sample.y, clustered_levels))

# Filter for D0s
to_plot1_arc0 <- to_plot1_arc[grepl("D0", to_plot1_arc$sample.x) & grepl("D0", to_plot1_arc$sample.y), ]
to_plot2_arc0 <- to_plot2_arc[grepl("D0", to_plot2_arc$sample.x) & grepl("D0", to_plot2_arc$sample.y), ]
# Reorder the D0 samples
sample_d0 <- unique(to_plot1_arc0$sample.x)
cluster_0 <- clustered_levels[clustered_levels %in% sample_d0]
# levels  
to_plot1_arc0 <- to_plot1_arc0 |>
  dplyr::mutate(sample.x = forcats::fct_relevel(sample.x, cluster_0),
                sample.y = forcats::fct_relevel(sample.y, cluster_0))

to_plot2_arc0 <- to_plot2_arc0 |>
  dplyr::mutate(sample.x = forcats::fct_relevel(sample.x, cluster_0),
                sample.y = forcats::fct_relevel(sample.y, cluster_0))
# D0 Heatmap 
g0_arc <- ggplot(to_plot1_arc0, aes(x = sample.x, y = sample.y)) +
  geom_tile(data = subset(to_plot1_arc0, as.numeric(sample.x) > as.numeric(sample.y)), aes(fill = value)) +
  scale_fill_gradientn(colors = colours_upper, name = "Relatedness") +
  theme(axis.text.x = element_blank()) +
  theme(axis.text.y = element_blank()) +
  new_scale_fill() + # Add the p-value layer
  geom_tile(data = subset(to_plot2_arc0, as.numeric(sample.x) < as.numeric(sample.y)),
            aes(fill = value)) +
  scale_fill_manual(values = colours_lower, name = "Significance", breaks = c("P << 0.001", "P < 0.001", "P < 0.01")) +
  scale_x_discrete(limits = cluster_0) +
  scale_y_discrete(limits = cluster_0) +
  ggtitle("Djibouti")
g0_arc

```

## Dcifer - Assossa/Benshangul

```{r dcifer_Assossa, echo=FALSE}

mydsmp <- dcifer::formatDat(filtered_dat_ba, "sample_id", "locus", "allele")
coi   <- coi_ba$post_coi_med
afreq_ba <- dcifer::calcAfreq(mydsmp, coi)


dsmp <- dcifer::matchAfreq(mydsmp, afreq_ba)
dres_ba <- ibdDat(dsmp, coi, afreq_ba, pval = TRUE, confint = TRUE, rnull = 0.125, alpha = 0.01, nr = 1e3)

## Output dcifer results
# readr::write_rds(dres_aa, "C:/Users/Yisihak/Desktop/tes_multiple_run_result/dcifer/dcifer_result_ba.rds")

# === Relatedness matrix ===
dmat <- dres_ba[, , "estimate"]
dmat[upper.tri(dmat)] <- t(dmat)[upper.tri(t(dmat))]

# === Adjust p-values ===
dres0 <- dres_ba[, , "p_value"]
pval_vec <- dres0[lower.tri(dres0)]                           # raw p-values
pval_adj <- p.adjust(pval_vec, method = "BH")                 # [MODIFIED]: BH-adjustment
adjusted_pmat <- matrix(NA, nrow = nrow(dres0), ncol = ncol(dres0),
                        dimnames = dimnames(dres0))           # [MODIFIED]
adjusted_pmat[lower.tri(adjusted_pmat)] <- pval_adj          # [MODIFIED]
adjusted_pmat[upper.tri(adjusted_pmat)] <- t(adjusted_pmat)[upper.tri(adjusted_pmat)]  # symmetric

# === Categorize adjusted p-values ===
dres_low <- -log10(adjusted_pmat)                            # [MODIFIED]
dres_low[dres_low > -log10(0.0001)] <- "P << 0.001"
dres_low[dres_low <= -log10(0.0001) & dres_low > -log10(0.001)] <- "P < 0.001"
dres_low[dres_low <= -log10(0.001) & dres_low > -log10(0.01)] <- "P < 0.01"
dres_low[dres_low <= -log10(0.01)] <- "Not Sig"
dres_low[upper.tri(dres_low)] <- t(dres_low)[upper.tri(dres_low)]

# === Matrix to long format ===
matrix_to_longformat <- function(mat) {
  as.data.frame(mat) |>
    rownames_to_column("sample.x") |>
    pivot_longer(-sample.x, names_to = "sample.y", values_to = "value")
}

to_plot1 <- matrix_to_longformat(dmat)

# === Clustering for ordering ===
hc <- hclust(as.dist(1 - dmat))
order <- hc$order
clustered_levels <- colnames(dmat)[order]

to_plot1 <- to_plot1 |>
  mutate(sample.x = fct_relevel(sample.x, clustered_levels),
         sample.y = fct_relevel(sample.y, clustered_levels))

to_plot2 <- matrix_to_longformat(dres_low)
to_plot2 <- to_plot2 |>
  mutate(sample.x = fct_relevel(sample.x, clustered_levels),
         sample.y = fct_relevel(sample.y, clustered_levels))

# === Define colors ===
colours_upper <- colorRampPalette(c("#FFFFFF", "#1520A6"))(100)
colours_lower <- c("P << 0.001" = "#C70039", "P < 0.001" = "#FF5733", "P < 0.01" = "#FFC300", "Not Sig" = "white")

# === Filter for D0 samples ===
to_plot1_0 <- to_plot1[grepl("D0", to_plot1$sample.x) & grepl("D0", to_plot1$sample.y), ]
to_plot2_0 <- to_plot2[grepl("D0", to_plot2$sample.x) & grepl("D0", to_plot2$sample.y), ]

sample_d0 <- unique(to_plot1_0$sample.x)
cluster_0 <- clustered_levels[clustered_levels %in% sample_d0]

to_plot1_0 <- to_plot1_0 |>
  mutate(sample.x = fct_relevel(sample.x, cluster_0),
         sample.y = fct_relevel(sample.y, cluster_0))

to_plot2_0 <- to_plot2_0 |>
  mutate(sample.x = fct_relevel(sample.x, cluster_0),
         sample.y = fct_relevel(sample.y, cluster_0))

# pdf("C:/Users/Yisihak/Pictures/Djbouti figures/Djibouti_figure_relatedness_pvalue_corrected.pdf", width = 10, height = 10)

# === Heatmap ===
g0_ba <- ggplot(to_plot1_0, aes(x = sample.x, y = sample.y)) +
  geom_tile(data = subset(to_plot1_0, as.numeric(sample.x) > as.numeric(sample.y)), aes(fill = value)) +
  scale_fill_gradientn(colors = colours_upper, name = "Relatedness") +
  theme(axis.text.x = element_blank()) +
  theme(axis.text.y = element_blank()) +
  new_scale_fill() +
  geom_tile(data = subset(to_plot2_0, as.numeric(sample.x) < as.numeric(sample.y)),
            aes(fill = value)) +
  scale_fill_manual(values = colours_lower, name = "Significance", 
                    breaks = c("P << 0.001", "P < 0.001", "P < 0.01")) +
  scale_x_discrete(limits = cluster_0) +
  scale_y_discrete(limits = cluster_0) +
  ggtitle("Benshangul (BH adjusted pvalue)")

g0_ba

# Step 1: Get all related pairs (IBD > 0)
related_pairs <- to_plot1_0 %>% filter(value > 0)

# Step 2: Get all unique samples involved in related pairs
related_samples <- unique(c(related_pairs$sample.x, related_pairs$sample.y))

# Step 3: Get all unique samples in the full dataset
all_samples <- unique(c(to_plot1_0$sample.x, to_plot1_0$sample.y))

# Step 4: Compute proportion
proportion_related <- length(related_samples) / length(all_samples)

# Show result
proportion_related







# Extract IBD estimates from rnull > 0.125 run
ibd_est_rp5 <- dres_ba[, , "estimate"]
ibd_lower_rp5 <- dres_ba[, , "CI_lower"]
ibd_upper_rp5 <- dres_ba[, , "CI_upper"]

ibd_df_rp5_df  <- as.data.frame(ibd_est_rp5)
ibd_df_rp5_df  <- tibble::rownames_to_column(ibd_df_rp5_df, var = "sample.x")

ibd_df_rp5_lower_df  <- as.data.frame(ibd_lower_rp5)
ibd_df_rp5_lower_df  <- tibble::rownames_to_column(ibd_df_rp5_lower_df, var = "sample.x")

ibd_df_rp5_upper_df  <- as.data.frame(ibd_upper_rp5)
ibd_df_rp5_upper_df  <- tibble::rownames_to_column(ibd_df_rp5_upper_df, var = "sample.x")

#this is a DF with all ibd data in longform between non-matching pairs of samples
ibd_df_long_rp5 <-  ibd_df_rp5_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd"
  ) %>% filter(!is.na(ibd))
hist(ibd_df_long_rp5$ibd)
# CIs for rnull >0.125
CI_lower_df_long_r5 <- ibd_df_rp5_lower_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_lower")
CI_upper_df_long_r5 <- ibd_df_rp5_upper_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_upper")
# bind CIs into long df for rnull > 0.125
ibd_df_long_rp5 <- ibd_df_long_rp5 %>%
  left_join(CI_lower_df_long_r5) %>%
  left_join(CI_upper_df_long_r5)
pvals_r5 <- dres_ba[, , "p_value"]
pvals_r5 <- as.data.frame(pvals_r5)
pvals_r5 <- tibble::rownames_to_column(pvals_r5, var = "sample.x")


# pval 0.05
pvals_df_long_r5 <- pvals_r5 %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "p_value"
  ) %>% filter(!is.na(p_value)) %>%
  mutate(p_value = p_value/2) # divide by 2 for 1 sided test
# bind pvals into long df for rnull > 0.125
ibd_df_long_rp5 <- ibd_df_long_rp5 %>%
  left_join(pvals_df_long_r5)
# make p-value very large if r <0.125, set = 1
ibd_df_long_rp5 <- ibd_df_long_rp5 %>% mutate(p_value = ifelse(ibd <0.125, 1, p_value))
#here is where i do corrections for rnull > 0.5 dataframe
long_ibd_rp5 <- ibd_df_long_rp5 %>%
  mutate(p_BH = p.adjust(p_value, method = "BH"),
         p_BF = p.adjust(p_value, method = "bonferroni")) %>%
  mutate(sig_0.01 = ifelse(p_value < 0.01, 1, 0), sig_0.01_BH = ifelse(p_BH < 0.01, 1, 0),
         sig_0.01_BF = ifelse(p_BF < 0.01, 1, 0)) %>%
  mutate(sig_0.05 = ifelse(p_value < 0.05, 1, 0), sig_0.05_BH = ifelse(p_BH < 0.05, 1, 0),
         sig_0.05_BF = ifelse(p_BF < 0.05, 1, 0))

long_ibd_rp5 %<>% 
  filter(grepl("D0", sample.x)) %>% 
  filter(grepl("D0", sample.y))

# write.csv(long_ibd_rp5,"C:/Users/Yisihak/Desktop/tes_multiple_run_result/dcifer/long_ibd_125_Asosa.csv", row.names = F)


threshold <- 0.5

# Get all sample names in the dataset
all_samples <- unique(c(long_ibd_rp5$sample.x, long_ibd_rp5$sample.y))

# Identify samples that have IBD >= threshold with at least one other sample
samples_with_ibd <- long_ibd_rp5 %>%
  filter(ibd >= threshold) %>%
  filter(p_BH <= 0.05) %>% 
  select(sample.x, sample.y) %>%
  tidyr::pivot_longer(cols = everything(), values_to = "sample") %>%
  distinct(sample) %>%
  pull(sample)

# Proportion of samples involved in such pairs
proportion <- length(samples_with_ibd) / length(all_samples)

print(length(all_samples))


print(length(samples_with_ibd))


proportion



```

## Dcifer - Gondar

```{r dcifer_Gondar, echo=FALSE}

mydsmp <- dcifer::formatDat(filtered_dat_gm, "sample_id", "locus", "allele")
coi   <- coi_gm$post_coi_med
afreq_gm <- dcifer::calcAfreq(mydsmp, coi)

dsmp <- dcifer::matchAfreq(mydsmp, afreq_gm)
dres_gm <- ibdDat(dsmp, coi, afreq_gm, pval = TRUE, confint = TRUE, rnull = 0.125, alpha = 0.01, nr = 1e3)

## Output dcifer results
# readr::write_rds(dres_aa, "C:/Users/Yisihak/Desktop/tes_multiple_run_result/dcifer/dcifer_result_gm.rds")

# === Relatedness matrix ===
dmat <- dres_gm[, , "estimate"]
dmat[upper.tri(dmat)] <- t(dmat)[upper.tri(t(dmat))]

# === Adjust p-values ===
dres0 <- dres_gm[, , "p_value"]
pval_vec <- dres0[lower.tri(dres0)]                           # raw p-values
pval_adj <- p.adjust(pval_vec, method = "BH")                 # [MODIFIED]: BH-adjustment
adjusted_pmat <- matrix(NA, nrow = nrow(dres0), ncol = ncol(dres0),
                        dimnames = dimnames(dres0))           # [MODIFIED]
adjusted_pmat[lower.tri(adjusted_pmat)] <- pval_adj          # [MODIFIED]
adjusted_pmat[upper.tri(adjusted_pmat)] <- t(adjusted_pmat)[upper.tri(adjusted_pmat)]  # symmetric

# === Categorize adjusted p-values ===
dres_low <- -log10(adjusted_pmat)                            # [MODIFIED]
dres_low[dres_low > -log10(0.0001)] <- "P << 0.001"
dres_low[dres_low <= -log10(0.0001) & dres_low > -log10(0.001)] <- "P < 0.001"
dres_low[dres_low <= -log10(0.001) & dres_low > -log10(0.01)] <- "P < 0.01"
dres_low[dres_low <= -log10(0.01)] <- "Not Sig"
dres_low[upper.tri(dres_low)] <- t(dres_low)[upper.tri(dres_low)]

# === Matrix to long format ===
matrix_to_longformat <- function(mat) {
  as.data.frame(mat) |>
    rownames_to_column("sample.x") |>
    pivot_longer(-sample.x, names_to = "sample.y", values_to = "value")
}

to_plot1 <- matrix_to_longformat(dmat)

# === Clustering for ordering ===
hc <- hclust(as.dist(1 - dmat))
order <- hc$order
clustered_levels <- colnames(dmat)[order]

to_plot1 <- to_plot1 |>
  mutate(sample.x = fct_relevel(sample.x, clustered_levels),
         sample.y = fct_relevel(sample.y, clustered_levels))

to_plot2 <- matrix_to_longformat(dres_low)
to_plot2 <- to_plot2 |>
  mutate(sample.x = fct_relevel(sample.x, clustered_levels),
         sample.y = fct_relevel(sample.y, clustered_levels))

# === Define colors ===
colours_upper <- colorRampPalette(c("#FFFFFF", "#1520A6"))(100)
colours_lower <- c("P << 0.001" = "#C70039", "P < 0.001" = "#FF5733", "P < 0.01" = "#FFC300", "Not Sig" = "white")

# === Filter for D0 samples ===
to_plot1_0 <- to_plot1[grepl("D0", to_plot1$sample.x) & grepl("D0", to_plot1$sample.y), ]
to_plot2_0 <- to_plot2[grepl("D0", to_plot2$sample.x) & grepl("D0", to_plot2$sample.y), ]

sample_d0 <- unique(to_plot1_0$sample.x)
cluster_0 <- clustered_levels[clustered_levels %in% sample_d0]

to_plot1_0 <- to_plot1_0 |>
  mutate(sample.x = fct_relevel(sample.x, cluster_0),
         sample.y = fct_relevel(sample.y, cluster_0))

to_plot2_0 <- to_plot2_0 |>
  mutate(sample.x = fct_relevel(sample.x, cluster_0),
         sample.y = fct_relevel(sample.y, cluster_0))

# pdf("C:/Users/Yisihak/Pictures/Djbouti figures/Djibouti_figure_relatedness_pvalue_corrected.pdf", width = 10, height = 10)

# === Heatmap ===
g0_gm <- ggplot(to_plot1_0, aes(x = sample.x, y = sample.y)) +
  geom_tile(data = subset(to_plot1_0, as.numeric(sample.x) > as.numeric(sample.y)), aes(fill = value)) +
  scale_fill_gradientn(colors = colours_upper, name = "Relatedness") +
  theme(axis.text.x = element_blank()) +
  theme(axis.text.y = element_blank()) +
  new_scale_fill() +
  geom_tile(data = subset(to_plot2_0, as.numeric(sample.x) < as.numeric(sample.y)),
            aes(fill = value)) +
  scale_fill_manual(values = colours_lower, name = "Significance", 
                    breaks = c("P << 0.001", "P < 0.001", "P < 0.01")) +
  scale_x_discrete(limits = cluster_0) +
  scale_y_discrete(limits = cluster_0) +
  ggtitle("Gondar (BH adjusted pvalue)")

g0_gm

# Step 1: Get all related pairs (IBD > 0)
related_pairs <- to_plot1_0 %>% filter(value > 0)

# Step 2: Get all unique samples involved in related pairs
related_samples <- unique(c(related_pairs$sample.x, related_pairs$sample.y))

# Step 3: Get all unique samples in the full dataset
all_samples <- unique(c(to_plot1_0$sample.x, to_plot1_0$sample.y))

# Step 4: Compute proportion
proportion_related <- length(related_samples) / length(all_samples)

# Show result
proportion_related







# Extract IBD estimates from rnull > 0.125 run
ibd_est_rp5 <- dres_gm[, , "estimate"]
ibd_lower_rp5 <- dres_gm[, , "CI_lower"]
ibd_upper_rp5 <- dres_gm[, , "CI_upper"]

ibd_df_rp5_df  <- as.data.frame(ibd_est_rp5)
ibd_df_rp5_df  <- tibble::rownames_to_column(ibd_df_rp5_df, var = "sample.x")

ibd_df_rp5_lower_df  <- as.data.frame(ibd_lower_rp5)
ibd_df_rp5_lower_df  <- tibble::rownames_to_column(ibd_df_rp5_lower_df, var = "sample.x")

ibd_df_rp5_upper_df  <- as.data.frame(ibd_upper_rp5)
ibd_df_rp5_upper_df  <- tibble::rownames_to_column(ibd_df_rp5_upper_df, var = "sample.x")

#this is a DF with all ibd data in longform between non-matching pairs of samples
ibd_df_long_rp5 <-  ibd_df_rp5_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd"
  ) %>% filter(!is.na(ibd))
hist(ibd_df_long_rp5$ibd)
# CIs for rnull >0.125
CI_lower_df_long_r5 <- ibd_df_rp5_lower_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_lower")
CI_upper_df_long_r5 <- ibd_df_rp5_upper_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_upper")
# bind CIs into long df for rnull > 0.125
ibd_df_long_rp5 <- ibd_df_long_rp5 %>%
  left_join(CI_lower_df_long_r5) %>%
  left_join(CI_upper_df_long_r5)
pvals_r5 <- dres_gm[, , "p_value"]
pvals_r5 <- as.data.frame(pvals_r5)
pvals_r5 <- tibble::rownames_to_column(pvals_r5, var = "sample.x")


# pval 0.05
pvals_df_long_r5 <- pvals_r5 %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "p_value"
  ) %>% filter(!is.na(p_value)) %>%
  mutate(p_value = p_value/2) # divide by 2 for 1 sided test
# bind pvals into long df for rnull > 0.125
ibd_df_long_rp5 <- ibd_df_long_rp5 %>%
  left_join(pvals_df_long_r5)
# make p-value very large if r <0.125, set = 1
ibd_df_long_rp5 <- ibd_df_long_rp5 %>% mutate(p_value = ifelse(ibd <0.125, 1, p_value))
#here is where i do corrections for rnull > 0.5 dataframe
long_ibd_rp5 <- ibd_df_long_rp5 %>%
  mutate(p_BH = p.adjust(p_value, method = "BH"),
         p_BF = p.adjust(p_value, method = "bonferroni")) %>%
  mutate(sig_0.01 = ifelse(p_value < 0.01, 1, 0), sig_0.01_BH = ifelse(p_BH < 0.01, 1, 0),
         sig_0.01_BF = ifelse(p_BF < 0.01, 1, 0)) %>%
  mutate(sig_0.05 = ifelse(p_value < 0.05, 1, 0), sig_0.05_BH = ifelse(p_BH < 0.05, 1, 0),
         sig_0.05_BF = ifelse(p_BF < 0.05, 1, 0))

long_ibd_rp5 %<>% 
  filter(grepl("D0", sample.x)) %>% 
  filter(grepl("D0", sample.y))


# write.csv(long_ibd_rp5,"C:/Users/Yisihak/Desktop/tes_multiple_run_result/dcifer/long_ibd_125_Maksegnit.csv", row.names = F)



threshold <- 0.5

# Get all sample names in the dataset
all_samples <- unique(c(long_ibd_rp5$sample.x, long_ibd_rp5$sample.y))

# Identify samples that have IBD >= threshold with at least one other sample
samples_with_ibd <- long_ibd_rp5 %>%
  filter(ibd >= threshold) %>%
  filter(p_BH <= 0.05) %>% 
  select(sample.x, sample.y) %>%
  tidyr::pivot_longer(cols = everything(), values_to = "sample") %>%
  distinct(sample) %>%
  pull(sample)

# Proportion of samples involved in such pairs
proportion <- length(samples_with_ibd) / length(all_samples)

print(length(all_samples))


print(length(samples_with_ibd))

proportion



```

## Dcifer - Mizan

```{r dcifer_Mizan, echo=FALSE}

mydsmp <- dcifer::formatDat(filtered_dat_mm, "sample_id", "locus", "allele")
coi   <- coi_mm$post_coi_med
afreq_mm <- dcifer::calcAfreq(mydsmp, coi)

dsmp <- dcifer::matchAfreq(mydsmp, afreq_mm)
dres_mm <- ibdDat(dsmp, coi, afreq_mm, pval = TRUE, confint = TRUE, rnull = 0.125, alpha = 0.05, nr = 1e3)

## Output dcifer results
# readr::write_rds(dres_aa, "C:/Users/Yisihak/Desktop/tes_multiple_run_result/dcifer/dcifer_result_mm.rds")

# === Relatedness matrix ===
dmat <- dres_mm[, , "estimate"]
dmat[upper.tri(dmat)] <- t(dmat)[upper.tri(t(dmat))]

# === Adjust p-values ===
dres0 <- dres_mm[, , "p_value"]
pval_vec <- dres0[lower.tri(dres0)]                           # raw p-values
pval_adj <- p.adjust(pval_vec, method = "BH")                 # [MODIFIED]: BH-adjustment
adjusted_pmat <- matrix(NA, nrow = nrow(dres0), ncol = ncol(dres0),
                        dimnames = dimnames(dres0))           # [MODIFIED]
adjusted_pmat[lower.tri(adjusted_pmat)] <- pval_adj          # [MODIFIED]
adjusted_pmat[upper.tri(adjusted_pmat)] <- t(adjusted_pmat)[upper.tri(adjusted_pmat)]  # symmetric

# === Categorize adjusted p-values ===
dres_low <- -log10(adjusted_pmat)                            # [MODIFIED]
dres_low[dres_low > -log10(0.0001)] <- "P << 0.001"
dres_low[dres_low <= -log10(0.0001) & dres_low > -log10(0.001)] <- "P < 0.001"
dres_low[dres_low <= -log10(0.001) & dres_low > -log10(0.01)] <- "P < 0.01"
dres_low[dres_low <= -log10(0.01)] <- "Not Sig"
dres_low[upper.tri(dres_low)] <- t(dres_low)[upper.tri(dres_low)]

# === Matrix to long format ===
matrix_to_longformat <- function(mat) {
  as.data.frame(mat) |>
    rownames_to_column("sample.x") |>
    pivot_longer(-sample.x, names_to = "sample.y", values_to = "value")
}

to_plot1 <- matrix_to_longformat(dmat)

# === Clustering for ordering ===
hc <- hclust(as.dist(1 - dmat))
order <- hc$order
clustered_levels <- colnames(dmat)[order]

to_plot1 <- to_plot1 |>
  mutate(sample.x = fct_relevel(sample.x, clustered_levels),
         sample.y = fct_relevel(sample.y, clustered_levels))

to_plot2 <- matrix_to_longformat(dres_low)
to_plot2 <- to_plot2 |>
  mutate(sample.x = fct_relevel(sample.x, clustered_levels),
         sample.y = fct_relevel(sample.y, clustered_levels))

# === Define colors ===
colours_upper <- colorRampPalette(c("#FFFFFF", "#1520A6"))(100)
colours_lower <- c("P << 0.001" = "#C70039", "P < 0.001" = "#FF5733", "P < 0.01" = "#FFC300", "Not Sig" = "white")

# === Filter for D0 samples ===
to_plot1_0 <- to_plot1[grepl("D0", to_plot1$sample.x) & grepl("D0", to_plot1$sample.y), ]
to_plot2_0 <- to_plot2[grepl("D0", to_plot2$sample.x) & grepl("D0", to_plot2$sample.y), ]

sample_d0 <- unique(to_plot1_0$sample.x)
cluster_0 <- clustered_levels[clustered_levels %in% sample_d0]

to_plot1_0 <- to_plot1_0 |>
  mutate(sample.x = fct_relevel(sample.x, cluster_0),
         sample.y = fct_relevel(sample.y, cluster_0))

to_plot2_0 <- to_plot2_0 |>
  mutate(sample.x = fct_relevel(sample.x, cluster_0),
         sample.y = fct_relevel(sample.y, cluster_0))

# pdf("C:/Users/Yisihak/Pictures/Djbouti figures/Djibouti_figure_relatedness_pvalue_corrected.pdf", width = 10, height = 10)

# === Heatmap ===
g0_mm <- ggplot(to_plot1_0, aes(x = sample.x, y = sample.y)) +
  geom_tile(data = subset(to_plot1_0, as.numeric(sample.x) > as.numeric(sample.y)), aes(fill = value)) +
  scale_fill_gradientn(colors = colours_upper, name = "Relatedness") +
  theme(axis.text.x = element_blank()) +
  theme(axis.text.y = element_blank()) +
  new_scale_fill() +
  geom_tile(data = subset(to_plot2_0, as.numeric(sample.x) < as.numeric(sample.y)),
            aes(fill = value)) +
  scale_fill_manual(values = colours_lower, name = "Significance", 
                    breaks = c("P << 0.001", "P < 0.001", "P < 0.01")) +
  scale_x_discrete(limits = cluster_0) +
  scale_y_discrete(limits = cluster_0) +
  ggtitle("Mizan (BH adjusted pvalue)")

g0_mm

# Step 1: Get all related pairs (IBD > 0)
related_pairs <- to_plot1_0 %>% filter(value > 0)

# Step 2: Get all unique samples involved in related pairs
related_samples <- unique(c(related_pairs$sample.x, related_pairs$sample.y))

# Step 3: Get all unique samples in the full dataset
all_samples <- unique(c(to_plot1_0$sample.x, to_plot1_0$sample.y))

# Step 4: Compute proportion
proportion_related <- length(related_samples) / length(all_samples)

# Show result
proportion_related







# Extract IBD estimates from rnull > 0.125 run
ibd_est_rp5 <- dres_mm[, , "estimate"]
ibd_lower_rp5 <- dres_mm[, , "CI_lower"]
ibd_upper_rp5 <- dres_mm[, , "CI_upper"]

ibd_df_rp5_df  <- as.data.frame(ibd_est_rp5)
ibd_df_rp5_df  <- tibble::rownames_to_column(ibd_df_rp5_df, var = "sample.x")

ibd_df_rp5_lower_df  <- as.data.frame(ibd_lower_rp5)
ibd_df_rp5_lower_df  <- tibble::rownames_to_column(ibd_df_rp5_lower_df, var = "sample.x")

ibd_df_rp5_upper_df  <- as.data.frame(ibd_upper_rp5)
ibd_df_rp5_upper_df  <- tibble::rownames_to_column(ibd_df_rp5_upper_df, var = "sample.x")

#this is a DF with all ibd data in longform between non-matching pairs of samples
ibd_df_long_rp5 <-  ibd_df_rp5_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd"
  ) %>% filter(!is.na(ibd))
hist(ibd_df_long_rp5$ibd)
# CIs for rnull >0.125
CI_lower_df_long_r5 <- ibd_df_rp5_lower_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_lower")
CI_upper_df_long_r5 <- ibd_df_rp5_upper_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_upper")
# bind CIs into long df for rnull > 0.125
ibd_df_long_rp5 <- ibd_df_long_rp5 %>%
  left_join(CI_lower_df_long_r5) %>%
  left_join(CI_upper_df_long_r5)
pvals_r5 <- dres_mm[, , "p_value"]
pvals_r5 <- as.data.frame(pvals_r5)
pvals_r5 <- tibble::rownames_to_column(pvals_r5, var = "sample.x")


# pval 0.05
pvals_df_long_r5 <- pvals_r5 %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "p_value"
  ) %>% filter(!is.na(p_value)) %>%
  mutate(p_value = p_value/2) # divide by 2 for 1 sided test
# bind pvals into long df for rnull > 0.125
ibd_df_long_rp5 <- ibd_df_long_rp5 %>%
  left_join(pvals_df_long_r5)
# make p-value very large if r <0.125, set = 1
ibd_df_long_rp5 <- ibd_df_long_rp5 %>% mutate(p_value = ifelse(ibd <0.125, 1, p_value))
#here is where i do corrections for rnull > 0.5 dataframe
long_ibd_rp5 <- ibd_df_long_rp5 %>%
  mutate(p_BH = p.adjust(p_value, method = "BH"),
         p_BF = p.adjust(p_value, method = "bonferroni")) %>%
  mutate(sig_0.01 = ifelse(p_value < 0.01, 1, 0), sig_0.01_BH = ifelse(p_BH < 0.01, 1, 0),
         sig_0.01_BF = ifelse(p_BF < 0.01, 1, 0)) %>%
  mutate(sig_0.05 = ifelse(p_value < 0.05, 1, 0), sig_0.05_BH = ifelse(p_BH < 0.05, 1, 0),
         sig_0.05_BF = ifelse(p_BF < 0.05, 1, 0))

long_ibd_rp5 %<>% 
  filter(grepl("D0", sample.x)) %>% 
  filter(grepl("D0", sample.y))

# write.csv(long_ibd_rp5,"C:/Users/Yisihak/Desktop/tes_multiple_run_result/dcifer/long_ibd_125_Mizan.csv", row.names = F)


threshold <- 0.5

# Get all sample names in the dataset
all_samples <- unique(c(long_ibd_rp5$sample.x, long_ibd_rp5$sample.y))

# Identify samples that have IBD >= threshold with at least one other sample
samples_with_ibd <- long_ibd_rp5 %>%
  filter(ibd >= threshold) %>%
  filter(p_BH <= 0.05) %>% 
  select(sample.x, sample.y) %>%
  tidyr::pivot_longer(cols = everything(), values_to = "sample") %>%
  distinct(sample) %>%
  pull(sample)

# Proportion of samples involved in such pairs
proportion <- length(samples_with_ibd) / length(all_samples)

print(length(all_samples))


print(length(samples_with_ibd))


proportion






```

## Heatmap combined
Note: this is NOT all the samples in one dataset but separate panels

```{r heatmap_each, echo=FALSE}

# Plot/save all the sites in separate panels
g0_all <- ggarrange(g0_aa, g0_ac, g0_ba, g0_gm, g0_mm, 
                    common.legend = TRUE, ncol = 3, nrow = 2, legend = "bottom")
g0_all
ggsave("C:/Users/Yisihak/Desktop/tes_multiple_run_result/dcifer/relatedness_by_site_BH_Adjusted_updated.pdf", plot = g0_all, height = 7, width = 10)

```

## Run Dcifer - all
Now with the averaged allele frequencies, we can run dcifer on all samples together

```{r dcifer_all, echo=FALSE}

# format data
mydsmp <- dcifer::formatDat(filtered_dat, "sample_id", "locus", "allele")
# Combined COI, sorted correctly
coi <- coi_summary$post_coi_med

afreq <- dcifer::calcAfreq(mydsmp, coi)

dsmp <- dcifer::matchAfreq(mydsmp, afreq)

# Run Dcifer on full dataset, with averaged AF
dres <- ibdDat(dsmp, coi, afreq, pval = TRUE, confint = TRUE, rnull = 0, alpha = 0.01, nr = 1e3)
## Output dcifer results
# readr::write_rds(dres, "C:/Users/Yisihak/Desktop/tes_multiple_run_result/dcifer/dcifer_result_all.rds")

# Do this also for rnull >0.125
# Run Dcifer on full dataset, with averaged AF
dres_rp <- ibdDat(dsmp, coi, afreq, pval = TRUE, confint = TRUE, rnull = 0.125, alpha = 0.01, nr = 1e3)
## Output dcifer results
# readr::write_rds(dres_rp, "C:/Users/Yisihak/Desktop/tes_multiple_run_result/dcifer/dcifer_result_125_all.rds")


```

### P-value corrections
Dcifer p-value with Benjamini-Hochberg of Bonferroni corrections, code from Jessica (UG_IRS)

```{r p correction, exercise=TRUE, echo=F}

ibd_est <- dres[, , "estimate"]
ibd_lower <- dres[, , "CI_lower"]
ibd_upper <- dres[, , "CI_upper"]

# do stupid stuff to make dataframe (there's probably a better way to to this)
# rnull = 0
ibd_df <- as.data.frame(ibd_est)
ibd_df <- tibble::rownames_to_column(ibd_df, var = "sample.x")

ibd_lower_df <- as.data.frame(ibd_lower)
ibd_lower_df <- tibble::rownames_to_column(ibd_lower_df, var = "sample.x")

ibd_upper_df <- as.data.frame(ibd_upper)
ibd_upper_df <- tibble::rownames_to_column(ibd_upper_df, var = "sample.x")

#this is a DF with all ibd data in longform between non-matching pairs of samples
ibd_df_long <- ibd_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd"
  ) %>% filter(!is.na(ibd)) 

hist(ibd_df_long$ibd)

# BIND IN CI VALUES TO EACH DF
CI_lower_df_long <- ibd_lower_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_lower")
CI_upper_df_long <- ibd_upper_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_upper")
# bind CIs into long df 
ibd_df_long <- ibd_df_long %>% 
  left_join(CI_lower_df_long) %>% 
  left_join(CI_upper_df_long)

#### BIND IN P VALUES TO EACH DF ###############
pvals <- dres[, , "p_value"]
pvals <- as.data.frame(pvals)
pvals <- tibble::rownames_to_column(pvals, var = "sample.x")

pvals_df_long <- pvals %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "p_value"
  ) %>% filter(!is.na(p_value)) 


# bind pvals into long df 
ibd_df_long <- ibd_df_long %>% 
  left_join(pvals_df_long)

# here is where i do corrections for rnull = 0 dataframe
long_ibd <- ibd_df_long %>% 
  mutate(p_BH = p.adjust(p_value, method = "BH"), 
         p_BF = p.adjust(p_value, method = "bonferroni")) %>% 
  mutate(sig_0.01 = ifelse(p_value < 0.01, 1, 0), sig_0.01_BH = ifelse(p_BH < 0.01, 1, 0), 
         sig_0.01_BF = ifelse(p_BF < 0.01, 1, 0)) %>% 
  mutate(sig_0.05 = ifelse(p_value < 0.05, 1, 0), sig_0.05_BH = ifelse(p_BH < 0.05, 1, 0),
         sig_0.05_BF= ifelse(p_BF < 0.05, 1, 0)) %>%
  mutate(same_site = ifelse(substring(sample.x, 1, 2) == substring(sample.y, 1, 2), 
                            "Same site", 
                            "Different site"))

# write_csv(long_ibd, "C:/Users/Yisihak/Desktop/tes_multiple_run_result/dcifer/long_ibd_updated.csv")


#### NOW DO THE SAME FOR THE RNULL > 0.125 DATA ###############

# Extract IBD estimates from rnull > 0.125 run
ibd_est_rp5 <- dres_rp[, , "estimate"]
ibd_lower_rp5 <- dres_rp[, , "CI_lower"]
ibd_upper_rp5 <- dres_rp[, , "CI_upper"]

ibd_df_rp5_df  <- as.data.frame(ibd_est_rp5)
ibd_df_rp5_df  <- tibble::rownames_to_column(ibd_df_rp5_df, var = "sample.x")

ibd_df_rp5_lower_df  <- as.data.frame(ibd_lower_rp5)
ibd_df_rp5_lower_df  <- tibble::rownames_to_column(ibd_df_rp5_lower_df, var = "sample.x")

ibd_df_rp5_upper_df  <- as.data.frame(ibd_upper_rp5)
ibd_df_rp5_upper_df  <- tibble::rownames_to_column(ibd_df_rp5_upper_df, var = "sample.x")

#this is a DF with all ibd data in longform between non-matching pairs of samples
ibd_df_long_rp5 <-  ibd_df_rp5_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd"
  ) %>% filter(!is.na(ibd))
hist(ibd_df_long_rp5$ibd)
# CIs for rnull >0.125
CI_lower_df_long_r5 <- ibd_df_rp5_lower_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_lower")
CI_upper_df_long_r5 <- ibd_df_rp5_upper_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_upper")
# bind CIs into long df for rnull > 0.125
ibd_df_long_rp5 <- ibd_df_long_rp5 %>%
  left_join(CI_lower_df_long_r5) %>%
  left_join(CI_upper_df_long_r5)
pvals_r5 <- dres_rp[, , "p_value"]
pvals_r5 <- as.data.frame(pvals_r5)
pvals_r5 <- tibble::rownames_to_column(pvals_r5, var = "sample.x")
# pval 0.05
pvals_df_long_r5 <- pvals_r5 %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "p_value"
  ) %>% filter(!is.na(p_value)) %>%
  mutate(p_value = p_value/2) # divide by 2 for 1 sided test
# bind pvals into long df for rnull > 0.125
ibd_df_long_rp5 <- ibd_df_long_rp5 %>%
  left_join(pvals_df_long_r5)
# make p-value very large if r <0.125, set = 1
ibd_df_long_rp5 <- ibd_df_long_rp5 %>% mutate(p_value = ifelse(ibd <0.125, 1, p_value))
#here is where i do corrections for rnull > 0.5 dataframe
long_ibd_rp5 <- ibd_df_long_rp5 %>%
  mutate(p_BH = p.adjust(p_value, method = "BH"),
         p_BF = p.adjust(p_value, method = "bonferroni")) %>%
  mutate(sig_0.01 = ifelse(p_value < 0.01, 1, 0), sig_0.01_BH = ifelse(p_BH < 0.01, 1, 0),
         sig_0.01_BF = ifelse(p_BF < 0.01, 1, 0)) %>%
  mutate(sig_0.05 = ifelse(p_value < 0.05, 1, 0), sig_0.05_BH = ifelse(p_BH < 0.05, 1, 0),
         sig_0.05_BF = ifelse(p_BF < 0.05, 1, 0)) %>%
  mutate(same_site = ifelse(substring(sample.x, 1, 2) == substring(sample.y, 1, 2), 
                            "Same site", 
                            "Different site"))

# write_csv(long_ibd_rp5,"C:/Users/Yisihak/Desktop/tes_multiple_run_result/dcifer/long_ibd_125_updated.csv")

```

### Look at different ways of defining highly related pairs 
##### Standard DCIFER run, adjusting for multiple comparisons (bonferroni and BH)
##### Testing null that r>0.125, no adjustment for multiple comparisons
##### Testing null that r>0.125, adjustment for multiple comparisons (bonferroni and BH)
##### Sig levels 0.05 and 0.01 examined for all methods 

```{r compare_methods, echo= FALSE, warning = FALSE, message = FALSE, fig.show = 'hide'}

################# BONFERRONI CORRECTION,  R NULL = 0 ##########

# for rnull = 0 ibd df (standard), using FWER (bonferroni)
check_sig_0.05_BF <- long_ibd %>% filter(sig_0.05_BF == 1)
check_sig_0.01_BF <- long_ibd %>% filter(sig_0.01_BF == 1)


hist(check_sig_0.05_BF$ibd, main = "Histogram of r values after BF correction, p<0.05") #40
print("Number of pairs with significant relatedness after correction after Bonferroni correction (p<0.05)")
table(check_sig_0.05_BF$sig_0.05_BF)

hist(check_sig_0.01_BF$ibd, main  = "Histogram of r values after BF correction, p<0.01")
print("Number of pairs with significant relatedness after correction after Bonferroni correction, (p<0.01)")
table(check_sig_0.01_BF$sig_0.01_BF)


################# BENJAMINI-HOCHBERG CORRECTION, RNULL = 0  ##########

#for rnull = 0 ibd df (standard), using FDR
check_sig_0.05 <- long_ibd %>% filter(sig_0.05_BH == 1)
check_sig_0.01 <- long_ibd %>% filter(sig_0.01_BH == 1)

hist(check_sig_0.05$ibd, main = "Histogram of r values after BH correction, p<0.05")
print("Number of pairs with significant relatedness after BH correction (p<0.05)")
table(check_sig_0.05$sig_0.05_BH)

hist(check_sig_0.01$ibd, main  = "Histogram of r values after BH correction, p<0.01")
print("Number of pairs with significant relatedness after correction after BH correction (p<0.01)")
table(check_sig_0.01$sig_0.01_BH)

#look at significance level BF 
check_BF <- long_ibd %>% 
 filter(p_value <= 0.05) %>% 
  select(p_value, p_BF) %>% 
  filter(p_BF <= 0.05) %>% summarize(p_value = max(p_value))
#bonferroni threshold for rnull = 0 df
p_val_BF <- check_BF$p_value

#look at significance level 
check_BH <- long_ibd %>% filter(p_value <= 0.05) %>% 
  select(p_value, p_BH) %>% 
  filter(p_BH <= 0.05) %>% 
  summarize(p_value = max(p_value))
# bonferroni threshold
p_val_BH <- check_BH$p_value


# Plot between site relatedness
long_ibd_between <- long_ibd %>%
  filter(same_site == "Different site")
g_between <- ggplot(data = long_ibd_between, aes(x = log10(p_value), y = ibd)) + 
  geom_point(color = "darkblue") + 
  geom_vline(aes(xintercept = log10(0.05), linetype = "0.05")) + 
  geom_vline(aes(xintercept = log10(p_val_BF), linetype = "Bonferroni")) + 
  geom_vline(aes(xintercept = log10(p_val_BH), linetype = "Benjamini-Hochberg")) + 
  scale_linetype_manual(name = "Adjustment Method",
                        values = c("0.05" = "solid", 
                                   "Bonferroni" = "dashed", 
                                   "Benjamini-Hochberg" = "dotdash"),
                        labels = c("0.05", "Benjamini-Hochberg", "Bonferroni")) +
  ggtitle("IBD between populations") +
  theme(legend.position = "none")
g_between

# Plot between site relatedness
long_ibd_within <- long_ibd %>%
  filter(same_site == "Same site") %>%
  mutate(site = case_when(
    grepl(pattern = "AA", sample.x) ~ "Asayita",
    grepl(pattern = "AC", sample.x) ~ "Gambella",
    grepl(pattern = "GM", sample.x) ~ "Gondar",
    grepl(pattern = "BA", sample.x) ~ "Assossa",
    grepl(pattern = "ARC", sample.x) ~ "Djibouti",
    grepl(pattern = "MM", sample.x) ~ "Mizan"))
g_within <- ggplot(data = long_ibd_within, aes(x = log10(p_value), y = ibd, color = site)) +
  geom_point() + 
  geom_vline(aes(xintercept = log10(0.05), linetype = "0.05")) + 
  geom_vline(aes(xintercept = log10(p_val_BF), linetype = "Bonferroni")) + 
  geom_vline(aes(xintercept = log10(p_val_BH), linetype = "Benjamini-Hochberg")) + 
  scale_linetype_manual(name = "Adjustment Method",
                        values = c("0.05" = "solid", 
                                   "Bonferroni" = "dashed", 
                                   "Benjamini-Hochberg" = "dotdash"),
                        labels = c("0.05", "Benjamini-Hochberg", "Bonferroni")) +
  ggtitle("IBD within populations")
g_within



################# BONFERRONI CORRECTION,  R NULL > 0.125 ##########

# for rnull = 0 ibd df (standard), using FWER (bonferroni)
check_sig_0.05_BF <- long_ibd_rp5 %>% filter(sig_0.05_BF == 1)
check_sig_0.01_BF <- long_ibd_rp5 %>% filter(sig_0.01_BF == 1)


hist(check_sig_0.05_BF$ibd, main = "Histogram of r values after BF correction, p<0.05") #40
print("Number of pairs with significant relatedness after correction after Bonferroni correction (p<0.05)")
table(check_sig_0.05_BF$sig_0.05_BF)

hist(check_sig_0.01_BF$ibd, main  = "Histogram of r values after BF correction, p<0.01")
print("Number of pairs with significant relatedness after correction after Bonferroni correction, (p<0.01)")
table(check_sig_0.01_BF$sig_0.01_BF)



########## #################  NOW FOR RNULL > 0.125 ################# #################

################# BENJAMINI-HOCHBERG CORRECTION ##########

#for rnull >0.125 ibd df (standard), using FDR
check_sig_0.05 <- long_ibd_rp5 %>% filter(sig_0.05_BH == 1)
check_sig_0.01 <- long_ibd_rp5 %>% filter(sig_0.01_BH == 1)

hist(check_sig_0.05$ibd, main = "Histogram of r values after BH correction, p<0.05")
print("Number of pairs with significant relatedness after BH correction (p<0.05)")
table(check_sig_0.05$sig_0.05_BH)

hist(check_sig_0.01$ibd, main  = "Histogram of r values after BH correction, p<0.01")
print("Number of pairs with significant relatedness after correction after BH correction (p<0.01)")
table(check_sig_0.01$sig_0.01_BH)

#look at significance level BF
check_BF <- long_ibd_rp5 %>%
 filter(p_value <= 0.05) %>%
  select(p_value, p_BF) %>%
  filter(p_BF <= 0.05) %>% summarize(p_value = max(p_value))
#bonferroni threshold
p_val_BF_0.125 <- check_BF$p_value

#look at significance level
check_BH <- long_ibd_rp5 %>% filter(p_value <= 0.05) %>%
  select(p_value, p_BH) %>%
  filter(p_BH <= 0.05) %>%
  summarize(p_value = max(p_value))

# bonferroni threshold
p_val_BH_0.125 <- check_BH$p_value


# Plot between pop. relatedness
long_ibd_rp5_between <- long_ibd_rp5 %>%
  filter(same_site == "Different site")
g.125_between <- ggplot(data = long_ibd_rp5_between, aes(x = log10(p_value), y = ibd)) +
  geom_point(color = "darkblue") +
  geom_vline(aes(xintercept = log10(0.05), linetype = "0.05")) +
  geom_vline(aes(xintercept = log10(p_val_BF), linetype = "Bonferroni")) +
  geom_vline(aes(xintercept = log10(p_val_BH), linetype = "Benjamini-Hochberg")) +
  scale_linetype_manual(name = "Adjustment Method",
                        values = c("0.05" = "solid",
                                   "Bonferroni" = "dashed",
                                   "Benjamini-Hochberg" = "dotdash"),
                        labels = c("0.05", "Benjamini-Hochberg", "Bonferroni")) +
  ggtitle("IBD between populations (rnull > 0.125)") +
  theme(legend.position = "none")
g.125_between

# Plot between pop. relatedness
long_ibd_rp5_within <- long_ibd_rp5 %>%
  filter(same_site == "Same site") %>%
  mutate(site = case_when(
    grepl(pattern = "AA", sample.x) ~ "Asayita",
    grepl(pattern = "AC", sample.x) ~ "Gambella",
    grepl(pattern = "GM", sample.x) ~ "Gondar",
    grepl(pattern = "BA", sample.x) ~ "Assossa",
    grepl(pattern = "ARC", sample.x) ~ "Djibouti",
    grepl(pattern = "MM", sample.x) ~ "Mizan"))
g.125_within <- ggplot(data = long_ibd_rp5_within, aes(x = log10(p_value), y = ibd, color = site)) +
  geom_point() + 
  geom_vline(aes(xintercept = log10(0.05), linetype = "0.05")) + 
  geom_vline(aes(xintercept = log10(p_val_BF), linetype = "Bonferroni")) + 
  geom_vline(aes(xintercept = log10(p_val_BH), linetype = "Benjamini-Hochberg")) + 
  scale_linetype_manual(name = "Adjustment Method",
                        values = c("0.05" = "solid", 
                                   "Bonferroni" = "dashed", 
                                   "Benjamini-Hochberg" = "dotdash"),
                        labels = c("0.05", "Benjamini-Hochberg", "Bonferroni")) +
  ggtitle("IBD within populations (rnull > 0.125)")
g.125_within


g_p.thres <- ggarrange(g_between, g_within, g.125_between, g.125_within,
                       ncol = 2, nrow = 2)
g_p.thres
# ggsave(plot = g_p.thres, "hamms_dcifer_NEW/using_coi/dcifer_threshold_comparisons.pdf", height = 15, width = 15)

```

