---
title: "HAMMS - pre-moire allele check"
author: "Created by: WL"
date: "`r Sys.Date()`"
output: html_document
---

### Load libraries and working directory

```{r load libraries, echo=FALSE, warning = FALSE, message= FALSE}


library(magrittr)
library(ggplot2)
library(reshape2)
library(tidyr)
library(tidyverse)
library(stringr)
library(tools)
library(purrr)
library(parallel)
library(data.table)
library(vctrs)
library(plotly)
# library(trelliscopejs)
library(shiny)
library(htmltools)
library(htmlwidgets)
library(viridis)
library(gridExtra)
library(ggrepel)
library(readxl)
library(scales)
library(ggbeeswarm)
library(gtsummary)
library(dplyr)
```

Import HAMMS dataset. Also, merge separate runs into single allele table.


```{r visualization-functions}
#| label: visualization-functions

plot_allele_distribution <- function(dat) {
  n_alleles_per_locus_by_sample <- dat |>
    dplyr::group_by(sample_id, locus) |>
    dplyr::summarize(n_alleles = dplyr::n_distinct(allele)) |>
    dplyr::ungroup()

  ggplot(n_alleles_per_locus_by_sample, aes(x = sample_id, group = n_alleles, fill = as.factor(n_alleles))) +
    geom_bar(position = "stack") +
    theme(axis.text.x = element_text(angle = 90)) +
    scale_x_discrete(limits = n_alleles_per_locus_by_sample |>
      dplyr::group_by(sample_id) |>
      dplyr::summarize(avg_n_alleles = mean(n_alleles)) |>
      dplyr::arrange(avg_n_alleles) |>
      dplyr::pull(sample_id)) +
    xlab("Sample") +
    ylab("Number of Alleles")
}

plot_allele_prev_metrics <- function(dat, use_plotly = TRUE) {
  total_samples <- dat |>
    dplyr::distinct(sample_id) |>
    nrow()

  allele_prev_metrics <- dat |>
    dplyr::group_by(sample_id, locus) |>
    dplyr::mutate(prop = reads / sum(reads)) |>
    dplyr::ungroup() |>
    dplyr::group_by(locus, allele) |>
    dplyr::summarize(
      within_host_avg_prop = mean(prop),
      pop_prev = dplyr::n_distinct(sample_id) / total_samples,
      n_samples = dplyr::n_distinct(sample_id)
    )

  g <- ggplot(
      allele_prev_metrics, 
      aes(x = pop_prev, y = within_host_avg_prop, color = locus)
    ) +
    geom_point() +
    guides(color = "none") +
    xlab("Population Prevalence") +
    ylab("Within-Host Average Proportion") +
    theme_minimal()
  g
}
```




#merge the raw allele data
```{r}

run4 <- read.table("C:/Users/Yisihak/Desktop/tes_multiple_run_result/inbetween days/allele_data_run4.txt", sep="\t", header = T) %>%
  # filter(SampleID != "BA056-D0", Locus != "Pf3D7_03_v3-891553-891827-1A", PseudoCIGAR != "93C116G177C") %>%
  # filter(SampleID != "MM024-D28", Locus != "Pf3D7_14_v3-1166107-1166371-1A", PseudoCIGAR != "6A28+9N64+8N88+9N100+8N179A") %>%
  select(-Allele)


run5 <- read.table("C:/Users/Yisihak/Desktop/tes_multiple_run_result/inbetween days/allele_data_run5.txt", sep="\t", header = T) %>%
  # filter(Locus != "Pf3D7_13_v3-1465616-1465890-1A") %>%
  select(-Allele)


run9 <- read.table("C:/Users/Yisihak/Desktop/tes_multiple_run_result/inbetween days/allele_data_run9.txt", sep="\t", header = T) %>%
  # filter(Locus != "Pf3D7_13_v3-1465616-1465890-1A") %>%
  select(-Allele)


data <- rbind(run4, run5, run9)

# Remove suffix not part of the sample name
data$SampleID <- sub("_S[0-9]+", "",data$SampleID)

pseudo_cigar_to_alleles <- function(df) {
  df %>% select(Locus, PseudoCIGAR, ASV) %>% 
    distinct() %>% 
    group_by(Locus) %>% 
    mutate(Allele = seq(1, n())) %>% 
    ungroup() %>% 
    mutate(Allele = paste0(Locus, ".", Allele))
}

allele_sequence <- pseudo_cigar_to_alleles(data)

data %<>% left_join(allele_sequence, by = c("PseudoCIGAR", "Locus", "ASV")) %>%
  group_by(SampleID, Locus, Allele, PseudoCIGAR) %>%
  mutate(Read = sum(Reads)) %>%
  ungroup() %>%
  select(-Reads) %>%
  rename(Reads = Read) %>%
  distinct() %>%
  filter(!grepl("3D7*", SampleID))








# data %<>% left_join(allele_sequence, by = c("PseudoCIGAR", "Locus")) %>% 
#   group_by(SampleID, Locus, Allele, PseudoCIGAR) %>% 
#   summarize(Reads = sum(Reads)) %>% 
#   ungroup() %>% 
#   filter(!grepl("3D7*", SampleID))


# Check for duplicates
any(duplicated(data)) # we good


# Overall range of the data
sample_size <- n_distinct(data$SampleID)
print(sprintf("Sample size= %s", sample_size))

Locus_no <- n_distinct(data$Locus)
print(sprintf("All Locus = %s", Locus_no))



# (plot_allele_prev_metrics(data |> dplyr::rename(sample_id = SampleID, locus = Locus, allele = Allele, reads = Reads)) +
#     labs(title = "Pre-QC Data")) |>
    # ggplotly(tooltip = c("locus", "pop_prev", "within_host_avg_prop"))



# Use only diversity markers
pools <- read_tsv("C:/Users/Yisihak/Desktop/tes_multiple_run_result/QC/mad4hatter_pool_info/Pool1A-Diversity.tsv", trim_ws = TRUE) %>%
  rename("Locus" = "locus-pool")

data <- data %>%
  left_join(pools, by = "Locus")

data_div <- data %>%
  filter(Category == "Diversity") %>% 
  mutate(site = case_when(
    grepl("AA", SampleID) ~ "Asayita",
    grepl("AC", SampleID) ~ "Gambella",
    grepl("BA", SampleID) ~ "Benshangul",
    grepl("GM", SampleID) ~ "Gondar",
    grepl("MM", SampleID) ~ "Mizan",
    TRUE ~ "NA"
  ))


print(sprintf("Sample size= %s", sample_size)) # 615 final
Locus_no <- n_distinct(data_div$Locus)
print(sprintf("All Locus = %s", Locus_no)) # 165 final

allele_data_export <- data_div %>% 
  select(SampleID, Locus,ASV, PseudoCIGAR,Allele, Reads,site)

write_csv(allele_data_export,"C:/Users/Yisihak/Desktop/tes_multiple_run_result/inbetween days/TES_multisite_allele_data_pre-qc.csv")



```
#merge the filtered allele data and remove minor alleles
#remove minor allele(preparing an input to moire)

```{r import, echo = F, warning = F, message= F}


run4 <- read_csv("C:/Users/Yisihak/Desktop/tes_multiple_run_result/inbetween days/filtered_allele_data_run4.csv") %>%
  filter(SampleID != "BA056-D0", Locus != "Pf3D7_03_v3-891553-891827-1A", PseudoCIGAR != "93C116G177C") %>%
  filter(SampleID != "MM024-D28", Locus != "Pf3D7_14_v3-1166107-1166371-1A", PseudoCIGAR != "6A28+9N64+8N88+9N100+8N179A") %>%
  select(-norm.reads.locus, -norm.Reads.Locus, -n.alleles, -Allele)


run5 <- read_csv("C:/Users/Yisihak/Desktop/tes_multiple_run_result/inbetween days/filtered_allele_data_run5.csv") %>%
  filter(Locus != "Pf3D7_13_v3-1465616-1465890-1A") %>%
  select(-norm.reads.locus, -norm.Reads.Locus, -n.alleles, -Allele)


run9 <- read_csv("C:/Users/Yisihak/Desktop/tes_multiple_run_result/inbetween days/filtered_allele_data_run9.csv") %>%
  filter(Locus != "Pf3D7_13_v3-1465616-1465890-1A") %>%
  select(-norm.reads.locus, -norm.Reads.Locus, -n.alleles, -Allele)


data <- rbind(run4, run5, run9)
# data <- rbind(run4)

# Remove suffix not part of the sample name
data$SampleID <- sub("_S[0-9]+", "",data$SampleID)

pseudo_cigar_to_alleles <- function(df) {
  df %>% select(Locus, PseudoCIGAR, ASV) %>% 
    distinct() %>% 
    group_by(Locus) %>% 
    mutate(Allele = seq(1, n())) %>% 
    ungroup() %>% 
    mutate(Allele = paste0(Locus, ".", Allele))
}


allele_sequence <- pseudo_cigar_to_alleles(data)


data %<>% left_join(allele_sequence, by = c("PseudoCIGAR", "Locus", "ASV")) %>%
  group_by(SampleID, Locus, Allele) %>%
  mutate(Read = sum(Reads)) %>%
  filter(Read > 5) %>%  # filter alleles with less than 5 reads
  ungroup() %>%
  select(-Reads) %>%
  dplyr::rename(Reads = Read) %>%
  distinct() %>%
  filter(!grepl("3D7*", SampleID))


# data %<>% left_join(allele_sequence, by = c("PseudoCIGAR", "Locus")) %>% 
#   group_by(SampleID, Locus, Allele) %>% 
#   summarize(Reads = sum(Reads)) %>% 
#   filter(Reads > 5) %>%  # filter alleles with less than 5 reads
#   ungroup() %>% 
#   filter(!grepl("3D7*", SampleID))


# Check for duplicates
any(duplicated(data)) # we good


# Overall range of the data
sample_size <- n_distinct(data$SampleID)
print(sprintf("Sample size= %s", sample_size))

Locus_no <- n_distinct(data$Locus)
print(sprintf("All Locus = %s", Locus_no))


(plot_allele_prev_metrics(data |> dplyr::rename(sample_id = SampleID, locus = Locus, allele = Allele, reads = Reads)) +
    labs(title = "Pre-QC Data"))


# Use only diversity markers
pools <- read_tsv("C:/Users/Yisihak/Desktop/tes_multiple_run_result/QC/mad4hatter_pool_info/Pool1A-Diversity.tsv", trim_ws = TRUE) %>%
  dplyr::rename("Locus" = "locus-pool")

data <- data %>%
  left_join(pools, by = "Locus")

data_div <- data %>%
  filter(Category.x == "Diversity") %>% 
  mutate(site = case_when(
    grepl("AA", SampleID) ~ "Asayita",
    grepl("AC", SampleID) ~ "Gambella",
    grepl("BA", SampleID) ~ "Benshangul",
    grepl("GM", SampleID) ~ "Gondar",
    grepl("MM", SampleID) ~ "Mizan",
    TRUE ~ "NA"
  ))


print(sprintf("Sample size= %s", sample_size)) # 615 final
Locus_no <- n_distinct(data_div$Locus)
print(sprintf("All Locus = %s", Locus_no)) # 165 final

allele_data_export <- data_div %>% 
  select(SampleID, Locus,ASV,PseudoCIGAR, Allele, Reads,site)


loci_to_remove <- read_csv("C:/Users/Yisihak/Desktop/tes_multiple_run_result/loci_to_remove.csv")


data_div %<>%
  filter(!Locus %in% loci_to_remove$Locus)


(plot_allele_prev_metrics(data_div |> dplyr::rename(sample_id = SampleID, locus = Locus, allele = Allele, reads = Reads)) +
    labs(title = "Pre-QC Data"))


# write_csv(allele_data_export,"C:/Users/Yisihak/Desktop/tes_multiple_run_result/inbetween days/TESP_multisite_allele_data_filtered.csv")



```

### Preview raw data

What is the allele distribution for polyclonal loci?

```{r preview, echo=FALSE, warning = FALSE, message= FALSE}

data_new <- data_div %>% 
  dplyr::rename(sample_id = SampleID)

# Count number of alleles per Locus per sample
allele_count <- data_new %>% 
    group_by(sample_id, Locus) %>%
  summarise(n.alleles = n()) %>%
  group_by(Locus) %>% 
  arrange(-n.alleles)
# Allele range in each sample
print(sprintf("%s to %s alleles per Locus per sample",
              min(allele_count$n.alleles), max(allele_count$n.alleles)))

n_allele <- data_new %>%
  distinct(sample_id, Locus, Allele, .keep_all = TRUE) %>%
  group_by(sample_id, Locus) %>%
  summarise(n.alleles = n(), norm.reads.locus = Reads/sum(Reads))

# allele stats
allele_stats <- n_allele %>%
  group_by(Locus) %>%
  summarise(mean = mean(n.alleles), median = median(n.alleles),
            min = min(n.alleles), max = max(n.alleles)) %>%
  arrange(-max)
print(head(allele_stats))

# Define categories for number of alleles
n_allele_poly <- n_allele %>%
  filter(n.alleles >1) %>%
  mutate(allele_category = case_when(
    n.alleles == 2 ~ "2",
    n.alleles == 3 ~ "3",
    n.alleles %in% 4:5 ~ "4-5",
    n.alleles %in% 6:8 ~ "6-8",
    n.alleles > 8 ~ "9+"
  ))
# Calculate proportion of samples with each allele category per locus
n_allele_plot <- n_allele_poly %>%
  group_by(Locus, allele_category) %>%
  summarise(count = n()) %>%
  group_by(Locus) %>%
  mutate(prop.allele = count / sum(count))

# Reorder the allele_category factor levels
n_allele_plot <- n_allele_plot %>%
  mutate(allele_category = factor(allele_category, levels = c("2", "3", "4-5", "6-8", "9+")))

# Set order to proportion of 2 alleles
level_2allele <- n_allele_plot %>%
    filter(allele_category == "2") %>%
    arrange(-prop.allele) %>%
    pull(Locus)
n_allele_plot <- n_allele_plot %>%
  mutate(Locus = factor(Locus, levels = level_2allele))

# Plotting the stacked bar plot
g1 <- ggplot(n_allele_plot, aes(x = factor(Locus, level = level_2allele), y = prop.allele, fill = allele_category)) +
  geom_bar(stat = "identity") +
  labs(x = "Locus", y = "Proportion", fill = "Number of Alleles") +
  theme_classic() +
  theme(axis.text.x = element_blank()) 
g1p <- ggplotly(g1)
g1p


```

### Filter minor alleles by 1% prevalence

How does the allele distribution change after filtering?

```{r filter, echo=FALSE, warning = FALSE, message= FALSE}

##### Population level filtering first ###########

# Filter alleles with less than 1% of total reads at the population level
filtered_alleles <- data_new %>%
  group_by(sample_id, site, Locus, Allele) %>%
  summarise(
    total_reads = sum(Reads), # Total reads per allele across samples
    locus_reads = sum(Reads, na.rm = TRUE) %>% first(), # Total reads in the locus per sample/site
    .groups = "drop"
  ) %>%
  mutate(population_frequency = total_reads / locus_reads * 100) %>%
  filter(population_frequency >= 2) %>%
  select(sample_id, site, Locus, Allele)

# Join back with the original dataframe to retain only filtered alleles
filtered_data <- data_new %>%
  inner_join(filtered_alleles, by = c("sample_id", "site", "Locus", "Allele"))

###### Now do straight 1% host-level allele frequency ###########

filtered_data %<>%
  distinct(sample_id, Locus, Allele, .keep_all = TRUE) %>%
  group_by(sample_id, Locus) %>%
  mutate(norm.reads.locus = Reads/sum(Reads))

ggplot(data = filtered_data, aes(x = norm.reads.locus, y = Reads)) +
  geom_point(aes(color = site)) +
  scale_y_log10() +
  facet_wrap(~site) +
  ggtitle("Allele read count vs. within-host frequency") +
  geom_vline(xintercept = 0.01)


# Filter by norm.reads.locus > 0.01
filtered_allele_data <- filtered_data %>%
  filter(norm.reads.locus > 0.02)
  # filter(norm.reads.locus > 0.10 & Reads > 10) # apply another, if <10% AND read count > 10 (1 out of 10 reads)

# Now how does it look now
ggplot(data = filtered_allele_data, aes(x = norm.reads.locus, y = Reads)) +
  geom_point(aes(color = site)) +
  scale_y_log10() +
  facet_wrap(~site) +
  ggtitle("Allele read count vs. within-host frequency")

filtered <- filtered_allele_data %>%
  group_by(sample_id) %>%
  summarise(total = sum(Reads))

# write_csv(filtered_allele_data, "TES_final_allele_data.csv")

sample_size <- n_distinct(filtered_allele_data$sample_id)
print(sprintf("Sample size= %s", sample_size))

# new allele stats
allele_count_filt <- filtered_allele_data %>%
  group_by(sample_id, Locus) %>%
  mutate(n.alleles = n())

# Allele range in each sample
print(sprintf("%s to %s alleles per Locus per sample",
              min(allele_count_filt$n.alleles), max(allele_count_filt$n.alleles)))

# new allele stats
allele_stats_filt <- filtered_allele_data %>%
  group_by(sample_id, Locus) %>%
  mutate(n.alleles = n()) %>%
  group_by(Locus) %>%
  summarize(mean = mean(n.alleles), median = median(n.alleles),
            min = min(n.alleles), max = max(n.alleles)) %>%
  arrange(-max)
print(head(allele_stats_filt))

# Define categories for number of alleles
allele_filt_poly <- allele_count_filt %>%
  filter(n.alleles >1) %>%
  mutate(allele_category = case_when(
    n.alleles == 2 ~ "2",
    n.alleles == 3 ~ "3",
    n.alleles %in% 4:5 ~ "4-5",
    n.alleles %in% 6:8 ~ "6-8",
    n.alleles > 8 ~ "9+")) %>%
  group_by(Locus, allele_category) %>%
  summarize(count = n()) %>%
  group_by(Locus) %>%
  mutate(prop.allele = count / sum(count))
# Reorder the allele_category factor levels
allele_filt_poly <- allele_filt_poly %>%
  mutate(allele_category = factor(allele_category, levels = c("2", "3", "4-5", "6-8", "9+")))
# Set order to proportion of 2 alleles
level_2allele <- allele_filt_poly %>%
    filter(allele_category == "2") %>%
    arrange(-prop.allele) %>%
    pull(Locus)
allele_filt_poly <- allele_filt_poly %>%
  mutate(Locus = factor(Locus, levels = level_2allele))
# Plotting the stacked bar plot
g2 <- ggplot(allele_filt_poly, aes(x = factor(Locus, level = level_2allele), y = prop.allele, fill = allele_category)) +
  geom_bar(stat = "identity") +
  labs(x = "Locus", y = "Proportion", fill = "Number of Alleles") +
  theme_classic() +
  theme(axis.text.x = element_blank()) 
g2p <- ggplotly(g2)
g2p


```

Looks like no loci have 9+ alleles after filtering. The max number of alleles is now 5. What are the loci with 4-5 alleles?

```{r investigate, echo=FALSE, warning = FALSE, message= FALSE}

loci_lot.al <- allele_filt_poly %>%
  filter(allele_category == "4-5") %>%
  pull(Locus)
print(loci_lot.al)

```

### Make moire input

```{r moire_input, echo=FALSE, warning = FALSE, message= FALSE}

# filtered_data <- filtered_allele_data %>% 
#   select(sample_id,Locus,ASV,PseudoCIGAR,Allele,site)
# 
# write_csv(filtered_data,"C:/Users/Yisihak/Desktop/tes_multiple_run_result/QC/TES_final_allele_data_filtered_updated.csv")

# loci_to_remove <- read_csv("C:/Users/Yisihak/Desktop/tes_multiple_run_result/loci_to_remove.csv")

# Make moire input
moire_input <- filtered_allele_data %>% 
  select (sample_id, Locus, Allele, site) %>%
  rename("locus" = "Locus") %>%
  rename("allele" = "Allele") %>% 
  distinct()

# moire_input_updated <- moire_input %>%
#   filter(!locus %in% loci_to_remove$Locus)


moire_input_aa <- moire_input %>% 
  filter(site == "Asayita")

moire_input_ac <- moire_input %>% 
  filter(site == "Gambella")

moire_input_ba <- moire_input %>% 
  filter(site == "Benshangul")

moire_input_gm <- moire_input %>% 
  filter(site == "Gondar")

moire_input_mm <- moire_input %>% 
  filter(site == "Mizan")

write.csv(moire_input, "C:/Users/Yisihak/Desktop/tes_multiple_run_result/moire/TES_moire_input_filtered__updated.csv", row.names = F)
write.csv(moire_input_aa, "C:/Users/Yisihak/Desktop/tes_multiple_run_result/moire/TES_moire_input_filtered_updated_aa.csv", row.names = F)
write.csv(moire_input_ac, "C:/Users/Yisihak/Desktop/tes_multiple_run_result/moire/TES_moire_input_filtered_updated_ac.csv", row.names = F)
write.csv(moire_input_ba, "C:/Users/Yisihak/Desktop/tes_multiple_run_result/moire/TES_moire_input_filtered_updated_ba.csv", row.names = F)
write.csv(moire_input_gm, "C:/Users/Yisihak/Desktop/tes_multiple_run_result/moire/TES_moire_input_filtered_updated_gm.csv", row.names = F)
write.csv(moire_input_mm, "C:/Users/Yisihak/Desktop/tes_multiple_run_result/moire/TES_moire_input_filtered_updated_mm.csv", row.names = F)




```


#run moire


```{r}

library(moire)
library(readr)
library(ggplot2)

# load your filtered data
data <- read_csv("C:/Users/Yisihak/Desktop/tes_multiple_run_result/moire/TES_moire_input_filtered_aa.csv")
moire_data <- moire::load_long_form_data(data)

#### MCMC parameters
# this is likely a sufficient amount of burnin and sampling, however it can
# be dataset specific. You will want to assess convergence by examining the 
# likelihood traces and see if there are any trends suggesting that the
# loglikelihood is continuing to improve during the sampling stage. A converged
# loglikelihood trace will be flat, looking like a "hairy caterpillar".
burnin <- 1e3
num_samples <- 1e3

# this is the thinning rate, meaning we will only keep every nth sample
# from the MCMC chain. This is useful for reducing the final size of the 
# MCMC result, however there generally isn't a reason to increase this
# because we usually have more than enough memory to store the results.
thin <- 1

# This is the total number of chains that will be run using the parallel tempering
# algorithm. Generally, the more chains you run, the more likely the MCMC will
# converge. However, the more chains you run, the more computationally expensive
# the analysis will be.
total_chains <- 80

# This is the lower temperature of the parallel tempering algorithm. If you run the
# MCMC algorithm and find that the chain swap plot (moire::plot_chain_swaps) shows
# that the chains are not swapping well (points fall below .25), you may want to 
# increase the lower temperature value. This may impact the ability of the algorithm
# to converge, but it will help the chains swap more effectively.
lower_pt_temp <- .5

# This is the total number of cores that will be used to run the MCMC when using parallel
# tempering. num_cores should be less than or equal to the number of chains, otherwise
# some cores will be idle.
num_cores <- 16

#### PRIORS
## Relatedness: prior distribution is beta, meaning values are between 0 and 1.
# In general, we believe that relatedness is likely to be away from 0 or 1 under
# a biological perspective. For relatedness to be close to 1, we would need to have
# repeated co-transmission without any superinfection occurring, leading to severe
# bottlenecking of the parasite population within individuals. For relatedness to be
# close to 0, we would need to have infections within individuals to be completely
# independent of and unrelated to each other, as if they came from a complete well 
# mixed background population. The following prior is a weakly informative prior
# that is centered around 0.5 and steers away from extreme values near 0 or 1.
r_alpha <- 3
r_beta <- 3

## False Positive/Negative Rates:
# False positives occur when an allele is incorrectly called as present when it is
# actually absent. False negatives occur when an allele is incorrectly called as
# absent when it is actually present.
# In general, the error rates are likely to be low, so it would be reasonable to 
# to use a prior that favors small values. However, the model is relatively robust
# to the error rates, so it is unlikely that our priors will have a large impact
# on the results.

# This is a flat prior on rates, meaning we have no prior beliefs or information
# available to say that any particular rate is more or less likely than any other.
# eps_pos_alpha <- 1
# eps_pos_beta <- 1
# eps_neg_alpha <- 1
# eps_neg_beta <- 1

# Alternatively, we can use a prior that favors small values as follows:
eps_pos_alpha <- .1
eps_pos_beta <- 9.9
eps_neg_alpha <- .1
eps_neg_beta <- 9.9

## Population mean COI: 
# Part of the model places a prior on the distribution of COI values (specifically 
# a zero truncated poisson distribution). We do not know the mean of this distribution,
# so we place a weakly informative prior on the mean of the distribution using a 
# gamma distribution. The following prior  is weakly informative in that it places
# the mean of the distribution at slightly above 1. In general, you should not need
# to change this prior as it has little overall impact on the results.
mean_coi_shape <- 0.1
mean_coi_scale <- 10


# Run MOIRE
moire_results <- moire::run_mcmc(
    moire_data, is_missing = moire_data$is_missing,
    burnin = burnin,
    thin = thin,
    samples_per_chain = num_samples,
    eps_pos_alpha = eps_pos_alpha,
    eps_pos_beta = eps_pos_beta,
    eps_neg_alpha = eps_neg_alpha,
    eps_neg_beta = eps_neg_beta,
    r_alpha = r_alpha,
    r_beta = r_beta,
    mean_coi_shape = mean_coi_shape,
    mean_coi_scale = mean_coi_scale,
    pt_chains = seq(1, lower_pt_temp, length.out = total_chains),
    pt_num_threads = num_cores
)

# save the results to a file so you don't have to re-run the MCMC
readr::write_rds(moire_results, "C:/Users/Yisihak/Desktop/tes_multiple_run_result/moire/TES_moire_result_aa.rds")

# After running the MCMC, you should examine the swap rate plot to ensure the chains
# are swapping well. Good swapping is indicated by all points being between .25 - .75.
# If the chains are not swapping well (many or most chains are below the .25 swap rate),
# you may want to increase the lower temperature value or increase the number of chains that
# are run.
moire::plot_chain_swaps(moire_results)

# You should also examine the loglikelihood trace to ensure the chains are converging.
# A converged loglikelihood trace will be flat, looking like a "hairy caterpillar".
qplot(1:length(moire_results$chains[[1]]$llik_sample), moire_results$chains[[1]]$llik_sample, geom = "line")

```

