---
title: "QC FOR a sequence run "
author: "Created by: William, Modified by: Yisihak"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

**Load libraries and working directory**
```{r load libraries, echo=FALSE, warning = FALSE, message= FALSE}

library(dplyr)
library(magrittr)
library(ggplot2)
library(reshape2)
library(tidyr)
library(tidyverse)
library(stringr)
library(tools)
library(purrr)
library(parallel)
library(data.table)
library(vctrs)
library(plotly)
# library(trelliscopejs)
library(shiny)
library(htmltools)
library(htmlwidgets)
library(viridis)
library(gridExtra)
library(ggrepel)
library(readxl)
library(scales)

```
+

# import Allele and pool data  

```{r assign Allele by pseudo cigars, echo=FALSE, warning = FALSE, message= FALSE}

allele_data_run4 <- read.table("C:/Users/Y/Desktop/tes_multiple_run_result/run4/allele_data.txt", sep = "\t", header = T) %>%
  select(SampleID) %>%
  distinct()

allele_data_run4$SampleID <- gsub("_S.*","",allele_data_run4$SampleID)


allele_data_run5 <- read.table("C:/Users/Y/Desktop/tes_multiple_run_result/run5/allele_data.txt", sep = "\t", header = T) %>%
  select(SampleID) %>%
  distinct()

allele_data_run5$SampleID <- gsub("_S.*","",allele_data_run5$SampleID)

allele_data_run9 <- read.table("C:/Users/Y/Desktop/tes_multiple_run_result/run9/allele_data.txt", sep = "\t", header = T) %>%
  select(SampleID) %>%
  distinct()

allele_data_run9$SampleID <- gsub("_S.*","",allele_data_run9$SampleID)


allele_data_combined <- rbind(allele_data_run4,allele_data_run5,allele_data_run9) %>%
  group_by(SampleID, Locus, ASV, Allele, PseudoCIGAR) %>%
  mutate(Reads = sum(Reads)) %>%
  ungroup() %>%
  distinct()


allele_data <- allele_data_combined %>% 
  group_by(SampleID, Locus) %>% 
  mutate(reads = sum(Reads)) %>% 
  ungroup() %>% 
  group_by(Locus) %>%
  filter(all(reads > 100)) %>%     # keep rows where the locus has >100 reads in all samples
  distinct(Locus) 

### Rearrange Allele by PseudoCIGAR (not asv - variations in masked regions are not considered!)
cigar <- allele_data %>% 
  select (Locus, PseudoCIGAR) %>% 
  group_by(Locus) %>% 
  distinct(PseudoCIGAR) %>% 
  arrange(Locus) %>%
  dplyr::mutate(type = paste0("t", dplyr::cur_group_id()))
cigar$Allele <- paste(cigar$type, rowid(cigar$type), sep = ".")

### Make concatenated allele data
cigar_select <- cigar %>% select(Locus, PseudoCIGAR, Allele)
allele_data <- left_join(allele_data, cigar_select)
allele_data %<>%
  distinct(SampleID, Locus, PseudoCIGAR, Allele, .keep_all = TRUE) %>%
  group_by(SampleID, Locus) %>%
  mutate(norm.reads.locus = Reads/sum(Reads))%>%
  mutate(n.alleles = n()) %>% 
  arrange(SampleID, Locus)


# Import pool1A info
pool_1A <- read_tsv("C:/Users/Y/Desktop/mad4hatter_pool_info/Pool1A-Diversity.tsv", trim_ws = TRUE) %>%
  rename("Locus" = "locus-pool")

# Import pool5 info
pool_1B <- read_tsv("C:/Users/Y/Desktop/mad4hatter_pool_info/Pool1B-Resitance+.tsv", trim_ws = TRUE) %>%
  rename("Locus" = "locus-pool")

# Import pool2 info
pool_2 <- read_tsv("C:/Users/Y/Desktop/mad4hatter_pool_info/Pool2-Resistance+.tsv", trim_ws = TRUE) %>%
  rename("Locus" = "locus-pool")

```


**Preview raw data - sample size, loci numbers, allele numbers**
```{r preview data info, echo=FALSE, warning = FALSE, message= FALSE}

data <- allele_data
  
# Overall range of the data
sample_size <- n_distinct(data$SampleID)
print(sprintf("Sample size= %s", sample_size))

Locus_no <- n_distinct(data$Locus)
print(sprintf("All Locus = %s", Locus_no))

# Count number of alleles per Locus per sample
allele_count <- data %>% 
  group_by(Locus) %>% 
  summarize(n.alleles = n_distinct(Allele)) %>% 
  arrange(-n.alleles)
# Allele range in each sample
print(sprintf("%s to %s alleles per Locus per sample",
              min(allele_count$n.alleles), max(allele_count$n.alleles)))

```


# Data QC and filtering 
(1) Filter out bad-quality samples and possible false positive alleles (off-targets, low-abundance)

**Step 1. Total Reads**
```{r filter total reads, echo=FALSE, warning = FALSE, message= FALSE}

## Plot total Reads distribution - NC, PC, Samples
plot_Reads_dist <- function(data){

  # Summarize total reads for a sample
  Reads_table <- data %>% 
  distinct(SampleID, Locus, Allele, .keep_all = TRUE) %>%
  group_by(SampleID) %>% 
  summarize(Total_Reads = sum(Reads)) %>% 
  arrange(Total_Reads) 

  #Plot
  g <- ggplot(Reads_table, aes(x = log10(Total_Reads))) +
  geom_density(alpha = .5) +
  ggtitle("Total Reads distribution")
  g
}

plot_Reads_dist(data)


## Visualize samples with low Reads: Pool1A5 has 217 loci, at least each loci has 50 Reads, so total reads cutoff will be 217*50 (1A5+2 has 237 loci)

plot_Reads_cutoff <- function(data, loci = 237, Reads_cutoff = 50){

  # Summarize total Reads
  Reads_table <- data %>% 
  distinct(SampleID, Locus, Allele, .keep_all = TRUE) %>%
  group_by(SampleID) %>% 
  summarize(Total_Reads = sum(Reads)) %>% 
  arrange(Total_Reads) 
  
  # Set up Reads criterion
  Reads_criterion <- loci*Reads_cutoff
  sample_low_Reads <- Reads_table %>% 
    filter(Total_Reads < Reads_criterion) %>% pull(SampleID)
  
  # Make plotting df
  Reads_table %<>%
   mutate(QC = case_when(
    SampleID %in% sample_low_Reads ~ "Low-Reads samples",
    TRUE ~ "QC-passed samples")) %>%
    mutate(color = case_when(
    SampleID %in% sample_low_Reads ~ "gray",
    TRUE ~ "black")) %>%
    arrange(factor(QC, levels = c("Low-Reads samples", "QC-passed samples"))) 

  color_axis <- Reads_table$color[order(Reads_table$SampleID)]
  color_fill <- c("Low-Reads samples" = "gray", "QC-passed samples" = "black")

  level_order <- Reads_table$SampleID

  # Plot
  g <- Reads_table %>%
    ggplot(aes(x = factor(SampleID, level = level_order), y = log10(Total_Reads), fill = QC)) +
    geom_bar(stat = "identity") + 
    coord_flip() + 
    scale_fill_manual(values = color_fill) +
    theme_classic() +
    theme(axis.text.y=element_blank()) +
    geom_hline(yintercept = log10(Reads_criterion), linetype='dashed', color = 'red')+
    ylab("Log10 Total Reads") +
    xlab("SampleID") +
    ggtitle(paste0("Filter out samples with <",Reads_cutoff," * 237 reads"))
  g
}


plot_Reads_cutoff(data, loci = 237, Reads_cutoff = 50)


### Filter 2: Remove samples with few reads

remove_low_Reads_samples <- function(data, loci = 237, Reads_cutoff = 50){
  # Summarize total Reads
  Reads_table <- data %>% 
    distinct(SampleID, Locus, Allele, .keep_all = TRUE) %>%
    group_by(SampleID) %>% 
    summarize(Total_Reads = sum(Reads)) %>% 
    arrange(Total_Reads)
  
  # Set up Reads criterion
  Reads_criterion <- loci * Reads_cutoff
  sample_low_Reads <- Reads_table %>%
    filter(Total_Reads < Reads_criterion) %>% pull(SampleID)

  clean_data <- data %>% filter(! SampleID %in% unique(c(sample_low_Reads))) 
  sample_size <- n_distinct(clean_data$SampleID)
  clean_data_info <- sprintf("Step 1 - Sample size after removing samples with insufficient reads = %s, Retained prop. = %s", sample_size, ... = round(sample_size/n_distinct(data$SampleID), 2))

  return(list(clean_data = clean_data, clean_data_info = clean_data_info, removed_SampleID = sample_low_Reads))
}

# Example usage:
result <- remove_low_Reads_samples(data)
filtered_data1 <- result$clean_data  # This will give you the data after removing low-Reads samples
print(result$clean_data_info)  # This will print the information of filtered data compared to original input data
sample_low_reads <- print(result$removed_SampleID) # This will print the information of removed SampleID

##WL: export names of the filtered samples
if (length(result$removed_SampleID) > 0) {
  filtered_list1 <- data.frame(as.list(result$removed_SampleID)) %>%
  transpose() %>%
  rename("SampleID" = "V1") %>%
  mutate(StudySubject = word(SampleID, 1, sep = "_")) %>%
  select(SampleID, StudySubject) %>%
  mutate(Reason = "Few Total Reads")
} else {
  filtered_list1 <- data.frame(SampleID = character(), StudySubject = character(), Reason = character())
}

```


**Step 2. Loci coverage** 

```{r filter loci coverage, echo=FALSE, warning = FALSE, message= FALSE}

## Total loci distribution and summary by RUN

plot_loci_dist <- function(data){

  # Data    
  clone <- data %>%
    group_by(SampleID, Locus) %>% 
    summarise(total_loci = n()) %>%
    mutate(Clone = ifelse(total_loci == 1, "Mono", "Poly"))
  clone %<>% group_by(SampleID) %>% count(Clone)

  clone_wide <- clone %>% pivot_wider(names_from = Clone, values_from = n)
  clone_wide[is.na(clone_wide)] <- 0
  clone_wide %<>% mutate(Total_loci = Mono + Poly) %>% 
    mutate(Mono_prop = Mono/Total_loci) %>% 
    mutate(Poly_prop = Poly/Total_loci) %>%
    arrange(- Poly_prop)

  clone_wide_select <- clone_wide %>% select(SampleID, Poly_prop) 
  clone <- merge(clone, clone_wide_select, by = "SampleID")
  clone %<>% arrange(- Poly_prop)

  # Plot
  g <- ggplot(clone_wide, aes(x = Total_loci)) +
    geom_histogram() +
    scale_x_continuous(breaks = seq(0, 300, 25)) 
  g

}

plot_loci_dist(filtered_data1)



### Visualize samples with insufficient coverage of loci

plot_loci_cutoff <- function(data, loci = 237, perc_covered = 0.7){

  # Data    
  clone <- data %>%
    group_by(SampleID, Locus) %>% 
    summarise(total_loci = n()) %>%
    mutate(Clone = ifelse(total_loci == 1, "Mono", "Poly"))
  clone %<>% group_by(SampleID) %>% count(Clone)

  clone_wide <- clone %>% pivot_wider(names_from = Clone, values_from = n)
  clone_wide[is.na(clone_wide)] <- 0
  clone_wide %<>% mutate(Total_loci = Mono + Poly) %>% 
    mutate(Mono_prop = Mono/Total_loci) %>% 
    mutate(Poly_prop = Poly/Total_loci) %>%
    arrange(- Poly_prop)

  clone_wide_select <- clone_wide %>% select(SampleID, Poly_prop) 
  clone <- merge(clone, clone_wide_select, by = "SampleID")
  clone %<>% arrange(- Poly_prop) 

  # Set up loci criterion
  loci_criterion <- round(loci*perc_covered)

  sample_less_loci <- clone %>%
    group_by(SampleID) %>%
    summarize(total_loci = sum(n)) %>%
    filter(total_loci < loci_criterion) %>% 
    pull(SampleID)

  # Plot
  g <- clone %>%
    ggplot(aes(x = reorder(SampleID, Poly_prop), y = n)) +
    geom_col(aes(fill = Clone)) +
    theme(axis.text.y=element_blank()) +
    coord_flip()+
    scale_y_continuous(name = "No. of Locus", breaks = seq(0,200, by = 50)) +
    geom_hline(yintercept = loci_criterion, linetype='dotted', col = 'black') +
    xlab("SampleID") +
    ggtitle("Filter out samples with insufficient loci coverage") 
  g
}



plot_loci_cutoff(filtered_data1, loci = 237, perc_covered = 0.7)


### Filter out samples with insufficient loci coverage
remove_low_loci_samples <- function(data, loci = 237, perc_covered = 0.7){
  
  # Data    
  clone <- data %>%
    group_by(SampleID, Locus) %>% 
    summarise(total_loci = n()) %>%
    mutate(Clone = ifelse(total_loci == 1, "Mono", "Poly"))
  clone %<>% group_by(SampleID) %>% count(Clone)
  
  clone_wide <- clone %>% pivot_wider(names_from = Clone, values_from = n)
  clone_wide[is.na(clone_wide)] <- 0
  clone_wide %<>% mutate(Total_loci = Mono + Poly) %>% 
    mutate(Mono_prop = Mono/Total_loci) %>% 
    mutate(Poly_prop = Poly/Total_loci) %>%
    arrange(- Poly_prop)
  
  clone_wide_select <- clone_wide %>% select(SampleID, Poly_prop) 
  clone <- merge(clone, clone_wide_select, by = "SampleID")
  clone %<>% arrange(- Poly_prop) 
  
  # Set up loci criterion
  loci_criterion <- round(loci*perc_covered)
  sample_less_loci <- clone %>%
    group_by(SampleID) %>%
    summarize(total_loci = sum(n)) %>%
    filter(total_loci < loci_criterion) %>% 
    pull(SampleID)
  
  clean_data <- data %>% filter(! SampleID %in% sample_less_loci) 
  sample_size <- n_distinct(clean_data$SampleID)
  clean_data_info <- sprintf("Step 2 - Sample size after removing samples with insufficeint loci coverage = %s, Retained prop. = %s", sample_size, round(sample_size/n_distinct(data$SampleID), 2))
  
  
  return(list(clean_data = clean_data, clean_data_info = clean_data_info, removed_SampleID = sample_less_loci))
}

# Example usage:
result <- remove_low_loci_samples(filtered_data1)
filtered_data2 <- result$clean_data  # This will give you the data after removing low-loci-coverage samples
print(result$clean_data_info)  # This will print the information of filtered data compared to original input data
sample_less_loci <- print(result$removed_SampleID) # This will print the information of removed SampleID

##WL: export names of the filtered samples
if (length(result$removed_SampleID) > 0) {
  filtered_list2 <- data.frame(as.list(result$removed_SampleID)) %>%
    transpose() %>%
    rename("SampleID" = "V1") %>%
    mutate(StudySubject = word(SampleID, 1, sep = "_")) %>%
    select(SampleID, StudySubject) %>%
    mutate(Reason = "Insufficient locus coverage")
} else {
  filtered_list2 <- data.frame(SampleID = character(), StudySubject = character(), Reason = character())
}


```


**Step 3: Proportion of loci > 100 Reads in samples**
```{r filter n100 proportion, echo=FALSE, warning = FALSE, message= FALSE}

pool_1Ax <- pool_1A %>% select(Locus, Category)
pool_1Bx <- pool_1B %>% select(Locus, Category)
pool_2x <- pool_2 %>%select(Locus, Category)
pool_cat <- rbind(pool_1Ax, pool_1Bx, pool_2x)

filtered_data2 <- filtered_data2 %>%
  group_by(Locus) %>%
  mutate(Pool = sapply(strsplit(Locus,"-"),tail,1)) %>%
  ungroup() %>%
  left_join(pool_cat, by = "Locus")
  

### Calculate loci having 100 Reads (n100) and plot n100 by gene categories

plot_n100_loci_cutoff <- function(data, n100_cutoff = 0.5){ 

  # Group by sample + loci category
  Locus_cov_100 <- data %>% 
    group_by(SampleID, Locus, Category, Pool) %>% 
    summarize(Reads = sum(Reads)) %>% 
    ungroup() 

 # Calculate n100 and n50 loci counts
  Locus_cov_100 %<>%
    group_by(SampleID, Category, Pool) %>% 
    summarize(total_Reads = sum(Reads), n100 = sum(Reads >= 100)) 
  
  # Normalization - calculate % loci with n100 or n50 (Species targets are not counted)
  amp_cov_category_norm <- Locus_cov_100 %>% 
    filter(Category != "Species") %>%
    mutate(norm_100 = case_when(
      Category == "Diversity" ~ n100/165,
      Category == "Resistance" ~ n100/38,
      Category == "Immune" ~ n100/5,
      Category == "Diagnostic" ~ n100/30))
  
  # Plot
  g <- ggplot(amp_cov_category_norm, aes(x=total_Reads, y = norm_100)) +
    geom_point(aes(color = "red", alpha = 0.5)) +
    scale_x_log10() + 
    geom_hline(yintercept = n100_cutoff, linetype = "dashed", color = "black") +
    facet_wrap(~ Category) +
    ggtitle("Normalized prop. of loci with > 100 Reads in each sample") +
    ylim(0,1) +
    ylab("Proportion of loci with >100 Reads") +
    xlab("Total Reads")
  g
}  

plot_n100_loci_cutoff(filtered_data2, n100_cutoff = 0.5) 


### Plot n100 proportion & sample retained

plot_n100_sample_cutoff <- function(data, n100_cutoff = 0.5){
  
  # Group by sample + loci category
Locus_cov_100 <- data %>% 
  group_by(SampleID, Locus, Category, Pool) %>% 
  summarize(Reads = sum(Reads)) %>% 
  ungroup() 

n_table <- Locus_cov_100 %>%
  filter(Category == "Diversity") %>%
  group_by(SampleID, Locus) %>% 
  summarize(total_Reads = sum(Reads)) %>%
  mutate(n100 = ifelse(total_Reads >= 100, 1, 0))

n_count <-  n_table  %>%
  group_by(SampleID) %>%
  summarise(
    n100_0_count = sum(n100 == 0),
    n100_1_count = sum(n100 == 1),
    n100_proportion = n100_1_count / n())

# Create a data frame to store the results
results <- data.frame(X = numeric(), SampleCount = numeric())

# Loop through different threshold values
for (X in 0:100) {
  filtered_data <- n_count %>%
    filter(n100_proportion >= (X / 100))  # Convert X to a proportion

  # Add X and the count of samples exceeding the threshold to the results data frame
  results <- rbind(results, data.frame(X = X, SampleCount = nrow(filtered_data)))
}

# Create the plot showing tradeoffs between n100 proportion & sample retained
g <- ggplot(results, aes(x = SampleCount, y = X)) +
  geom_line() +
  geom_hline(yintercept = n100_cutoff*100, linetype = "dashed", color = "black", alpha = .5) +
  labs(y = "Proportion of loci over 100 Reads", x = "Sample count") +
  scale_x_continuous(breaks = seq(0, 1250, 100))
g

}




plot_n100_sample_cutoff(filtered_data2, n100_cutoff = 0.5)


### Remove samples with too low n100
remove_n100_cutoff <- function(data, n100_cutoff = 0.5){
  # Group by sample + loci category
  Locus_cov_100 <- data %>% 
    group_by(SampleID, Locus, Category, Pool) %>% 
    summarize(Reads = sum(Reads)) %>% 
    ungroup() 
  
  # Calculate n100 and n50 loci counts
  Locus_cov_100 <- Locus_cov_100 %>%
    group_by(SampleID, Category, Pool) %>% 
    summarize(total_Reads = sum(Reads), n100 = sum(Reads >= 100)) 
  
  # Normalization - calculate % loci with n100 or n50 (Species targets are not counted)
  amp_cov_category_norm <- Locus_cov_100 %>% 
    filter(Category != "Species") %>%
    mutate(norm_100 = case_when(
    Category == "Diversity" ~ n100/165,
    Category == "Resistance" ~ n100/38,
    Category == "Immune" ~ n100/5,
    Category == "Diagnostic" ~ n100/30))
  
  # Remove samples with diversity n100 < 50%
  sample_n100_fail <-  amp_cov_category_norm %>% filter(Category == "Diversity") %>% filter(norm_100 < n100_cutoff) %>% pull(SampleID)
  clean_data <- data %>% filter(! SampleID %in% sample_n100_fail) 
  sample_size <- n_distinct(clean_data$SampleID)
  clean_data_info <- sprintf("Step 3 - Sample size after removing samples with less than n100 cutoff = %s, Retained prop. = %s",   sample_size, round(sample_size/n_distinct(data$SampleID), 2))
  
  return(list(clean_data = clean_data, clean_data_info = clean_data_info, removed_SampleID = sample_n100_fail))
}


# Example usage:
result <- remove_n100_cutoff(filtered_data2, n100_cutoff = 0.5)
filtered_data3 <- result$clean_data  # This will give you the data after removing low-loci-coverage samples
print(result$clean_data_info)  # This will print the information of filtered data compared to original input data
sample_n100_fail <- print(result$removed_SampleID) # This will print the information of removed SampleID

##WL: export names of the filtered samples
filtered_list3 <- data.frame(as.list(result$removed_SampleID)) %>%
  transpose() %>%
  rename("SampleID" = "V1") %>%
  mutate(StudySubject = word(SampleID, 1, sep = "_")) %>%
  select(SampleID, StudySubject) %>%
  mutate(Reason = "Less than 50% loci covered")

```


**Export list of sample IDs that need to be redone**
```{r redos, echo=FALSE, warning = FALSE, message= FALSE}

### Filtered data: Remove bad-quality samples only 
redos <- rbind(filtered_list1, filtered_list2, filtered_list3)

filtered_data_final <- filtered_data3 %>% 
  group_by(SampleID, Locus) %>%
  mutate(norm.Reads.Locus = Reads/sum(Reads)) %>% 
  ungroup() %>% 
  filter(Category == "Diversity")


```


