---
title: "Amplicon Sequence Run QC"
author: "William and Yisihak"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
suppressPackageStartupMessages({
  library(dplyr)
  library(ggplot2)
  library(tidyr)
  library(tidyverse)
  library(stringr)
  library(purrr)
  library(data.table)
  library(vctrs)
  library(plotly)
  library(readxl)
  library(scales)
  library(htmltools)
})

options(stringsAsFactors = FALSE)

```

# import allele data and set configurations

Each run directory should contain `allele_data.txt`.

```{r configuration}
# One directory per sequence run.
# The directory name is used as the default RunID.

run_dirs <- c(
  "../data/QC/Input/run4",
  "../data/QC/Input/run5",
  "../data/QC/Input/run9"
)

# Optional: explicitly give RunIDs in the same order as run_dirs.
# Leave as NULL to use the directory names.
run_ids <- NULL

# Shared pool metadata files.
pool_1A_file <- "../data/QC/Input/mad4hatter_pool_info/Pool1A-Diversity.tsv"
pool_1B_file <- "../data/QC/Input/mad4hatter_pool_info/Pool1B-Resitance+.tsv"
pool_2_file  <- "../data/QC/Input/mad4hatter_pool_info/Pool2-Resistance+.tsv"

# Output directory. A separate subdirectory is created for every run.
output_dir <- "../data/QC/Results"

# Set TRUE if Allele should be generated from PseudoCIGAR.
# Set FALSE if the input already contains the desired Allele column.
assign_allele_from_pseudocigar <- FALSE

# QC thresholds retained from the original script.
TOTAL_READS_LOCI <- 248
TOTAL_READS_PER_LOCI <- 50
LOCUS_COVERAGE_FRACTION <- 0.70
N100_CUTOFF <- 0.50
LOW_READ_LOCUS_CUTOFF <- 10
LOW_READ_LOCUS_POPULATION_FRACTION <- 1/3
```

# Helper functions

The following helpers are used throughout the QC. They provide **empty-data protection**, so an empty run/control/filter does not cause the whole report to fail.

```{r helpers}
is_empty_df <- function(x) {
  is.null(x) || !is.data.frame(x) || nrow(x) == 0
}

safe_n_distinct <- function(x) {
  if (length(x) == 0) 0L else dplyr::n_distinct(x)
}

safe_mean <- function(x, digits = 2) {
  if (length(x) == 0 || all(is.na(x))) return(NA_real_)
  round(mean(x, na.rm = TRUE), digits)
}

safe_min <- function(x) {
  if (length(x) == 0 || all(is.na(x))) return(NA_real_)
  min(x, na.rm = TRUE)
}

safe_max <- function(x) {
  if (length(x) == 0 || all(is.na(x))) return(NA_real_)
}

safe_plot <- function(plot_obj, empty_title = "No data available") {
  if (inherits(plot_obj, "ggplot")) return(plot_obj)
  ggplot() +
    theme_void() +
    ggtitle(empty_title)
}

safe_ggplotly <- function(plot_obj, empty_title = "No data available") {
  if (inherits(plot_obj, "ggplot")) {
    return(ggplotly(plot_obj))
  }
  ggplotly(
    ggplot() +
      theme_void() +
      ggtitle(empty_title)
  )
}

read_tsv_safe <- function(path, ...) {
  if (!file.exists(path)) {
    stop("File not found: ", path)
  }
  readr::read_tsv(path, trim_ws = TRUE, show_col_types = FALSE, ...)
}

make_empty_redo <- function() {
  data.frame(
    SampleID = character(),
    StudySubject = character(),
    Reason = character(),
    stringsAsFactors = FALSE
  )
}

make_redo_table <- function(ids, reason) {
  ids <- unique(ids)
  if (length(ids) == 0) return(make_empty_redo())

  data.frame(
    SampleID = ids,
    StudySubject = stringr::word(ids, 1, sep = "_"),
    Reason = reason,
    stringsAsFactors = FALSE
  )
}

safe_write_csv <- function(x, path) {
  if (is.null(x) || !is.data.frame(x)) x <- data.frame()
  dir.create(dirname(path), recursive = TRUE, showWarnings = FALSE)
  write.csv(x, path, row.names = FALSE)
}

run_name_from_path <- function(path) {
  nm <- basename(normalizePath(path, winslash = "/", mustWork = FALSE))
  if (is.na(nm) || nm == "") basename(dirname(path)) else nm
}

# Return a ggplot object only when the input contains data.
empty_plot <- function(title) {
  ggplot() +
    theme_void() +
    ggtitle(title)
}
```

# Load shared pool metadata

```{r load-pool-metadata}
pool_1A <- read_tsv_safe(pool_1A_file) %>%
  rename(Locus = `locus-pool`)

pool_1B <- read_tsv_safe(pool_1B_file) %>%
  rename(Locus = `locus-pool`)

pool_2 <- read_tsv_safe(pool_2_file) %>%
  rename(Locus = `locus-pool`)

pool_1A_neg <- pool_1A %>% select(Locus, Category)
pool_1B_neg <- pool_1B %>% select(Locus, `Reason to include (if drug resistance: aminoacids)`)
pool_2_neg  <- pool_2 %>% select(Locus, `Reason to include (if drug resistance: aminoacids)`)

pool_neg <- bind_rows(
  pool_1A_neg,
  pool_1B_neg,
  pool_2_neg
)

pool_1Ax <- pool_1A %>% select(Locus, Category)
pool_1Bx <- pool_1B %>% select(Locus, Category)
pool_2x  <- pool_2 %>% select(Locus, Category)
pool_cat <- bind_rows(pool_1Ax, pool_1Bx, pool_2x)

pool_gene <- bind_rows(
  pool_1B %>% select(Locus, Category, Gene),
  pool_2 %>% select(Locus, Category, Gene)
)
```

# Input validation and run loader

```{r input-functions}
validate_configuration <- function(run_dirs, run_ids = NULL) {
  if (length(run_dirs) == 0) {
    stop("No sequence-run directories were supplied.")
  }

  if (!is.null(run_ids) && length(run_ids) != length(run_dirs)) {
    stop("run_ids must have the same length as run_dirs.")
  }

  missing_dirs <- run_dirs[!dir.exists(run_dirs)]
  if (length(missing_dirs) > 0) {
    stop(
      "The following run directories do not exist:\n",
      paste(missing_dirs, collapse = "\n")
    )
  }
}

load_allele_data <- function(run_dir, run_id, assign_allele = TRUE) {
  allele_file <- file.path(run_dir, "allele_data.txt")

  if (!file.exists(allele_file)) {
    warning("Skipping run '", run_id, "': allele_data.txt was not found.")
    return(NULL)
  }

  allele_data <- read.table(
    allele_file,
    sep = "\t",
    header = TRUE,
    check.names = FALSE,
    stringsAsFactors = FALSE
  )

  if (is_empty_df(allele_data)) {
    warning("Run '", run_id, "' contains an empty allele_data.txt.")
    return(NULL)
  }

  required_cols <- c("SampleID", "Locus", "Reads")
  missing_cols <- setdiff(required_cols, names(allele_data))

  if (length(missing_cols) > 0) {
    stop(
      "Run '", run_id, "' is missing required columns: ",
      paste(missing_cols, collapse = ", ")
    )
  }

  if (!"Allele" %in% names(allele_data)) {
    if (!"PseudoCIGAR" %in% names(allele_data)) {
      stop(
        "Run '", run_id,
        "' has neither Allele nor PseudoCIGAR. Cannot create Allele."
      )
    }
    assign_allele <- TRUE
  }

  if (assign_allele) {
    if (!"PseudoCIGAR" %in% names(allele_data)) {
      stop(
        "assign_allele_from_pseudocigar=TRUE but PseudoCIGAR is missing in run '",
        run_id, "'."
      )
    }

    cigar <- allele_data %>%
      select(Locus, PseudoCIGAR) %>%
      distinct() %>%
      group_by(Locus) %>%
      arrange(Locus, PseudoCIGAR, .by_group = TRUE) %>%
      mutate(type = paste0("t", cur_group_id())) %>%
      ungroup()

    cigar$Allele <- paste(cigar$type, ave(cigar$type, cigar$type, FUN = seq_along), sep = ".")

    allele_data <- allele_data %>%
      left_join(cigar %>% select(Locus, PseudoCIGAR, Allele),
                by = c("Locus", "PseudoCIGAR"))
  }

  allele_data <- allele_data %>%
    distinct(SampleID, Locus, PseudoCIGAR, Allele, .keep_all = TRUE) %>%
    group_by(SampleID, Locus) %>%
    mutate(
      norm.reads.locus = ifelse(sum(Reads, na.rm = TRUE) > 0,
                                Reads / sum(Reads, na.rm = TRUE),
                                0),
      n.alleles = n()
    ) %>%
    arrange(SampleID, Locus) %>%
    ungroup()

  allele_data
}
```

# QC functions

## Preview data structure

```{r preview-functions}
preview_data <- function(data) {
  if (is_empty_df(data)) {
    return(list(
      sample_size = 0,
      Locus_no = 0,
      allele_count = data.frame(),
      clone_wide = data.frame()
    ))
  }

  sample_size <- safe_n_distinct(data$SampleID)
  Locus_no <- safe_n_distinct(data$Locus)

  allele_count <- data %>%
    group_by(Locus) %>%
    summarize(n.alleles = n_distinct(Allele), .groups = "drop") %>%
    arrange(desc(n.alleles))

  clone <- data %>%
    group_by(SampleID, Locus) %>%
    summarise(total_alleles = n(), .groups = "drop") %>%
    mutate(Clone = ifelse(total_alleles == 1, "Mono", "Poly")) %>%
    group_by(SampleID) %>%
    count(Clone) %>%
    pivot_wider(names_from = Clone, values_from = n, values_fill = 0)

  if (!"Mono" %in% names(clone)) clone$Mono <- 0
  if (!"Poly" %in% names(clone)) clone$Poly <- 0

  clone_wide <- clone %>%
    mutate(
      Total_alleles = Mono + Poly,
      Mono_prop = ifelse(Total_alleles > 0, Mono / Total_alleles, 0),
      Poly_prop = ifelse(Total_alleles > 0, Poly / Total_alleles, 0)
    )

  list(
    sample_size = sample_size,
    Locus_no = Locus_no,
    allele_count = allele_count,
    clone_wide = clone_wide
  )
}

plot_total_alleles <- function(data) {
  if (is_empty_df(data)) return(NULL)

  allele_table <- data %>%
    group_by(Locus) %>%
    summarize(total_alleles = n_distinct(Allele), .groups = "drop") %>%
    arrange(desc(total_alleles))

  if (nrow(allele_table) == 0) return(NULL)

  q95 <- quantile(allele_table$total_alleles, 0.95, na.rm = TRUE)

  allele_table %>%
    mutate(alleles_over_q95 = ifelse(total_alleles > q95, "Y", "N")) %>%
    ggplot(aes(x = Locus, y = total_alleles, color = alleles_over_q95)) +
    geom_point() +
    theme(axis.text.x = element_text(angle = 90, hjust = 0, size = 5)) +
    scale_color_hue(direction = -1) +
    ggtitle("Which loci are polyallelic?")
}
```

## Negative controls

```{r negative-controls-functions}
negative_control_check <- function(data) {
  if (is_empty_df(data)) {
    return(list(
      negatives = data.frame(),
      bad_negatives = data.frame(),
      loci_neg = data.frame(),
      loci_neg_info = data.frame()
    ))
  }

  negatives <- data %>%
    filter(group == "NC") %>%
    group_by(SampleID, Locus) %>%
    summarize(Reads = sum(Reads, na.rm = TRUE), .groups = "drop")

  bad_negatives <- negatives %>%
    filter(Reads > 100) %>%
    arrange(Reads)

  loci_neg <- negatives %>%
    group_by(Locus) %>%
    summarise(Reads = sum(Reads, na.rm = TRUE), .groups = "drop") %>%
    mutate(pool = sapply(strsplit(Locus, "-"), tail, 1))

  loci_neg_info <- loci_neg %>%
    left_join(pool_neg, by = "Locus") %>%
    filter(Reads >= 100) %>%
    arrange(desc(Reads))

  list(
    negatives = negatives,
    bad_negatives = bad_negatives,
    loci_neg = loci_neg,
    loci_neg_info = loci_neg_info
  )
}

plot_negative_read_histogram <- function(negatives) {
  if (is_empty_df(negatives)) return(NULL)

  ggplot(negatives, aes(x = Reads)) +
    geom_histogram() +
    geom_vline(xintercept = 100, linetype = "dotted", col = "darkred") +
    xlab("Number of Reads at Locus") +
    ggtitle("How many total Reads do neg.ctrls have across each Locus?")
}

plot_negative_loci <- function(loci_neg) {
  if (is_empty_df(loci_neg)) return(NULL)

  ggplot(loci_neg, aes(y = Reads, x = Locus)) +
    geom_bar(position = "dodge", stat = "identity") +
    geom_hline(yintercept = 100, linetype = "dotted", col = "darkred") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 3)) +
    ggtitle("What is the total number of neg.ctrl Reads, by pool?") +
    facet_wrap(~pool, scales = "free_x")
}
```

## Positive controls

```{r positive-control-functions}
plot_pc_clones <- function(data) {
  if (is_empty_df(data)) return(NULL)

  data_pc <- data %>% filter(group == "PC")
  if (is_empty_df(data_pc)) return(NULL)

  pc_clone <- data_pc %>%
    group_by(SampleID, Locus) %>%
    summarise(total_alleles = n(), .groups = "drop") %>%
    mutate(Clone = ifelse(total_alleles == 1, "Mono", "Poly")) %>%
    group_by(SampleID) %>%
    count(Clone)

  if (is_empty_df(pc_clone)) return(NULL)

  ggplot(pc_clone, aes(x = SampleID, y = n)) +
    geom_col(aes(fill = Clone)) +
    theme(
      axis.text.y = element_text(size = 8),
      axis.text.x = element_text(angle = 90, size = 7)
    ) +
    scale_y_continuous(name = "Locus No.", breaks = seq(0, 200, by = 50))
}

plot_pc_polyclonal_loci <- function(data) {
  if (is_empty_df(data)) return(NULL)

  positives <- data %>%
    filter(group == "PC") %>%
    group_by(SampleID, Locus) %>%
    mutate(
      read_fraction = ifelse(sum(Reads, na.rm = TRUE) > 0,
                             100 * Reads / sum(Reads, na.rm = TRUE),
                             0),
      n.alleles = n()
    ) %>%
    filter(norm.reads.locus > 0.01) %>%
    mutate(Clone = ifelse(n.alleles == 1, "Mono", "Poly")) %>%
    ungroup()

  pc_poly_allele <- positives %>% filter(n.alleles > 1)
  if (is_empty_df(pc_poly_allele)) return(NULL)

  ggplot(pc_poly_allele) +
    geom_point(aes(x = read_fraction, y = log10(Reads), color = Locus)) +
    geom_hline(yintercept = 2, color = "black", linetype = "dashed")
}

pc_consistency <- function(data) {
  if (is_empty_df(data)) return(data.frame())

  allele_pc <- data %>%
    filter(group == "PC") %>%
    group_by(SampleID, Locus, Allele, Reads) %>%
    summarise(.groups = "drop") %>%
    group_by(Locus, Allele) %>%
    mutate(duplicated = n() > 1) %>%
    ungroup()

  if (is_empty_df(allele_pc)) return(data.frame())

  allele_pc %>%
    group_by(SampleID) %>%
    summarize(
      pc_dup_prop = mean(duplicated == TRUE),
      .groups = "drop"
    )
}

plot_pc_clones_filt <- function(data) {
  if (is_empty_df(data)) return(NULL)

  data_pc_filt <- data %>%
    filter(group == "PC", Reads > 10) %>%
    group_by(SampleID, Locus) %>%
    mutate(
      read_fraction = ifelse(sum(Reads, na.rm = TRUE) > 0,
                             Reads / sum(Reads, na.rm = TRUE),
                             0)
    ) %>%
    filter(read_fraction > 0.01) %>%
    mutate(
      n.alleles = n(),
      Clone = ifelse(n.alleles == 1, "Mono", "Poly")
    ) %>%
    ungroup()

  if (is_empty_df(data_pc_filt)) return(NULL)

  pc_clone <- data_pc_filt %>%
    group_by(SampleID) %>%
    count(Clone)

  ggplot(pc_clone, aes(x = SampleID, y = n)) +
    geom_col(aes(fill = Clone)) +
    theme(
      axis.text.y = element_text(size = 8),
      axis.text.x = element_text(angle = 90, size = 7)
    ) +
    scale_y_continuous(name = "Locus No.", breaks = seq(0, 200, by = 50))
}

plot_pc_polyclonal_loci_filt <- function(data) {
  if (is_empty_df(data)) return(NULL)

  data_pc_filt <- data %>%
    filter(group == "PC", Reads > 10) %>%
    group_by(SampleID, Locus) %>%
    mutate(read_fraction = ifelse(sum(Reads, na.rm = TRUE) > 0,
                                   Reads / sum(Reads, na.rm = TRUE), 0)) %>%
    filter(read_fraction > 0.01) %>%
    mutate(
      n.alleles = n(),
      Clone = ifelse(n.alleles == 1, "Mono", "Poly")
    ) %>%
    ungroup()

  pc_poly_allele <- data_pc_filt %>% filter(n.alleles > 1)
  if (is_empty_df(pc_poly_allele)) return(NULL)

  ggplot(pc_poly_allele) +
    geom_point(aes(x = read_fraction, y = log10(Reads), color = Locus))
}
```

# Step 1 — Total reads

```{r total-read-functions}
plot_Reads_dist <- function(data) {
  if (is_empty_df(data)) return(NULL)

  Reads_table <- data %>%
    distinct(SampleID, Locus, Allele, .keep_all = TRUE) %>%
    group_by(SampleID) %>%
    summarize(Total_Reads = sum(Reads, na.rm = TRUE), .groups = "drop") %>%
    arrange(Total_Reads)

  if (nrow(Reads_table) == 0) return(NULL)

  Reads_table <- Reads_table %>%
    mutate(label = case_when(
      grepl("NC", SampleID) ~ "NC",
      grepl("NTC", SampleID) ~ "NC",
      grepl("3D7", SampleID) ~ "PC",
      grepl("Dd2", SampleID) ~ "PC",
      grepl("DD2", SampleID) ~ "PC",
      grepl("HB3", SampleID) ~ "PC",
      grepl("K1", SampleID) ~ "PC",
      grepl("UCSF", SampleID) ~ "PC",
      grepl("PC", SampleID) ~ "PC",
      TRUE ~ "Sample"
    ))

  ggplot(Reads_table, aes(x = log10(Total_Reads + 1), fill = label)) +
    geom_density(alpha = .5) +
    ggtitle("Total Reads distribution")
}

plot_Reads_cutoff <- function(data, loci = 248, Reads_cutoff = 50) {
  if (is_empty_df(data)) return(NULL)

  Reads_table <- data %>%
    distinct(SampleID, Locus, Allele, .keep_all = TRUE) %>%
    group_by(SampleID) %>%
    summarize(Total_Reads = sum(Reads, na.rm = TRUE), .groups = "drop") %>%
    arrange(Total_Reads)

  if (nrow(Reads_table) == 0) return(NULL)

  pc <- unique(data$SampleID[data$group == "PC"])
  nc <- unique(data$SampleID[data$group == "NC"])
  controls <- unique(c(pc, nc))

  Reads_criterion <- loci * Reads_cutoff

  sample_low_Reads <- Reads_table %>%
    filter(!SampleID %in% controls, Total_Reads < Reads_criterion) %>%
    pull(SampleID)

  Reads_table <- Reads_table %>%
    mutate(
      QC = case_when(
        SampleID %in% pc ~ "Positive controls",
        SampleID %in% nc ~ "Negative controls",
        SampleID %in% sample_low_Reads ~ "Low-Reads samples",
        TRUE ~ "QC-passed samples"
      )
    )

  level_order <- Reads_table %>%
    arrange(factor(QC, levels = c(
      "Negative controls",
      "Low-Reads samples",
      "Positive controls",
      "QC-passed samples"
    ))) %>%
    pull(SampleID)

  color_fill <- c(
    "Positive controls" = "red",
    "Negative controls" = "green",
    "Low-Reads samples" = "gray",
    "QC-passed samples" = "black"
  )

  ggplot(
    Reads_table,
    aes(x = factor(SampleID, levels = level_order),
        y = log10(Total_Reads + 1),
        fill = QC)
  ) +
    geom_bar(stat = "identity") +
    coord_flip() +
    scale_fill_manual(values = color_fill) +
    theme_classic() +
    theme(axis.text.y = element_blank()) +
    geom_hline(
      yintercept = log10(Reads_criterion + 1),
      linetype = "dashed",
      color = "red"
    ) +
    ylab("Log10 Total Reads") +
    xlab("SampleID") +
    ggtitle("Filter out samples with <50*248 reads")
}

remove_low_Reads_samples <- function(data, loci = 248, Reads_cutoff = 50) {
  if (is_empty_df(data)) {
    return(list(
      clean_data = data,
      clean_data_info = "Step 1 - No input data.",
      removed_SampleID = character()
    ))
  }

  Reads_table <- data %>%
    distinct(SampleID, Locus, Allele, .keep_all = TRUE) %>%
    group_by(SampleID) %>%
    summarize(Total_Reads = sum(Reads, na.rm = TRUE), .groups = "drop")

  pc <- unique(data$SampleID[data$group == "PC"])
  nc <- unique(data$SampleID[data$group == "NC"])
  controls <- unique(c(pc, nc))

  Reads_criterion <- loci * Reads_cutoff

  sample_low_Reads <- Reads_table %>%
    filter(!SampleID %in% controls, Total_Reads < Reads_criterion) %>%
    pull(SampleID)

  clean_data <- data %>%
    filter(!SampleID %in% unique(c(controls, sample_low_Reads)))

  original_n <- safe_n_distinct(data$SampleID)
  final_n <- safe_n_distinct(clean_data$SampleID)

  list(
    clean_data = clean_data,
    clean_data_info = sprintf(
      "Step 1 - Sample size after removing samples with insufficient reads = %s, Retained prop. = %s",
      final_n,
      ifelse(original_n > 0, round(final_n / original_n, 2), NA)
    ),
    removed_SampleID = sample_low_Reads
  )
}
```

# Step 2 — Loci coverage

```{r loci-functions}
plot_loci_dist <- function(data) {

  if (is_empty_df(data)) {
    return(NULL)
  }

  # Count the number of loci represented for each sample
  loci_by_sample <- data %>%
    group_by(SampleID, Locus) %>%
    summarise(
      total_loci = n(),
      .groups = "drop"
    ) %>%
    group_by(SampleID) %>%
    summarise(
      Total_loci = n_distinct(Locus),
      .groups = "drop"
    )

  if (is_empty_df(loci_by_sample)) {
    return(NULL)
  }

  ggplot(loci_by_sample, aes(x = Total_loci)) +
    geom_histogram() +
    scale_x_continuous(breaks = seq(0, 300, 25)) +
    xlab("Total loci") +
    ggtitle("Loci distribution")
}


plot_loci_cutoff <- function(
    data,
    loci = 248,
    perc_covered = 0.7
) {

  if (is_empty_df(data)) {
    return(NULL)
  }

  # Count unique loci represented in each sample
  loci_by_sample <- data %>%
    group_by(SampleID) %>%
    summarise(
      Total_loci = n_distinct(Locus),
      .groups = "drop"
    )

  if (is_empty_df(loci_by_sample)) {
    return(NULL)
  }

  # Calculate cutoff
  loci_criterion <- round(loci * perc_covered)

  # Classify samples
  loci_by_sample <- loci_by_sample %>%
    mutate(
      QC = ifelse(
        Total_loci < loci_criterion,
        "Below cutoff",
        "QC-passed"
      )
    ) %>%
    arrange(Total_loci)

  ggplot(
    loci_by_sample,
    aes(
      x = reorder(SampleID, Total_loci),
      y = Total_loci
    )
  ) +
    geom_col(aes(fill = QC)) +
    coord_flip() +
    scale_y_continuous(
      name = "No. of Locus",
      breaks = seq(0, 300, by = 50)
    ) +
    geom_hline(
      yintercept = loci_criterion,
      linetype = "dotted",
      color = "black"
    ) +
    xlab("SampleID") +
    ggtitle("Filter out samples with insufficient loci coverage")
}

remove_low_loci_samples <- function(data, loci = 248, perc_covered = 0.7) {
  if (is_empty_df(data)) {
    return(list(
      clean_data = data,
      clean_data_info = "Step 2 - No input data.",
      removed_SampleID = character()
    ))
  }

  clone <- data %>%
    group_by(SampleID, Locus) %>%
    summarise(total_loci = n(), .groups = "drop") %>%
    mutate(Clone = ifelse(total_loci == 1, "Mono", "Poly")) %>%
    group_by(SampleID) %>%
    count(Clone)

  loci_criterion <- round(loci * perc_covered)

  sample_less_loci <- clone %>%
    group_by(SampleID) %>%
    summarise(total_loci = sum(n), .groups = "drop") %>%
    filter(total_loci < loci_criterion) %>%
    pull(SampleID)

  clean_data <- data %>%
    filter(!SampleID %in% sample_less_loci)

  original_n <- safe_n_distinct(data$SampleID)
  final_n <- safe_n_distinct(clean_data$SampleID)

  list(
    clean_data = clean_data,
    clean_data_info = sprintf(
      "Step 2 - Sample size after removing samples with insufficient loci coverage = %s, Retained prop. = %s",
      final_n,
      ifelse(original_n > 0, round(final_n / original_n, 2), NA)
    ),
    removed_SampleID = sample_less_loci
  )
}
```

# Step 3 — Proportion of loci with >100 Reads

```{r n100-functions}
plot_n100_loci_cutoff <- function(data, n100_cutoff = 0.5) {
  if (is_empty_df(data)) return(NULL)

  Locus_cov_100 <- data %>%
    group_by(SampleID, Locus, Category, Pool) %>%
    summarize(Reads = sum(Reads, na.rm = TRUE), .groups = "drop") %>%
    group_by(SampleID, Category, Pool) %>%
    summarize(
      total_Reads = sum(Reads, na.rm = TRUE),
      n100 = sum(Reads >= 100),
      .groups = "drop"
    )

  amp_cov_category_norm <- Locus_cov_100 %>%
    filter(Category != "Species") %>%
    mutate(norm_100 = case_when(
      Category == "Diversity" ~ n100 / 165,
      Category == "Resistance" ~ n100 / 38,
      Category == "Immune" ~ n100 / 5,
      Category == "Diagnostic" ~ n100 / 30,
      TRUE ~ NA_real_
    )) %>%
    filter(!is.na(norm_100))

  if (is_empty_df(amp_cov_category_norm)) return(NULL)

  ggplot(amp_cov_category_norm, aes(x = total_Reads, y = norm_100)) +
    geom_point(aes(color = "red", alpha = 0.5)) +
    scale_x_log10() +
    geom_hline(yintercept = n100_cutoff, linetype = "dashed", color = "black") +
    facet_wrap(~Category) +
    ggtitle("Normalized prop. of loci with > 100 Reads in each sample") +
    ylim(0, 1) +
    ylab("Proportion of loci with >100 Reads") +
    xlab("Total Reads")
}

plot_n100_sample_cutoff <- function(data, n100_cutoff = 0.5) {
  if (is_empty_df(data)) return(NULL)

  Locus_cov_100 <- data %>%
    group_by(SampleID, Locus, Category, Pool) %>%
    summarize(Reads = sum(Reads, na.rm = TRUE), .groups = "drop")

  n_table <- Locus_cov_100 %>%
    filter(Category == "Diversity") %>%
    group_by(SampleID, Locus) %>%
    summarize(total_Reads = sum(Reads, na.rm = TRUE), .groups = "drop") %>%
    mutate(n100 = ifelse(total_Reads >= 100, 1, 0))

  n_count <- n_table %>%
    group_by(SampleID) %>%
    summarise(
      n100_0_count = sum(n100 == 0),
      n100_1_count = sum(n100 == 1),
      n100_proportion = n100_1_count / n(),
      .groups = "drop"
    )

  if (is_empty_df(n_count)) return(NULL)

  results <- tibble(
    X = 0:100,
    SampleCount = map_int(
      0:100,
      ~ sum(n_count$n100_proportion >= (.x / 100), na.rm = TRUE)
    )
  )

  ggplot(results, aes(x = SampleCount, y = X)) +
    geom_line() +
    geom_hline(
      yintercept = n100_cutoff * 100,
      linetype = "dashed",
      color = "black",
      alpha = .5
    ) +
    labs(
      y = "Proportion of loci over 100 Reads",
      x = "Sample count"
    ) +
    scale_x_continuous(breaks = seq(0, 1250, 100))
}

remove_n100_cutoff <- function(data, n100_cutoff = 0.5) {
  if (is_empty_df(data)) {
    return(list(
      clean_data = data,
      clean_data_info = "Step 3 - No input data.",
      removed_SampleID = character()
    ))
  }

  Locus_cov_100 <- data %>%
    group_by(SampleID, Locus, Category, Pool) %>%
    summarize(Reads = sum(Reads, na.rm = TRUE), .groups = "drop") %>%
    group_by(SampleID, Category, Pool) %>%
    summarize(
      total_Reads = sum(Reads, na.rm = TRUE),
      n100 = sum(Reads >= 100),
      .groups = "drop"
    )

  amp_cov_category_norm <- Locus_cov_100 %>%
    filter(Category != "Species") %>%
    mutate(norm_100 = case_when(
      Category == "Diversity" ~ n100 / 165,
      Category == "Resistance" ~ n100 / 38,
      Category == "Immune" ~ n100 / 5,
      Category == "Diagnostic" ~ n100 / 30,
      TRUE ~ NA_real_
    )) %>%
    filter(!is.na(norm_100))

  sample_n100_fail <- amp_cov_category_norm %>%
    filter(Category == "Diversity", norm_100 < n100_cutoff) %>%
    pull(SampleID) %>%
    unique()

  clean_data <- data %>%
    filter(!SampleID %in% sample_n100_fail)

  original_n <- safe_n_distinct(data$SampleID)
  final_n <- safe_n_distinct(clean_data$SampleID)

  list(
    clean_data = clean_data,
    clean_data_info = sprintf(
      "Step 3 - Sample size after removing samples with less than n100 cutoff = %s, Retained prop. = %s",
      final_n,
      ifelse(original_n > 0, round(final_n / original_n, 2), NA)
    ),
    removed_SampleID = sample_n100_fail
  )
}
```

# Locus performance check

```{r locus-performance}
locus_performance_check <- function(data) {
  if (is_empty_df(data)) {
    return(list(
      target_low_Reads = data.frame(),
      gene_low_Reads = data.frame(),
      gene_low_Reads_freq = data.frame(),
      gene_low_Reads_popcut = data.frame()
    ))
  }

  Locus_check <- data %>%
    group_by(SampleID, Locus) %>%
    summarize(total_Reads = sum(Reads, na.rm = TRUE), .groups = "drop")

  target_low_Reads <- Locus_check %>%
    filter(total_Reads <= LOW_READ_LOCUS_CUTOFF)

  gene_low_Reads <- target_low_Reads %>%
    left_join(pool_gene, by = "Locus")

  gene_low_Reads_freq <- gene_low_Reads %>%
    count(Locus, name = "count") %>%
    mutate(
      freq = count / safe_n_distinct(data$SampleID)
    )

  gene_low_Reads_popcut <- gene_low_Reads_freq %>%
    filter(freq > LOW_READ_LOCUS_POPULATION_FRACTION)

  list(
    target_low_Reads = target_low_Reads,
    gene_low_Reads = gene_low_Reads,
    gene_low_Reads_freq = gene_low_Reads_freq,
    gene_low_Reads_popcut = gene_low_Reads_popcut
  )
}
```

# Per-run QC pipeline

```{r run-qc-function}
process_run <- function(run_dir, run_id) {

  message("================================================")
  message("Processing run: ", run_id)
  message("================================================")

  raw <- load_allele_data(
    run_dir = run_dir,
    run_id = run_id,
    assign_allele = assign_allele_from_pseudocigar
  )

  if (is.null(raw) || is_empty_df(raw)) {
    return(list(
      RunID = run_id,
      RunDir = run_dir,
      status = "EMPTY_OR_MISSING",
      data = data.frame()
    ))
  }

  # Merge/control classification retained from original script.
  data <- raw %>%
    mutate(
      group = case_when(
      grepl("NC", SampleID) ~ "NC",
      grepl("NTC", SampleID) ~ "NC",
      grepl("3D7", SampleID) ~ "PC",
      grepl("Dd2", SampleID) ~ "PC",
      grepl("DD2", SampleID) ~ "PC",
      grepl("HB3", SampleID) ~ "PC",
      grepl("K1", SampleID) ~ "PC",
      grepl("UCSF", SampleID) ~ "PC",
      grepl("PC", SampleID) ~ "PC",
        TRUE ~ "Sample"
      )
    )

  # Preview.
  preview <- preview_data(data)

  # -----------------------------
  # Control checks
  # -----------------------------
  neg <- negative_control_check(data)
  pc_consistency_table <- pc_consistency(data)

  # -----------------------------
  # Step 1: total reads
  # -----------------------------
  step1 <- remove_low_Reads_samples(
    data,
    loci = TOTAL_READS_LOCI,
    Reads_cutoff = TOTAL_READS_PER_LOCI
  )
  filtered_data1 <- step1$clean_data

  filtered_list1 <- make_redo_table(
    step1$removed_SampleID,
    "Few Total Reads"
  )

  # -----------------------------
  # Step 2: loci coverage
  # -----------------------------
  step2 <- remove_low_loci_samples(
    filtered_data1,
    loci = TOTAL_READS_LOCI,
    perc_covered = LOCUS_COVERAGE_FRACTION
  )
  filtered_data2 <- step2$clean_data

  filtered_list2 <- make_redo_table(
    step2$removed_SampleID,
    "Insufficient locus coverage"
  )

  # Add pool/category metadata.
  if (!is_empty_df(filtered_data2)) {
    filtered_data2 <- filtered_data2 %>%
      mutate(Pool = sapply(strsplit(Locus, "-"), tail, 1)) %>%
      left_join(pool_cat, by = "Locus")
  }

  # -----------------------------
  # Step 3: n100
  # -----------------------------
  step3 <- remove_n100_cutoff(
    filtered_data2,
    n100_cutoff = N100_CUTOFF
  )
  filtered_data3 <- step3$clean_data

  filtered_list3 <- make_redo_table(
    step3$removed_SampleID,
    "Less than 50% loci covered"
  )

  # -----------------------------
  # Locus performance
  # -----------------------------
  locus_check <- locus_performance_check(filtered_data3)

  redos <- bind_rows(
    filtered_list1,
    filtered_list2,
    filtered_list3
  ) %>%
    distinct()

  list(
    RunID = run_id,
    RunDir = run_dir,
    status = "OK",
    data = data,
    preview = preview,
    negatives = neg,
    pc_consistency = pc_consistency_table,
    filtered_data1 = filtered_data1,
    filtered_data2 = filtered_data2,
    filtered_data3 = filtered_data3,
    filtered_list1 = filtered_list1,
    filtered_list2 = filtered_list2,
    filtered_list3 = filtered_list3,
    redos = redos,
    locus_check = locus_check
  )
}
```

# Run all sequence runs

```{r run-all}
validate_configuration(run_dirs, run_ids)

if (is.null(run_ids)) {
  run_ids <- vapply(run_dirs, run_name_from_path, character(1))
}

dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

run_results <- map2(
  run_dirs,
  run_ids,
  ~ process_run(.x, .y)
)

names(run_results) <- run_ids


run_summary <- tibble(
  RunID = run_ids,
  RunDir = run_dirs,
  Status = map_chr(run_results, function(x) {
    x$status
  }),
  Samples = map_int(run_results, function(x) {
    if (x$status == "OK") {
      safe_n_distinct(x$data$SampleID)
    } else {
      0L
    }
  }),
  Loci = map_int(run_results, function(x) {
    if (x$status == "OK") {
      safe_n_distinct(x$data$Locus)
    } else {
      0L
    }
  }),
  Samples_Redo = map_int(run_results, function(x) {
    if (x$status == "OK") {
      nrow(x$redos)
    } else {
      0L
    }
  })
)

run_summary

```

# QC results — one section per sequence run

```{r render-run-results, results='asis'}
for (run_id in names(run_results)) {

  r <- run_results[[run_id]]

  cat("\n\n# Sequence run: ", run_id, "\n\n", sep = "")

  if (r$status != "OK") {
    cat(
      "**Status:** ", r$status,
      ". No QC calculations were performed because the run input was missing or empty.\n\n",
      sep = ""
    )
    next
  }

  # -----------------------------
  # Preview
  # -----------------------------
  cat("## Preview data structure\n\n")
  cat(
    sprintf(
      "- Sample size = %s\n- All Locus = %s\n",
      r$preview$sample_size,
      r$preview$Locus_no
    )
  )

  if (!is_empty_df(r$preview$allele_count)) {
    cat(
      sprintf(
        "- Alleles per locus range = %s to %s\n",
        safe_min(r$preview$allele_count$n.alleles),
        safe_max(r$preview$allele_count$n.alleles)
      )
    )
  } else {
    cat("- Allele summary: no data\n")
  }

  if (!is_empty_df(r$preview$clone_wide)) {
    cat(
      sprintf(
        "- Mean total loci = %s\n- Mean proportion monoclonal = %s\n- Mean proportion polyclonal = %s\n",
        safe_mean(r$preview$clone_wide$Total_alleles),
        safe_mean(r$preview$clone_wide$Mono_prop),
        safe_mean(r$preview$clone_wide$Poly_prop)
      )
    )
  }

  cat("\n### Which loci are polyallelic?\n\n")
  p <- plot_total_alleles(r$data)
  if (is.null(p)) print(empty_plot("Which loci are polyallelic? — No data")) else print(ggplotly(p))

  # -----------------------------
  # Negative controls
  # -----------------------------
  cat("\n## Check negative controls\n\n")

  p <- plot_negative_read_histogram(r$negatives$negatives)
  if (is.null(p)) print(empty_plot("Negative controls — No negative controls found")) else print(p)

  p <- plot_negative_loci(r$negatives$loci_neg)
  if (is.null(p)) print(empty_plot("Negative controls by locus — No data")) else print(p)

  cat("\n**Negative controls with >100 reads/allele:**\n\n")
  print(r$negatives$bad_negatives)

  cat("\n**Loci with >=100 reads summed over negative controls:**\n\n")
  print(r$negatives$loci_neg_info)

  # -----------------------------
  # Positive controls
  # -----------------------------
  cat("\n## Check positive controls\n\n")

  p <- plot_pc_clones(r$data)
  if (is.null(p)) print(empty_plot("Positive control clonality — No positive controls found")) else print(ggplotly(p))

  p <- plot_pc_polyclonal_loci(r$data)
  if (is.null(p)) print(empty_plot("Positive control polyclonal loci — No data")) else print(ggplotly(p))

  cat("\n### Positive-control consistency\n\n")
  print(r$pc_consistency)

  p <- plot_pc_clones_filt(r$data)
  if (is.null(p)) print(empty_plot("Filtered positive-control clonality — No data")) else print(ggplotly(p))

  p <- plot_pc_polyclonal_loci_filt(r$data)
  if (is.null(p)) print(empty_plot("Filtered positive-control polyclonal loci — No data")) else print(ggplotly(p))

  # -----------------------------
  # Step 1
  # -----------------------------
  cat("\n## Step 1: Total Reads\n\n")
  
  if (is_empty_df(r$filtered_list1)) {
    cat("No samples were removed at this step.\n\n")
  } else {
    print(r$filtered_list1)
  }

  p <- plot_Reads_dist(r$data)
  if (is.null(p)) print(empty_plot("Total Reads distribution — No data")) else print(p)

  p <- plot_Reads_cutoff(
    r$data,
    loci = TOTAL_READS_LOCI,
    Reads_cutoff = TOTAL_READS_PER_LOCI
  )
  if (is.null(p)) print(empty_plot("Total Reads cutoff — No data")) else print(ggplotly(p))

  # cat("\n", r$filtered_list1, "\n", sep = "")

  # -----------------------------
  # Step 2
  # -----------------------------
  cat("\n## Step 2: Loci coverage\n\n")
  
  if (is_empty_df(r$filtered_list2)) {
    cat("No samples were removed at this step.\n\n")
  } else {
    print(r$filtered_list2)
  }

  p <- plot_loci_dist(r$filtered_data1)
  if (is.null(p)) print(empty_plot("Loci distribution — No data")) else print(p)

  p <- plot_loci_cutoff(
    r$filtered_data1,
    loci = TOTAL_READS_LOCI,
    perc_covered = LOCUS_COVERAGE_FRACTION
  )
  if (is.null(p)) print(empty_plot("Loci coverage cutoff — No data")) else print(ggplotly(p))

  # cat("\n", r$filtered_list2, "\n", sep = "")

  # -----------------------------
  # Step 3
  # -----------------------------
  cat("\n## Step 3: Proportion of loci > 100 Reads\n\n")
  
  if (is_empty_df(r$filtered_list3)) {
    cat("No samples were removed at this step.\n\n")
  } else {
    print(r$filtered_list3)
  }

  p <- plot_n100_loci_cutoff(
    r$filtered_data2,
    n100_cutoff = N100_CUTOFF
  )
  if (is.null(p)) print(empty_plot("Normalized proportion of loci >100 Reads — No data")) else print(p)

  p <- plot_n100_sample_cutoff(
    r$filtered_data2,
    n100_cutoff = N100_CUTOFF
  )
  if (is.null(p)) print(empty_plot("n100 proportion vs sample retention — No data")) else print(p)

  # cat("\n", r$filtered_list3, "\n", sep = "")

  # -----------------------------
  # Locus performance
  # -----------------------------
  cat("\n## Locus performance check\n\n")

  lp <- r$locus_check

  cat(
    sprintf(
      "%s loci had <= %s Reads at least in one sample.\n",
      safe_n_distinct(lp$gene_low_Reads$Locus),
      LOW_READ_LOCUS_CUTOFF
    )
  )

  cat(
    sprintf(
      "%s loci had <= %s Reads across more than 1/3 of total samples.\n",
      nrow(lp$gene_low_Reads_popcut),
      LOW_READ_LOCUS_CUTOFF
    )
  )

  cat("\n**Which loci had <=10 Reads across 1/3 of total samples?**\n\n")
  if (is_empty_df(lp$gene_low_Reads_popcut)) {
    cat("None.\n")
  } else {
    print(lp$gene_low_Reads_popcut)
  }

  # -----------------------------
  # Redos
  # -----------------------------
  cat("\n## Samples that need to be redone\n\n")
  print(r$redos)
}
```

# Export results

Each run gets its **own output folder**, so results from different sequence runs cannot overwrite each other.

```{r export-results}
for (run_id in names(run_results)) {

  r <- run_results[[run_id]]

  run_output <- file.path(output_dir, make.names(run_id))
  dir.create(run_output, recursive = TRUE, showWarnings = FALSE)

  if (r$status != "OK") {
    safe_write_csv(
      data.frame(
        RunID = run_id,
        RunDir = r$RunDir,
        Status = r$status
      ),
      file.path(run_output, "run_status.csv")
    )
    next
  }

  # Run summary.
  safe_write_csv(
    data.frame(
      RunID = run_id,
      Samples_input = safe_n_distinct(r$data$SampleID),
      Loci_input = safe_n_distinct(r$data$Locus),
      Samples_after_step1 = safe_n_distinct(r$filtered_data1$SampleID),
      Samples_after_step2 = safe_n_distinct(r$filtered_data2$SampleID),
      Samples_after_step3 = safe_n_distinct(r$filtered_data3$SampleID),
      Samples_redo = safe_n_distinct(r$redos$SampleID)
    ),
    file.path(run_output, "run_summary.csv")
  )

  # Control results.
  safe_write_csv(
    r$negatives$bad_negatives,
    file.path(run_output, "negative_controls_over_100_reads.csv")
  )

  safe_write_csv(
    r$negatives$loci_neg_info,
    file.path(run_output, "negative_control_loci_over_100_reads.csv")
  )

  safe_write_csv(
    r$pc_consistency,
    file.path(run_output, "positive_control_consistency.csv")
  )

  # Redo lists.
  safe_write_csv(
    r$redos,
    file.path(run_output, "QC_failed_samples.csv")
  )

  # Filtered allele table.
  safe_write_csv(
    r$filtered_data3,
    file.path(run_output, "TES_filtered_allele_data.csv")
  )

  # Locus performance.
  safe_write_csv(
    r$locus_check$gene_low_Reads_popcut,
    file.path(run_output, "loci_low_reads_over_1_3_samples.csv")
  )
}

# Combined overview across all runs.
safe_write_csv(
  run_summary,
  file.path(output_dir, "all_runs_summary.csv")
)

run_summary
```


# Merge Allele Data for Re-sequenced Samples, Remove False Positive Alleles and Prepare an Input for Moire

allele frequency visualization 

```{r visualization-functions}
#| label: visualization-functions

plot_allele_distribution <- function(dat) {
  n_alleles_per_locus_by_sample <- dat |>
    dplyr::group_by(sample_id, locus) |>
    dplyr::summarize(n_alleles = dplyr::n_distinct(allele)) |>
    dplyr::ungroup()

  ggplot(n_alleles_per_locus_by_sample, aes(x = sample_id, group = n_alleles, fill = as.factor(n_alleles))) +
    geom_bar(position = "stack") +
    theme(axis.text.x = element_text(angle = 90)) +
    scale_x_discrete(limits = n_alleles_per_locus_by_sample |>
      dplyr::group_by(sample_id) |>
      dplyr::summarize(avg_n_alleles = mean(n_alleles)) |>
      dplyr::arrange(avg_n_alleles) |>
      dplyr::pull(sample_id)) +
    xlab("Sample") +
    ylab("Number of Alleles")
}

plot_allele_prev_metrics <- function(dat, use_plotly = TRUE) {
  total_samples <- dat |>
    dplyr::distinct(sample_id) |>
    nrow()

  allele_prev_metrics <- dat |>
    dplyr::group_by(sample_id, locus) |>
    dplyr::mutate(prop = reads / sum(reads)) |>
    dplyr::ungroup() |>
    dplyr::group_by(locus, allele) |>
    dplyr::summarize(
      within_host_avg_prop = mean(prop),
      pop_prev = dplyr::n_distinct(sample_id) / total_samples,
      n_samples = dplyr::n_distinct(sample_id)
    )

  g <- ggplot(
      allele_prev_metrics, 
      aes(x = pop_prev, y = within_host_avg_prop, color = locus)
    ) +
    geom_point() +
    guides(color = "none") +
    xlab("Population Prevalence") +
    ylab("Within-Host Average Proportion") +
    theme_minimal()
  g
}
```


# merge the filtered allele data from each run

```{r import, echo = F, warning = F, message= F}
# looking at the raw alelle data and identifing potential alleles and loci that could prove difficult when merging and filter those out 

run4 <- read_csv("../data/QC/Results/run4/TES_filtered_allele_data.csv") %>%
  filter(SampleID != "BA056-D0", Locus != "Pf3D7_03_v3-891553-891827-1A", PseudoCIGAR != "93C116G177C") %>%
  filter(SampleID != "MM024-D28", Locus != "Pf3D7_14_v3-1166107-1166371-1A", PseudoCIGAR != "6A28+9N64+8N88+9N100+8N179A") %>%
  select(-norm.reads.locus, -n.alleles, -Allele)


run5 <- read_csv("../data/QC/Results/run5/TES_filtered_allele_data.csv") %>%
  filter(Locus != "Pf3D7_13_v3-1465616-1465890-1A") %>%
  select(-norm.reads.locus, -n.alleles, -Allele)


run9 <- read_csv("../data/QC/Results/run9/TES_filtered_allele_data.csv") %>%
  filter(Locus != "Pf3D7_13_v3-1465616-1465890-1A") %>%
  select(-norm.reads.locus, -n.alleles, -Allele)


data <- rbind(run4, run5, run9)


# Remove suffix not part of the sample name
data$SampleID <- sub("_S[0-9]+", "",data$SampleID)

pseudo_cigar_to_alleles <- function(df) {
  df %>% select(Locus, PseudoCIGAR, ASV) %>% 
    distinct() %>% 
    group_by(Locus) %>% 
    mutate(Allele = seq(1, n())) %>% 
    ungroup() %>% 
    mutate(Allele = paste0(Locus, ".", Allele))
}


# generate allele from pseudocigar
allele_sequence <- pseudo_cigar_to_alleles(data)


data %<>% left_join(allele_sequence, by = c("PseudoCIGAR", "Locus", "ASV")) %>%
  group_by(SampleID, Locus, Allele) %>%
  mutate(Read = sum(Reads)) %>%
  filter(Read > 5) %>%  # filter alleles with less than 5 reads
  ungroup() %>%
  select(-Reads) %>%
  dplyr::rename(Reads = Read) %>%
  distinct() %>%
  filter(!grepl("3D7*", SampleID))


# Check for duplicates
any(duplicated(data)) # we good

print("Pre-filtering")
# Overall range of the data
sample_size <- n_distinct(data$SampleID)
print(sprintf("Sample size= %s", sample_size))

Locus_no <- n_distinct(data$Locus)
print(sprintf("All Locus = %s", Locus_no))


(plot_allele_prev_metrics(data |> dplyr::rename(sample_id = SampleID, locus = Locus, allele = Allele, reads = Reads)) +
    labs(title = "Pre-filtering"))


# Use only diversity markers
pools <- pool_1A

data <- data %>%
  left_join(pools, by = "Locus")

data_div <- data %>%
  filter(Category.x == "Diversity") %>% 
  mutate(site = case_when(
    grepl("AA", SampleID) ~ "Asayita",
    grepl("AC", SampleID) ~ "Gambella",
    grepl("BA", SampleID) ~ "Benshangul",
    grepl("GM", SampleID) ~ "Gondar",
    grepl("MM", SampleID) ~ "Mizan",
    TRUE ~ "NA"
  ))

print("Post-filtering")
print(sprintf("Sample size= %s", sample_size)) # 615 final
Locus_no <- n_distinct(data_div$Locus)
print(sprintf("All Locus = %s", Locus_no)) # 165 final


# loci with many false positive alleles
loci_to_remove <- read_csv("../data/QC/Results/loci_to_remove.csv")

data_div %<>%
  filter(!Locus %in% loci_to_remove$Locus)


(plot_allele_prev_metrics(data_div |> dplyr::rename(sample_id = SampleID, locus = Locus, allele = Allele, reads = Reads)) +
    labs(title = "Post-filtering"))




```

### Preview raw data

What is the allele distribution for polyclonal loci?

```{r preview, echo=FALSE, warning = FALSE, message= FALSE}

data_new <- data_div %>% 
  dplyr::rename(sample_id = SampleID)

# Count number of alleles per Locus per sample
allele_count <- data_new %>% 
    group_by(sample_id, Locus) %>%
  summarise(n.alleles = n()) %>%
  group_by(Locus) %>% 
  arrange(-n.alleles)
# Allele range in each sample
print(sprintf("%s to %s alleles per Locus per sample",
              min(allele_count$n.alleles), max(allele_count$n.alleles)))

n_allele <- data_new %>%
  distinct(sample_id, Locus, Allele, .keep_all = TRUE) %>%
  group_by(sample_id, Locus) %>%
  reframe(n.alleles = n(), norm.reads.locus = Reads/sum(Reads))

# allele stats
allele_stats <- n_allele %>%
  group_by(Locus) %>%
  summarise(mean = mean(n.alleles), median = median(n.alleles),
            min = min(n.alleles), max = max(n.alleles)) %>%
  arrange(-max)
print(head(allele_stats))

# Define categories for number of alleles
n_allele_poly <- n_allele %>%
  filter(n.alleles >1) %>%
  mutate(allele_category = case_when(
    n.alleles == 2 ~ "2",
    n.alleles == 3 ~ "3",
    n.alleles %in% 4:5 ~ "4-5",
    n.alleles %in% 6:8 ~ "6-8",
    n.alleles > 8 ~ "9+"
  ))
# Calculate proportion of samples with each allele category per locus
n_allele_plot <- n_allele_poly %>%
  group_by(Locus, allele_category) %>%
  summarise(count = n()) %>%
  group_by(Locus) %>%
  mutate(prop.allele = count / sum(count))

# Reorder the allele_category factor levels
n_allele_plot <- n_allele_plot %>%
  mutate(allele_category = factor(allele_category, levels = c("2", "3", "4-5", "6-8", "9+")))

# Set order to proportion of 2 alleles
level_2allele <- n_allele_plot %>%
    filter(allele_category == "2") %>%
    arrange(-prop.allele) %>%
    pull(Locus)
n_allele_plot <- n_allele_plot %>%
  mutate(Locus = factor(Locus, levels = level_2allele))

# Plotting the stacked bar plot
g1 <- ggplot(n_allele_plot, aes(x = factor(Locus, level = level_2allele), y = prop.allele, fill = allele_category)) +
  geom_bar(stat = "identity") +
  labs(x = "Locus", y = "Proportion", fill = "Number of Alleles") +
  theme_classic() +
  theme(axis.text.x = element_blank()) 
g1p <- ggplotly(g1)
g1p


```

### Filter minor alleles by 1% prevalence

How does the allele distribution change after filtering?

```{r filter, echo=FALSE, warning = FALSE, message= FALSE}

##### Population level filtering first ###########

# Filter alleles with less than 1% of total reads at the population level
filtered_alleles <- data_new %>%
  group_by(sample_id, site, Locus, Allele) %>%
  summarise(
    total_reads = sum(Reads), # Total reads per allele across samples
    locus_reads = sum(Reads, na.rm = TRUE) %>% first(), # Total reads in the locus per sample/site
    .groups = "drop"
  ) %>%
  mutate(population_frequency = total_reads / locus_reads * 100) %>%
  filter(population_frequency >= 2) %>%
  select(sample_id, site, Locus, Allele)

# Join back with the original dataframe to retain only filtered alleles
filtered_data <- data_new %>%
  inner_join(filtered_alleles, by = c("sample_id", "site", "Locus", "Allele"))

###### Now do straight 1% host-level allele frequency ###########

filtered_data %<>%
  distinct(sample_id, Locus, Allele, .keep_all = TRUE) %>%
  group_by(sample_id, Locus) %>%
  mutate(norm.reads.locus = Reads/sum(Reads))

ggplot(data = filtered_data, aes(x = norm.reads.locus, y = Reads)) +
  geom_point(aes(color = site)) +
  scale_y_log10() +
  facet_wrap(~site) +
  ggtitle("Allele read count vs. within-host frequency") +
  geom_vline(xintercept = 0.01)


# Filter by norm.reads.locus > 0.01
filtered_allele_data <- filtered_data %>%
  filter(norm.reads.locus > 0.02)
  # filter(norm.reads.locus > 0.10 & Reads > 10) # apply another, if <10% AND read count > 10 (1 out of 10 reads)

# Now how does it look now
ggplot(data = filtered_allele_data, aes(x = norm.reads.locus, y = Reads)) +
  geom_point(aes(color = site)) +
  scale_y_log10() +
  facet_wrap(~site) +
  ggtitle("Allele read count vs. within-host frequency")

filtered <- filtered_allele_data %>%
  group_by(sample_id) %>%
  summarise(total = sum(Reads))

# write_csv(filtered_allele_data, "TES_final_allele_data.csv")

sample_size <- n_distinct(filtered_allele_data$sample_id)
print(sprintf("Sample size= %s", sample_size))

# new allele stats
allele_count_filt <- filtered_allele_data %>%
  group_by(sample_id, Locus) %>%
  mutate(n.alleles = n())

# Allele range in each sample
print(sprintf("%s to %s alleles per Locus per sample",
              min(allele_count_filt$n.alleles), max(allele_count_filt$n.alleles)))

# new allele stats
allele_stats_filt <- filtered_allele_data %>%
  group_by(sample_id, Locus) %>%
  mutate(n.alleles = n()) %>%
  group_by(Locus) %>%
  summarize(mean = mean(n.alleles), median = median(n.alleles),
            min = min(n.alleles), max = max(n.alleles)) %>%
  arrange(-max)
print(head(allele_stats_filt))

# Define categories for number of alleles
allele_filt_poly <- allele_count_filt %>%
  filter(n.alleles >1) %>%
  mutate(allele_category = case_when(
    n.alleles == 2 ~ "2",
    n.alleles == 3 ~ "3",
    n.alleles %in% 4:5 ~ "4-5",
    n.alleles %in% 6:8 ~ "6-8",
    n.alleles > 8 ~ "9+")) %>%
  group_by(Locus, allele_category) %>%
  summarize(count = n()) %>%
  group_by(Locus) %>%
  mutate(prop.allele = count / sum(count))
# Reorder the allele_category factor levels
allele_filt_poly <- allele_filt_poly %>%
  mutate(allele_category = factor(allele_category, levels = c("2", "3", "4-5", "6-8", "9+")))
# Set order to proportion of 2 alleles
level_2allele <- allele_filt_poly %>%
    filter(allele_category == "2") %>%
    arrange(-prop.allele) %>%
    pull(Locus)
allele_filt_poly <- allele_filt_poly %>%
  mutate(Locus = factor(Locus, levels = level_2allele))
# Plotting the stacked bar plot
g2 <- ggplot(allele_filt_poly, aes(x = factor(Locus, level = level_2allele), y = prop.allele, fill = allele_category)) +
  geom_bar(stat = "identity") +
  labs(x = "Locus", y = "Proportion", fill = "Number of Alleles") +
  theme_classic() +
  theme(axis.text.x = element_blank()) 
g2p <- ggplotly(g2)
g2p


```

Looks like no loci have 9+ alleles after filtering. The max number of alleles is now 8. What are the loci with 6-8 alleles?

```{r investigate, echo=FALSE, warning = FALSE, message= FALSE}

loci_lot.al <- allele_filt_poly %>%
  filter(allele_category == "6-8") %>%
  pull(Locus)
print(loci_lot.al)

```

### Make moire input

```{r moire_input, echo=FALSE, warning = FALSE, message= FALSE}

# Make moire input
moire_input <- filtered_allele_data %>% 
  select (sample_id, Locus, Allele, site) %>%
  rename("locus" = "Locus") %>%
  rename("allele" = "Allele") %>% 
  distinct()


moire_input_aa <- moire_input %>% 
  filter(site == "Asayita")

moire_input_ac <- moire_input %>% 
  filter(site == "Gambella")

moire_input_ba <- moire_input %>% 
  filter(site == "Benshangul")

moire_input_gm <- moire_input %>% 
  filter(site == "Gondar")

moire_input_mm <- moire_input %>% 
  filter(site == "Mizan")

write.csv(moire_input, "../data/Moire/Input/TES_moire_input_filtered.csv", row.names = F)
write.csv(moire_input_aa, "../data/Moire/Input/TES_moire_input_filtered_aa.csv", row.names = F)
write.csv(moire_input_ac, "../data/Moire/Input/TES_moire_input_filtered_ac.csv", row.names = F)
write.csv(moire_input_ba, "../data/Moire/Input/TES_moire_input_filtered_ba.csv", row.names = F)
write.csv(moire_input_gm, "../data/Moire/Input/TES_moire_input_filtered_gm.csv", row.names = F)
write.csv(moire_input_mm, "../data/Moire/Input/TES_moire_input_filtered_mm.csv", row.names = F)


```
