---
title: "Moire diagnostics, Dcifer & P-adjustments"
author: "William Louie and Yisihak Abreham"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

## Load Libraries

```{r load libraries, warning=FALSE, message=FALSE}

library(tidyverse)
library(readr)
library(moire)
library(plotly)
library(magrittr)
library(gtsummary)
library(gt)
library(ggpubr)
library(webshot2)
library(grid)
library(patchwork)
library(dcifer)
library(ggnewscale)
library(igraph)

```

## Import data
Import & manage pre- and post-moire data

```{r import, echo=FALSE, message=FALSE, warning=FALSE}

moire_input <- read_csv("../data/Moire/Input/TES_moire_input_filtered.csv")

# Split into sitess
moire_input_aa <- moire_input %>%
  filter(site == "Asayita")
moire_input_ac <- moire_input %>%
  filter(site == "Gambella")
moire_input_ba <- moire_input %>%
  filter(site == "Benshangul")
moire_input_gm <- moire_input %>%
  filter(site == "Gondar")
moire_input_mm <- moire_input %>%
  filter(site == "Mizan")

# Load long form
dat <- moire::load_long_form_data(moire_input)
dat_aa <- moire::load_long_form_data(moire_input_aa)
dat_ac <- moire::load_long_form_data(moire_input_ac)
dat_ba <- moire::load_long_form_data(moire_input_ba)
dat_gm <- moire::load_long_form_data(moire_input_gm)
dat_mm <- moire::load_long_form_data(moire_input_mm)

# Import moire results
moire_results_aa <- readRDS("../data/Moire/Results/TES_moire_result_updated_aa.rds")
moire_results_ac <- readRDS("../data/Moire/Results/TES_moire_result_updated_ac.rds")
moire_results_ba <- readRDS("../data/Moire/Results/TES_moire_result_updated_ba.rds")
moire_results_gm <- readRDS("../data/Moire/Results/TES_moire_result_updated_gm.rds")
moire_results_mm <- readRDS("../data/Moire/Results/TES_moire_result_updated_mm.rds")

# Filter uninformative loci
filtered_dat <- moire_input %>% filter(! locus %in% dat$uninformative_loci)

filtered_dat_aa <- moire_input_aa %>% filter(! locus %in% dat_aa$uninformative_loci)
setdiff(moire_results_aa[["args"]][["data"]][["loci"]], unique(filtered_dat_aa$locus))

filtered_dat_ac <- moire_input_ac %>% filter(! locus %in% dat_ac$uninformative_loci)
setdiff(moire_results_ac[["args"]][["data"]][["loci"]], unique(filtered_dat_ac$locus))

filtered_dat_ba <- moire_input_ba %>% filter(! locus %in% dat_ba$uninformative_loci)
setdiff(moire_results_ba[["args"]][["data"]][["loci"]], unique(filtered_dat_ba$locus))

filtered_dat_gm <- moire_input_gm %>% filter(! locus %in% dat_gm$uninformative_loci)
setdiff(moire_results_gm[["args"]][["data"]][["loci"]], unique(filtered_dat_gm$locus))

filtered_dat_mm <- moire_input_mm %>% filter(! locus %in% dat_mm$uninformative_loci)
setdiff(moire_results_mm[["args"]][["data"]][["loci"]], unique(filtered_dat_mm$locus))

```



# Post-moire exploration

**MCMC diagnostics**
```{r moire chain diagnostics, echo=FALSE, warning = FALSE, message= FALSE}

moire_results <- readr::read_rds("../data/Moire/Results/TES_moire_result_updated.rds")

moire_data <- read.csv("../data/Moire/Input/TES_moire_input_filtered.csv", header = T)

dat <- moire::load_long_form_data(moire_data)

# Convergence
plot_posterior_llik <- function(moire_results) {
  # does not currently support multiple chains
  posterior_burnin <- moire_results$chains[[1]]$posterior_burnin
  posterior_sample <- moire_results$chains[[1]]$posterior_sample
  prior_burnin <- moire_results$chains[[1]]$prior_burnin
  prior_sample <- moire_results$chains[[1]]$prior_sample
  likelihood_burnin <- moire_results$chains[[1]]$llik_burnin
  likelihood_sample <- moire_results$chains[[1]]$llik_sample
  total = length(posterior_burnin) + length(posterior_sample)
  dat <- rbind(
    data.frame(llik = "posterior", value = c(posterior_burnin, posterior_sample), sample = 1:total),
    data.frame(llik = "prior", value = c(prior_burnin, prior_sample), sample = 1:total),
    data.frame(llik = "likelihood", value = c(likelihood_burnin, likelihood_sample), sample = 1:total)
  )
  g <- ggplot(dat, aes(x = sample, y = value, color = llik, group = llik)) +
    geom_line() +
    geom_vline(xintercept = length(posterior_burnin), linetype = "dashed", alpha = .5)
  g
}

ggplotly(plot_posterior_llik(moire_results))

# Chain swap
plot_chain_swap_dist <- function(moire_results) {
  swap_dist <- moire_results$chains[[1]]$swap_acceptances / moire_results$args$samples_per_chain
  temps <- moire_results$chains[[1]]$temp_gradient
  swap_idx <- (temps[1:length(temps)-1] + temps[2:length(temps)]) / 2
  dat <- data.frame(swap_rate = swap_dist, temp = swap_idx)
  g <- ggplot(dat, aes(x = temp, y = swap_rate)) +
    geom_point() +
    geom_vline(data = data.frame(x = temps), aes(xintercept=x), linetype = "dashed", alpha = 0.25) +
    coord_cartesian(ylim=c(0,1))
  g
}
ggplotly(plot_chain_swap_dist(moire_results))

#Check eps_pos/eps_neg distributions
plot_eps_pos <- function(moire_results) {
  eps_pos <- moire::summarize_epsilon_pos(moire_results)

  g <- ggplot(eps_pos, aes(x = sample_id, y = post_eps_pos_med,
                           ymin = post_eps_pos_lower,
                           ymax = post_eps_pos_upper)) +
    geom_errorbar() +
    geom_point() +
    theme(axis.text.x = element_blank())

  g
}

plot_eps_neg <- function(moire_results) {
  eps_neg <- moire::summarize_epsilon_neg(moire_results)

  g <- ggplot(eps_neg, aes(x = sample_id, y = post_eps_neg_med,
                           ymin = post_eps_neg_lower,
                           ymax = post_eps_neg_upper)) +
    geom_errorbar() +
    geom_point() +
    theme(axis.text.x = element_blank())

  g
}

ggplotly(plot_eps_pos(moire_results))
ggplotly(plot_eps_neg(moire_results))

```

## Summarize COI

```{r moire COI_eCOI, echo=FALSE, warning = FALSE, message= FALSE} 

# Summarize COI for each site
coi_aa <- moire::summarize_coi(moire_results_aa)
coi_ac <- moire::summarize_coi(moire_results_ac)
coi_ba <- moire::summarize_coi(moire_results_ba)
coi_gm <- moire::summarize_coi(moire_results_gm)
coi_mm <- moire::summarize_coi(moire_results_mm)
coi_summary <- rbind(coi_aa, coi_ac, coi_ba, coi_gm, coi_mm)

coi_summary <- coi_summary %<>%
    mutate(site = case_when(
    grepl(pattern = "AA", sample_id) ~ "Asayita",
    grepl(pattern = "AC", sample_id) ~ "Gambella",
    grepl(pattern = "GM", sample_id) ~ "Gondar",
    grepl(pattern = "BA", sample_id) ~ "Assossa",
    grepl(pattern = "MM", sample_id) ~ "Mizan")) 


# Summarize eCOI for each site
ecoi_aa <- moire::summarize_effective_coi(moire_results_aa)
ecoi_ac <- moire::summarize_effective_coi(moire_results_ac)
ecoi_ba <- moire::summarize_effective_coi(moire_results_ba)
ecoi_gm <- moire::summarize_effective_coi(moire_results_gm)
ecoi_mm <- moire::summarize_effective_coi(moire_results_mm)
ecoi_summary <- rbind(ecoi_aa, ecoi_ac, ecoi_ba, ecoi_gm, ecoi_mm)
rm(ecoi_aa, ecoi_ac, ecoi_arc, ecoi_ba, ecoi_gm, ecoi_mm)

coi_summary <- coi_summary %>%
  full_join(ecoi_summary, by = "sample_id")

ecoi_summary %<>%
  mutate(site = case_when(
    grepl(pattern = "AA", sample_id) ~ "Asayita",
    grepl(pattern = "AC", sample_id) ~ "Gambella",
    grepl(pattern = "GM", sample_id) ~ "Gondar",
    grepl(pattern = "BA", sample_id) ~ "Assossa",
    grepl(pattern = "MM", sample_id) ~ "Mizan")) 

write_csv(coi_summary, "../data/Moire/Results/coi_summary.csv")

print("What is the distribution of median COI?")
# Plot individual coi median
ggplot(coi_summary, aes(x = sample_id, y = post_coi_med,
                           ymin = post_coi_lower,
                           ymax = post_coi_upper)) +
    geom_errorbar() +
    geom_point() +
    theme_classic(base_size = 10) +
    theme(axis.text.x = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank()) +
    facet_wrap(~site, scales = "free") +
    ggtitle("Median COI, by site")

print("What is the distribution of mean COI?")
# Plot individual coi mean
ggplot(coi_summary, aes(x=post_coi_mean,
                            ymin = post_coi_lower,
                            ymax = post_coi_upper)) + 
  geom_histogram(bins = 20) +
  scale_x_continuous(breaks=seq(1,30,1)) +
  geom_vline(aes(xintercept = mean(post_coi_mean)), color = "gray", linetype ='dashed') +
  xlab("Mean COI") +
  ylab ("Count") +
  ggtitle("Distribution of Mean COI") +
  theme_classic() +
  facet_wrap(~site, scales = "free") +
  ggtitle("Mean COI, by site")


print("Plot distribution of median effective COI")
# Plot individual effective coi median
ggplot(coi_summary, aes(x = sample_id, y = post_effective_coi_med,
                           ymin = post_effective_coi_lower,
                           ymax = post_effective_coi_upper)) +
    geom_errorbar() +
    geom_point() +
    theme_classic(base_size = 10) +
    theme(axis.text.x = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank()) +
    facet_wrap(~site, scales = "free") +
    ggtitle("Median effective COI, by site")

print("Plot distribution of mean effective COI")
# Plot individual effective coi mean
ggplot(ecoi_summary, aes(x=post_effective_coi_mean)) + 
  geom_histogram(bins = 20) +
  scale_x_continuous(breaks=seq(1,15,1)) +
  geom_vline(aes(xintercept = mean(post_effective_coi_mean)), color = "gray", linetype = "dashed") +
  xlab("Mean Effective COI") + 
  ylab("Count") +
  ggtitle("Distribution of Mean Effective COI") +
  theme_classic() +
  facet_wrap(~site, scales = "free") +
  ggtitle("Mean effective COI, by site")


```


### COI dot plot

```{r}

## COI plot of Day 0s only
coi_summary_d0 <- coi_summary %>%
  filter(grepl("D0", sample_id))


coi_arrange <- coi_summary_d0 %>%
  mutate(mean_COI = as.numeric(post_coi_mean)) %>%
  mutate(median_COI = as.numeric(post_coi_med)) %>%
  mutate(mean_eCOI = as.numeric(post_effective_coi_mean)) %>%
  mutate(median_eCOI = as.numeric(post_effective_coi_med)) %>%
  select(sample_id, mean_COI, median_COI, mean_eCOI, median_eCOI, site) %>%
  mutate(median_COI = paste0(median_COI, ".00")) %>%
  mutate(median_COI = as.numeric(median_COI)) 



# COI table summary
table_coi <- coi_arrange |>
  tbl_summary(by = site, 
    include = c(mean_COI, mean_eCOI),
    statistic = list(
      mean_COI ~ "{mean} ({sd})",
      mean_eCOI ~ "{mean} ({sd})")) %>%
  add_p()
table_coi


# Save as gt
## We will combine this with the plot later
gt_table <- table_coi %>%
  as_gt()


# Site colors
site_color <- c("Gondar" = "blue","Asayita" = "darkred","Assossa" = "darkorange", "Mizan" = "darkviolet","Gambella" = "yellow3" )

# Site order
site_order <- c("Asayita", "Assossa", "Gambella", "Gondar", "Mizan")
coi_summary_d0$site <- factor(coi_summary_d0$site, levels = site_order)

# Calc the standard deviations
coi_summary_d0_sd <- coi_summary_d0 %>%
  group_by(site) %>%
  dplyr::summarize(
    mean = mean(post_coi_mean),
    sd = sd(post_coi_mean),
    ymin = mean - sd,
    ymax = mean + sd,
    .groups = 'drop') %>%
  mutate(site = factor(site, levels = site_order))

# Use coi_summary
g_coi <- ggplot(coi_summary_d0, 
  aes(x = site, y = post_coi_mean)) +
  geom_point(aes(color = site), position = "jitter") +
  geom_boxplot(aes(color = site, fill = site), alpha = 0.1, coef = 0, outlier.shape = NA) +
  geom_errorbar(data = coi_summary_d0_sd, aes(y = mean, ymin = ymin, ymax = ymax, x = site), width = 0.5,
                color = "black") +
  geom_segment(data = coi_summary_d0_sd, 
               aes(x = as.numeric(site) - 0.25, 
                   xend = as.numeric(site) + 0.25, 
                   y = mean, yend = mean), 
               inherit.aes = FALSE, color = "black", size = 1) +
  theme_minimal() +
  scale_y_continuous(breaks = seq(1, 10, by = 1)) +
  ylab("posterior mean COI") +
  scale_color_manual(values = site_color) +
  scale_fill_manual(values = site_color) +
  theme(axis.title.x = element_blank(),
        axis.text = element_text(size = 14),
        legend.text = element_text(size = 16),
      panel.grid = element_blank(), 
      axis.line = element_line(size = 1.1))
g_coi

ggplotly(g_coi)


# Do the stats
comparisons <- compare_means(post_coi_mean ~ site, data = coi_summary_d0, method = "wilcox.test")
# Add a y.position column to set where the significance bars should appear
y_max <- max(coi_summary_d0$post_coi_mean)
spacing <- (y_max * 0.02)  # Space lines at 2% of the maximum y value
comparisons <- comparisons %>%
  mutate(y.position = y_max + (seq(0.5, nrow(comparisons)) - 1) * spacing)

g_coi_stat <- g_coi + stat_pvalue_manual(
  comparisons,
  method = "wilcox.test",
  label = "p.signif", hide.ns = TRUE,
  size = 6)
g_coi_stat


Sys.setenv(
  CHROMOTE_CHROME = "C:/Program Files/Google/Chrome/Application/chrome.exe"
)


# Save the gt table as an image
gtsave(gt_table, filename = "Figure_d0_coi_distribution.png", path = "../data/Moire/Results/Figures/")
# # Read the saved table image as a grob (graphical object)
table_grob <- grid::rasterGrob(png::readPNG(file.path("../data/Moire/Results/Figures/Figure_d0_coi_distribution.png")), interpolate = TRUE)
# # Combine the plot and table using patchwork
combined_plot <- g_coi_stat + table_grob + plot_layout(ncol = 1, heights = c(2, 1))
# # Final save
ggsave("../data/Moire/Results/Figures/Figure_d0_coi_distribution.pdf", plot = combined_plot, width = 9, height = 11, dpi = 300, device = "pdf")

```


## Dcifer - Asayita
Heatmap shows relatedness between D0s only

```{r dcifer_Asayita, echo=FALSE}

mydsmp <- dcifer::formatDat(filtered_dat_aa, "sample_id", "locus", "allele")
coi   <- coi_aa$post_coi_med
afreq_aa <- dcifer::calcAfreq(mydsmp, coi)

# Match alleles
dsmp <- dcifer::matchAfreq(mydsmp, afreq_aa)

# Run ibdDat
dres_aa <- ibdDat(dsmp$dsmp, coi, dsmp$afreq,
                  pval = TRUE, confint = TRUE,
                  rnull = 0.125, alpha = 0.01, nr = 1e3)

## Output dcifer results
readr::write_rds(dres_aa, "../data/Dcifer/Results/dcifer_result_aa.rds")

# === Relatedness matrix ===
dmat <- dres_aa[, , "estimate"]
dmat[upper.tri(dmat)] <- t(dmat)[upper.tri(t(dmat))]

# === Adjust p-values ===
dres0 <- dres_aa[, , "p_value"]
pval_vec <- dres0[lower.tri(dres0)]                           # raw p-values
pval_adj <- p.adjust(pval_vec, method = "BH")                 # [MODIFIED]: BH-adjustment
adjusted_pmat <- matrix(NA, nrow = nrow(dres0), ncol = ncol(dres0),
                        dimnames = dimnames(dres0))           # [MODIFIED]
adjusted_pmat[lower.tri(adjusted_pmat)] <- pval_adj          # [MODIFIED]
adjusted_pmat[upper.tri(adjusted_pmat)] <- t(adjusted_pmat)[upper.tri(adjusted_pmat)]  # symmetric

# === Categorize adjusted p-values ===
dres_low <- -log10(adjusted_pmat)                            # [MODIFIED]
dres_low[dres_low > -log10(0.0001)] <- "P << 0.001"
dres_low[dres_low <= -log10(0.0001) & dres_low > -log10(0.001)] <- "P < 0.001"
dres_low[dres_low <= -log10(0.001) & dres_low > -log10(0.01)] <- "P < 0.01"
dres_low[dres_low <= -log10(0.01)] <- "Not Sig"
dres_low[upper.tri(dres_low)] <- t(dres_low)[upper.tri(dres_low)]

# === Matrix to long format ===
matrix_to_longformat <- function(mat) {
  as.data.frame(mat) |>
    rownames_to_column("sample.x") |>
    pivot_longer(-sample.x, names_to = "sample.y", values_to = "value")
}

to_plot1 <- matrix_to_longformat(dmat)

# === Clustering for ordering ===
hc <- hclust(as.dist(1 - dmat))
order <- hc$order
clustered_levels <- colnames(dmat)[order]

to_plot1 <- to_plot1 |>
  mutate(sample.x = fct_relevel(sample.x, clustered_levels),
         sample.y = fct_relevel(sample.y, clustered_levels))

to_plot2 <- matrix_to_longformat(dres_low)
to_plot2 <- to_plot2 |>
  mutate(sample.x = fct_relevel(sample.x, clustered_levels),
         sample.y = fct_relevel(sample.y, clustered_levels))

# === Define colors ===
colours_upper <- colorRampPalette(c("#FFFFFF", "#1520A6"))(100)
colours_lower <- c("P << 0.001" = "#C70039", "P < 0.001" = "#FF5733", "P < 0.01" = "#FFC300", "Not Sig" = "white")

# === Filter for D0 samples ===
to_plot1_0 <- to_plot1[grepl("D0", to_plot1$sample.x) & grepl("D0", to_plot1$sample.y), ]
to_plot2_0 <- to_plot2[grepl("D0", to_plot2$sample.x) & grepl("D0", to_plot2$sample.y), ]

sample_d0 <- unique(to_plot1_0$sample.x)
cluster_0 <- clustered_levels[clustered_levels %in% sample_d0]

to_plot1_0 <- to_plot1_0 |>
  mutate(sample.x = fct_relevel(sample.x, cluster_0),
         sample.y = fct_relevel(sample.y, cluster_0))

to_plot2_0 <- to_plot2_0 |>
  mutate(sample.x = fct_relevel(sample.x, cluster_0),
         sample.y = fct_relevel(sample.y, cluster_0))

pdf("../data/Dcifer/Results/figures/Asayita_relatedness_pvalue_corrected.pdf", width = 10, height = 10)

# === Heatmap ===
g0_aa <- ggplot(to_plot1_0, aes(x = sample.x, y = sample.y)) +
  geom_tile(data = subset(to_plot1_0, as.numeric(sample.x) > as.numeric(sample.y)), aes(fill = value)) +
  scale_fill_gradientn(colors = colours_upper, name = "Relatedness") +
  theme(axis.text.x = element_blank()) +
  theme(axis.text.y = element_blank()) +
  new_scale_fill() +
  geom_tile(data = subset(to_plot2_0, as.numeric(sample.x) < as.numeric(sample.y)),
            aes(fill = value)) +
  scale_fill_manual(values = colours_lower, name = "Significance", 
                    breaks = c("P << 0.001", "P < 0.001", "P < 0.01")) +
  scale_x_discrete(limits = cluster_0) +
  scale_y_discrete(limits = cluster_0) +
  ggtitle("Asayita (BH adjusted pvalue)")

g0_aa

dev.off()




# Step 1: Get all related pairs (IBD > 0)
related_pairs <- to_plot1_0 %>% filter(value > 0.5)

# Step 2: Get all unique samples involved in related pairs
related_samples <- unique(c(related_pairs$sample.x, related_pairs$sample.y))

# Step 3: Get all unique samples in the full dataset
all_samples <- unique(c(to_plot1_0$sample.x, to_plot1_0$sample.y))

# Step 4: Compute proportion
proportion_related <- length(related_samples) / length(all_samples)

print("proportion_related")
# Show result
proportion_related





# Extract IBD estimates from rnull > 0.125 run
ibd_est_rp5 <- dres_aa[, , "estimate"]
ibd_lower_rp5 <- dres_aa[, , "CI_lower"]
ibd_upper_rp5 <- dres_aa[, , "CI_upper"]

ibd_df_rp5_df  <- as.data.frame(ibd_est_rp5)
ibd_df_rp5_df  <- tibble::rownames_to_column(ibd_df_rp5_df, var = "sample.x")

ibd_df_rp5_lower_df  <- as.data.frame(ibd_lower_rp5)
ibd_df_rp5_lower_df  <- tibble::rownames_to_column(ibd_df_rp5_lower_df, var = "sample.x")

ibd_df_rp5_upper_df  <- as.data.frame(ibd_upper_rp5)
ibd_df_rp5_upper_df  <- tibble::rownames_to_column(ibd_df_rp5_upper_df, var = "sample.x")

#this is a DF with all ibd data in longform between non-matching pairs of samples
ibd_df_long_rp5 <-  ibd_df_rp5_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd"
  ) %>% filter(!is.na(ibd))
hist(ibd_df_long_rp5$ibd)
# CIs for rnull >0.125
CI_lower_df_long_r5 <- ibd_df_rp5_lower_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_lower")
CI_upper_df_long_r5 <- ibd_df_rp5_upper_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_upper")
# bind CIs into long df for rnull > 0.125
ibd_df_long_rp5 <- ibd_df_long_rp5 %>%
  left_join(CI_lower_df_long_r5) %>%
  left_join(CI_upper_df_long_r5)
pvals_r5 <- dres_aa[, , "p_value"]
pvals_r5 <- as.data.frame(pvals_r5)
pvals_r5 <- tibble::rownames_to_column(pvals_r5, var = "sample.x")


# pval 0.05
pvals_df_long_r5 <- pvals_r5 %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "p_value"
  ) %>% filter(!is.na(p_value)) %>%
  mutate(p_value = p_value/2) # divide by 2 for 1 sided test
# bind pvals into long df for rnull > 0.125
ibd_df_long_rp5 <- ibd_df_long_rp5 %>%
  left_join(pvals_df_long_r5)
# make p-value very large if r <0.125, set = 1
ibd_df_long_rp5 <- ibd_df_long_rp5 %>% mutate(p_value = ifelse(ibd <0.125, 1, p_value))
#here is where i do corrections for rnull > 0.5 dataframe
long_ibd_rp5 <- ibd_df_long_rp5 %>%
  mutate(p_BH = p.adjust(p_value, method = "BH"),
         p_BF = p.adjust(p_value, method = "bonferroni")) %>%
  mutate(sig_0.01 = ifelse(p_value < 0.01, 1, 0), sig_0.01_BH = ifelse(p_BH < 0.01, 1, 0),
         sig_0.01_BF = ifelse(p_BF < 0.01, 1, 0)) %>%
  mutate(sig_0.05 = ifelse(p_value < 0.05, 1, 0), sig_0.05_BH = ifelse(p_BH < 0.05, 1, 0),
         sig_0.05_BF = ifelse(p_BF < 0.05, 1, 0))

long_ibd_rp5 %<>% 
  filter(grepl("D0", sample.x)) %>% 
  filter(grepl("D0", sample.y)) %>% 
  filter(p_BH <= 0.05)



threshold <- 0.5

# Get all sample names in the dataset
all_samples <- unique(c(long_ibd_rp5$sample.x, long_ibd_rp5$sample.y))

# Identify samples that have IBD >= threshold with at least one other sample
samples_with_ibd <- long_ibd_rp5 %>%
  filter(ibd >= threshold) %>%
  select(sample.x, sample.y) %>%
  tidyr::pivot_longer(cols = everything(), values_to = "sample") %>%
  distinct(sample) %>%
  pull(sample)

# Proportion of samples involved in such pairs
proportion <- length(samples_with_ibd) / length(all_samples)



print(length(all_samples))


print(length(samples_with_ibd))

proportion



```

## Dcifer - Abobo

```{r dcifer_Gambella, echo=FALSE}

mydsmp <- dcifer::formatDat(filtered_dat_ac, "sample_id", "locus", "allele")
coi   <- coi_ac$post_coi_med
afreq_ac <- dcifer::calcAfreq(mydsmp, coi)

dsmp <- dcifer::matchAfreq(mydsmp, afreq_ac)
dres_ac <- ibdDat(dsmp$dsmp, coi, dsmp$afreq, pval = TRUE, confint = TRUE, rnull = 0.125, alpha = 0.01, nr = 1e3)

## Output dcifer results
readr::write_rds(dres_ac, "../data/Dcifer/Results/dcifer_result_ac.rds")

# === Relatedness matrix ===
dmat <- dres_ac[, , "estimate"]
dmat[upper.tri(dmat)] <- t(dmat)[upper.tri(t(dmat))]

# === Adjust p-values ===
dres0 <- dres_ac[, , "p_value"]
pval_vec <- dres0[lower.tri(dres0)]                           # raw p-values
pval_adj <- p.adjust(pval_vec, method = "BH")                 # [MODIFIED]: BH-adjustment
adjusted_pmat <- matrix(NA, nrow = nrow(dres0), ncol = ncol(dres0),
                        dimnames = dimnames(dres0))           # [MODIFIED]
adjusted_pmat[lower.tri(adjusted_pmat)] <- pval_adj          # [MODIFIED]
adjusted_pmat[upper.tri(adjusted_pmat)] <- t(adjusted_pmat)[upper.tri(adjusted_pmat)]  # symmetric

# === Categorize adjusted p-values ===
dres_low <- -log10(adjusted_pmat)                            # [MODIFIED]
dres_low[dres_low > -log10(0.0001)] <- "P << 0.001"
dres_low[dres_low <= -log10(0.0001) & dres_low > -log10(0.001)] <- "P < 0.001"
dres_low[dres_low <= -log10(0.001) & dres_low > -log10(0.01)] <- "P < 0.01"
dres_low[dres_low <= -log10(0.01)] <- "Not Sig"
dres_low[upper.tri(dres_low)] <- t(dres_low)[upper.tri(dres_low)]

# === Matrix to long format ===
matrix_to_longformat <- function(mat) {
  as.data.frame(mat) |>
    rownames_to_column("sample.x") |>
    pivot_longer(-sample.x, names_to = "sample.y", values_to = "value")
}

to_plot1 <- matrix_to_longformat(dmat)

# === Clustering for ordering ===
hc <- hclust(as.dist(1 - dmat))
order <- hc$order
clustered_levels <- colnames(dmat)[order]

to_plot1 <- to_plot1 |>
  mutate(sample.x = fct_relevel(sample.x, clustered_levels),
         sample.y = fct_relevel(sample.y, clustered_levels))

to_plot2 <- matrix_to_longformat(dres_low)
to_plot2 <- to_plot2 |>
  mutate(sample.x = fct_relevel(sample.x, clustered_levels),
         sample.y = fct_relevel(sample.y, clustered_levels))

# === Define colors ===
colours_upper <- colorRampPalette(c("#FFFFFF", "#1520A6"))(100)
colours_lower <- c("P << 0.001" = "#C70039", "P < 0.001" = "#FF5733", "P < 0.01" = "#FFC300", "Not Sig" = "white")

# === Filter for D0 samples ===
to_plot1_0 <- to_plot1[grepl("D0", to_plot1$sample.x) & grepl("D0", to_plot1$sample.y), ]
to_plot2_0 <- to_plot2[grepl("D0", to_plot2$sample.x) & grepl("D0", to_plot2$sample.y), ]

sample_d0 <- unique(to_plot1_0$sample.x)
cluster_0 <- clustered_levels[clustered_levels %in% sample_d0]

to_plot1_0 <- to_plot1_0 |>
  mutate(sample.x = fct_relevel(sample.x, cluster_0),
         sample.y = fct_relevel(sample.y, cluster_0))

to_plot2_0 <- to_plot2_0 |>
  mutate(sample.x = fct_relevel(sample.x, cluster_0),
         sample.y = fct_relevel(sample.y, cluster_0))

pdf("../data/Dcifer/Results/figures/Abobo_relatedness_pvalue_corrected.pdf", width = 10, height = 10)

# === Heatmap ===
g0_ac <- ggplot(to_plot1_0, aes(x = sample.x, y = sample.y)) +
  geom_tile(data = subset(to_plot1_0, as.numeric(sample.x) > as.numeric(sample.y)), aes(fill = value)) +
  scale_fill_gradientn(colors = colours_upper, name = "Relatedness") +
  theme(axis.text.x = element_blank()) +
  theme(axis.text.y = element_blank()) +
  new_scale_fill() +
  geom_tile(data = subset(to_plot2_0, as.numeric(sample.x) < as.numeric(sample.y)),
            aes(fill = value)) +
  scale_fill_manual(values = colours_lower, name = "Significance", 
                    breaks = c("P << 0.001", "P < 0.001", "P < 0.01")) +
  scale_x_discrete(limits = cluster_0) +
  scale_y_discrete(limits = cluster_0) +
  ggtitle("Abobo (BH adjusted pvalue)")

g0_ac


# Step 1: Get all related pairs (IBD > 0)
related_pairs <- to_plot1_0 %>% filter(value > 0)

# Step 2: Get all unique samples involved in related pairs
related_samples <- unique(c(related_pairs$sample.x, related_pairs$sample.y))

# Step 3: Get all unique samples in the full dataset
all_samples <- unique(c(to_plot1_0$sample.x, to_plot1_0$sample.y))

# Step 4: Compute proportion
proportion_related <- length(related_samples) / length(all_samples)

# Show result
proportion_related





# Extract IBD estimates from rnull > 0.125 run
ibd_est_rp5 <- dres_ac[, , "estimate"]
ibd_lower_rp5 <- dres_ac[, , "CI_lower"]
ibd_upper_rp5 <- dres_ac[, , "CI_upper"]

ibd_df_rp5_df  <- as.data.frame(ibd_est_rp5)
ibd_df_rp5_df  <- tibble::rownames_to_column(ibd_df_rp5_df, var = "sample.x")

ibd_df_rp5_lower_df  <- as.data.frame(ibd_lower_rp5)
ibd_df_rp5_lower_df  <- tibble::rownames_to_column(ibd_df_rp5_lower_df, var = "sample.x")

ibd_df_rp5_upper_df  <- as.data.frame(ibd_upper_rp5)
ibd_df_rp5_upper_df  <- tibble::rownames_to_column(ibd_df_rp5_upper_df, var = "sample.x")

#this is a DF with all ibd data in longform between non-matching pairs of samples
ibd_df_long_rp5 <-  ibd_df_rp5_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd"
  ) %>% filter(!is.na(ibd))
hist(ibd_df_long_rp5$ibd)
# CIs for rnull >0.125
CI_lower_df_long_r5 <- ibd_df_rp5_lower_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_lower")
CI_upper_df_long_r5 <- ibd_df_rp5_upper_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_upper")
# bind CIs into long df for rnull > 0.125
ibd_df_long_rp5 <- ibd_df_long_rp5 %>%
  left_join(CI_lower_df_long_r5) %>%
  left_join(CI_upper_df_long_r5)
pvals_r5 <- dres_ac[, , "p_value"]
pvals_r5 <- as.data.frame(pvals_r5)
pvals_r5 <- tibble::rownames_to_column(pvals_r5, var = "sample.x")


# pval 0.05
pvals_df_long_r5 <- pvals_r5 %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "p_value"
  ) %>% filter(!is.na(p_value)) %>%
  mutate(p_value = p_value/2) # divide by 2 for 1 sided test
# bind pvals into long df for rnull > 0.125
ibd_df_long_rp5 <- ibd_df_long_rp5 %>%
  left_join(pvals_df_long_r5)
# make p-value very large if r <0.125, set = 1
ibd_df_long_rp5 <- ibd_df_long_rp5 %>% mutate(p_value = ifelse(ibd <0.125, 1, p_value))
#here is where i do corrections for rnull > 0.5 dataframe
long_ibd_rp5 <- ibd_df_long_rp5 %>%
  mutate(p_BH = p.adjust(p_value, method = "BH"),
         p_BF = p.adjust(p_value, method = "bonferroni")) %>%
  mutate(sig_0.01 = ifelse(p_value < 0.01, 1, 0), sig_0.01_BH = ifelse(p_BH < 0.01, 1, 0),
         sig_0.01_BF = ifelse(p_BF < 0.01, 1, 0)) %>%
  mutate(sig_0.05 = ifelse(p_value < 0.05, 1, 0), sig_0.05_BH = ifelse(p_BH < 0.05, 1, 0),
         sig_0.05_BF = ifelse(p_BF < 0.05, 1, 0))

long_ibd_rp5 %<>% 
  filter(grepl("D0", sample.x)) %>% 
  filter(grepl("D0", sample.y))




threshold <- 0.5

# Get all sample names in the dataset
all_samples <- unique(c(long_ibd_rp5$sample.x, long_ibd_rp5$sample.y))

# Identify samples that have IBD >= threshold with at least one other sample
samples_with_ibd <- long_ibd_rp5 %>%
  filter(ibd >= threshold) %>%
  filter(p_BH <= 0.05) %>% 
  select(sample.x, sample.y) %>%
  tidyr::pivot_longer(cols = everything(), values_to = "sample") %>%
  distinct(sample) %>%
  pull(sample)

# Proportion of samples involved in such pairs
proportion <- length(samples_with_ibd) / length(all_samples)


print(length(all_samples))


print(length(samples_with_ibd))


proportion




```

## Dcifer - Assossa/Benshangul

```{r dcifer_Assossa, echo=FALSE}

mydsmp <- dcifer::formatDat(filtered_dat_ba, "sample_id", "locus", "allele")
coi   <- coi_ba$post_coi_med
afreq_ba <- dcifer::calcAfreq(mydsmp, coi)

dsmp <- dcifer::matchAfreq(mydsmp, afreq_ba)

dres_ba <- ibdDat(dsmp$dsmp, coi, dsmp$afreq, pval = TRUE, confint = TRUE, rnull = 0.125, alpha = 0.01, nr = 1e3)

## Output dcifer results
readr::write_rds(dres_ba, "../data/Dcifer/Results/dcifer_result_ba.rds")

# === Relatedness matrix ===
dmat <- dres_ba[, , "estimate"]
dmat[upper.tri(dmat)] <- t(dmat)[upper.tri(t(dmat))]

# === Adjust p-values ===
dres0 <- dres_ba[, , "p_value"]
pval_vec <- dres0[lower.tri(dres0)]                           # raw p-values
pval_adj <- p.adjust(pval_vec, method = "BH")                 # [MODIFIED]: BH-adjustment
adjusted_pmat <- matrix(NA, nrow = nrow(dres0), ncol = ncol(dres0),
                        dimnames = dimnames(dres0))           # [MODIFIED]
adjusted_pmat[lower.tri(adjusted_pmat)] <- pval_adj          # [MODIFIED]
adjusted_pmat[upper.tri(adjusted_pmat)] <- t(adjusted_pmat)[upper.tri(adjusted_pmat)]  # symmetric

# === Categorize adjusted p-values ===
dres_low <- -log10(adjusted_pmat)                            # [MODIFIED]
dres_low[dres_low > -log10(0.0001)] <- "P << 0.001"
dres_low[dres_low <= -log10(0.0001) & dres_low > -log10(0.001)] <- "P < 0.001"
dres_low[dres_low <= -log10(0.001) & dres_low > -log10(0.01)] <- "P < 0.01"
dres_low[dres_low <= -log10(0.01)] <- "Not Sig"
dres_low[upper.tri(dres_low)] <- t(dres_low)[upper.tri(dres_low)]

# === Matrix to long format ===
matrix_to_longformat <- function(mat) {
  as.data.frame(mat) |>
    rownames_to_column("sample.x") |>
    pivot_longer(-sample.x, names_to = "sample.y", values_to = "value")
}

to_plot1 <- matrix_to_longformat(dmat)

# === Clustering for ordering ===
hc <- hclust(as.dist(1 - dmat))
order <- hc$order
clustered_levels <- colnames(dmat)[order]

to_plot1 <- to_plot1 |>
  mutate(sample.x = fct_relevel(sample.x, clustered_levels),
         sample.y = fct_relevel(sample.y, clustered_levels))

to_plot2 <- matrix_to_longformat(dres_low)
to_plot2 <- to_plot2 |>
  mutate(sample.x = fct_relevel(sample.x, clustered_levels),
         sample.y = fct_relevel(sample.y, clustered_levels))

# === Define colors ===
colours_upper <- colorRampPalette(c("#FFFFFF", "#1520A6"))(100)
colours_lower <- c("P << 0.001" = "#C70039", "P < 0.001" = "#FF5733", "P < 0.01" = "#FFC300", "Not Sig" = "white")

# === Filter for D0 samples ===
to_plot1_0 <- to_plot1[grepl("D0", to_plot1$sample.x) & grepl("D0", to_plot1$sample.y), ]
to_plot2_0 <- to_plot2[grepl("D0", to_plot2$sample.x) & grepl("D0", to_plot2$sample.y), ]

sample_d0 <- unique(to_plot1_0$sample.x)
cluster_0 <- clustered_levels[clustered_levels %in% sample_d0]

to_plot1_0 <- to_plot1_0 |>
  mutate(sample.x = fct_relevel(sample.x, cluster_0),
         sample.y = fct_relevel(sample.y, cluster_0))

to_plot2_0 <- to_plot2_0 |>
  mutate(sample.x = fct_relevel(sample.x, cluster_0),
         sample.y = fct_relevel(sample.y, cluster_0))

pdf("../data/Dcifer/Results/figures/Assosa_relatedness_pvalue_corrected.pdf", width = 10, height = 10)

# === Heatmap ===
g0_ba <- ggplot(to_plot1_0, aes(x = sample.x, y = sample.y)) +
  geom_tile(data = subset(to_plot1_0, as.numeric(sample.x) > as.numeric(sample.y)), aes(fill = value)) +
  scale_fill_gradientn(colors = colours_upper, name = "Relatedness") +
  theme(axis.text.x = element_blank()) +
  theme(axis.text.y = element_blank()) +
  new_scale_fill() +
  geom_tile(data = subset(to_plot2_0, as.numeric(sample.x) < as.numeric(sample.y)),
            aes(fill = value)) +
  scale_fill_manual(values = colours_lower, name = "Significance", 
                    breaks = c("P << 0.001", "P < 0.001", "P < 0.01")) +
  scale_x_discrete(limits = cluster_0) +
  scale_y_discrete(limits = cluster_0) +
  ggtitle("Benshangul (BH adjusted pvalue)")

g0_ba

# Step 1: Get all related pairs (IBD > 0)
related_pairs <- to_plot1_0 %>% filter(value > 0)

# Step 2: Get all unique samples involved in related pairs
related_samples <- unique(c(related_pairs$sample.x, related_pairs$sample.y))

# Step 3: Get all unique samples in the full dataset
all_samples <- unique(c(to_plot1_0$sample.x, to_plot1_0$sample.y))

# Step 4: Compute proportion
proportion_related <- length(related_samples) / length(all_samples)

# Show result
proportion_related







# Extract IBD estimates from rnull > 0.125 run
ibd_est_rp5 <- dres_ba[, , "estimate"]
ibd_lower_rp5 <- dres_ba[, , "CI_lower"]
ibd_upper_rp5 <- dres_ba[, , "CI_upper"]

ibd_df_rp5_df  <- as.data.frame(ibd_est_rp5)
ibd_df_rp5_df  <- tibble::rownames_to_column(ibd_df_rp5_df, var = "sample.x")

ibd_df_rp5_lower_df  <- as.data.frame(ibd_lower_rp5)
ibd_df_rp5_lower_df  <- tibble::rownames_to_column(ibd_df_rp5_lower_df, var = "sample.x")

ibd_df_rp5_upper_df  <- as.data.frame(ibd_upper_rp5)
ibd_df_rp5_upper_df  <- tibble::rownames_to_column(ibd_df_rp5_upper_df, var = "sample.x")

#this is a DF with all ibd data in longform between non-matching pairs of samples
ibd_df_long_rp5 <-  ibd_df_rp5_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd"
  ) %>% filter(!is.na(ibd))
hist(ibd_df_long_rp5$ibd)
# CIs for rnull >0.125
CI_lower_df_long_r5 <- ibd_df_rp5_lower_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_lower")
CI_upper_df_long_r5 <- ibd_df_rp5_upper_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_upper")
# bind CIs into long df for rnull > 0.125
ibd_df_long_rp5 <- ibd_df_long_rp5 %>%
  left_join(CI_lower_df_long_r5) %>%
  left_join(CI_upper_df_long_r5)
pvals_r5 <- dres_ba[, , "p_value"]
pvals_r5 <- as.data.frame(pvals_r5)
pvals_r5 <- tibble::rownames_to_column(pvals_r5, var = "sample.x")


# pval 0.05
pvals_df_long_r5 <- pvals_r5 %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "p_value"
  ) %>% filter(!is.na(p_value)) %>%
  mutate(p_value = p_value/2) # divide by 2 for 1 sided test
# bind pvals into long df for rnull > 0.125
ibd_df_long_rp5 <- ibd_df_long_rp5 %>%
  left_join(pvals_df_long_r5)
# make p-value very large if r <0.125, set = 1
ibd_df_long_rp5 <- ibd_df_long_rp5 %>% mutate(p_value = ifelse(ibd <0.125, 1, p_value))
#here is where i do corrections for rnull > 0.5 dataframe
long_ibd_rp5 <- ibd_df_long_rp5 %>%
  mutate(p_BH = p.adjust(p_value, method = "BH"),
         p_BF = p.adjust(p_value, method = "bonferroni")) %>%
  mutate(sig_0.01 = ifelse(p_value < 0.01, 1, 0), sig_0.01_BH = ifelse(p_BH < 0.01, 1, 0),
         sig_0.01_BF = ifelse(p_BF < 0.01, 1, 0)) %>%
  mutate(sig_0.05 = ifelse(p_value < 0.05, 1, 0), sig_0.05_BH = ifelse(p_BH < 0.05, 1, 0),
         sig_0.05_BF = ifelse(p_BF < 0.05, 1, 0))

long_ibd_rp5 %<>% 
  filter(grepl("D0", sample.x)) %>% 
  filter(grepl("D0", sample.y))


threshold <- 0.5

# Get all sample names in the dataset
all_samples <- unique(c(long_ibd_rp5$sample.x, long_ibd_rp5$sample.y))

# Identify samples that have IBD >= threshold with at least one other sample
samples_with_ibd <- long_ibd_rp5 %>%
  filter(ibd >= threshold) %>%
  filter(p_BH <= 0.05) %>% 
  select(sample.x, sample.y) %>%
  tidyr::pivot_longer(cols = everything(), values_to = "sample") %>%
  distinct(sample) %>%
  pull(sample)

# Proportion of samples involved in such pairs
proportion <- length(samples_with_ibd) / length(all_samples)

print(length(all_samples))


print(length(samples_with_ibd))


proportion



```

## Dcifer - Maksegnit

```{r dcifer_Gondar, echo=FALSE}

mydsmp <- dcifer::formatDat(filtered_dat_gm, "sample_id", "locus", "allele")
coi   <- coi_gm$post_coi_med
afreq_gm <- dcifer::calcAfreq(mydsmp, coi)

dsmp <- dcifer::matchAfreq(mydsmp, afreq_gm)
dres_gm <- ibdDat(dsmp$dsmp, coi, dsmp$afreq, pval = TRUE, confint = TRUE, rnull = 0.125, alpha = 0.01, nr = 1e3)

## Output dcifer results
readr::write_rds(dres_gm, "../data/Dcifer/Results/dcifer_result_gm.rds")

# === Relatedness matrix ===
dmat <- dres_gm[, , "estimate"]
dmat[upper.tri(dmat)] <- t(dmat)[upper.tri(t(dmat))]

# === Adjust p-values ===
dres0 <- dres_gm[, , "p_value"]
pval_vec <- dres0[lower.tri(dres0)]                           # raw p-values
pval_adj <- p.adjust(pval_vec, method = "BH")                 # [MODIFIED]: BH-adjustment
adjusted_pmat <- matrix(NA, nrow = nrow(dres0), ncol = ncol(dres0),
                        dimnames = dimnames(dres0))           # [MODIFIED]
adjusted_pmat[lower.tri(adjusted_pmat)] <- pval_adj          # [MODIFIED]
adjusted_pmat[upper.tri(adjusted_pmat)] <- t(adjusted_pmat)[upper.tri(adjusted_pmat)]  # symmetric

# === Categorize adjusted p-values ===
dres_low <- -log10(adjusted_pmat)                            # [MODIFIED]
dres_low[dres_low > -log10(0.0001)] <- "P << 0.001"
dres_low[dres_low <= -log10(0.0001) & dres_low > -log10(0.001)] <- "P < 0.001"
dres_low[dres_low <= -log10(0.001) & dres_low > -log10(0.01)] <- "P < 0.01"
dres_low[dres_low <= -log10(0.01)] <- "Not Sig"
dres_low[upper.tri(dres_low)] <- t(dres_low)[upper.tri(dres_low)]

# === Matrix to long format ===
matrix_to_longformat <- function(mat) {
  as.data.frame(mat) |>
    rownames_to_column("sample.x") |>
    pivot_longer(-sample.x, names_to = "sample.y", values_to = "value")
}

to_plot1 <- matrix_to_longformat(dmat)

# === Clustering for ordering ===
hc <- hclust(as.dist(1 - dmat))
order <- hc$order
clustered_levels <- colnames(dmat)[order]

to_plot1 <- to_plot1 |>
  mutate(sample.x = fct_relevel(sample.x, clustered_levels),
         sample.y = fct_relevel(sample.y, clustered_levels))

to_plot2 <- matrix_to_longformat(dres_low)
to_plot2 <- to_plot2 |>
  mutate(sample.x = fct_relevel(sample.x, clustered_levels),
         sample.y = fct_relevel(sample.y, clustered_levels))

# === Define colors ===
colours_upper <- colorRampPalette(c("#FFFFFF", "#1520A6"))(100)
colours_lower <- c("P << 0.001" = "#C70039", "P < 0.001" = "#FF5733", "P < 0.01" = "#FFC300", "Not Sig" = "white")

# === Filter for D0 samples ===
to_plot1_0 <- to_plot1[grepl("D0", to_plot1$sample.x) & grepl("D0", to_plot1$sample.y), ]
to_plot2_0 <- to_plot2[grepl("D0", to_plot2$sample.x) & grepl("D0", to_plot2$sample.y), ]

sample_d0 <- unique(to_plot1_0$sample.x)
cluster_0 <- clustered_levels[clustered_levels %in% sample_d0]

to_plot1_0 <- to_plot1_0 |>
  mutate(sample.x = fct_relevel(sample.x, cluster_0),
         sample.y = fct_relevel(sample.y, cluster_0))

to_plot2_0 <- to_plot2_0 |>
  mutate(sample.x = fct_relevel(sample.x, cluster_0),
         sample.y = fct_relevel(sample.y, cluster_0))

pdf("../data/Dcifer/Results/figures/Maksegnit_relatedness_pvalue_corrected.pdf", width = 10, height = 10)

# === Heatmap ===
g0_gm <- ggplot(to_plot1_0, aes(x = sample.x, y = sample.y)) +
  geom_tile(data = subset(to_plot1_0, as.numeric(sample.x) > as.numeric(sample.y)), aes(fill = value)) +
  scale_fill_gradientn(colors = colours_upper, name = "Relatedness") +
  theme(axis.text.x = element_blank()) +
  theme(axis.text.y = element_blank()) +
  new_scale_fill() +
  geom_tile(data = subset(to_plot2_0, as.numeric(sample.x) < as.numeric(sample.y)),
            aes(fill = value)) +
  scale_fill_manual(values = colours_lower, name = "Significance", 
                    breaks = c("P << 0.001", "P < 0.001", "P < 0.01")) +
  scale_x_discrete(limits = cluster_0) +
  scale_y_discrete(limits = cluster_0) +
  ggtitle("Gondar (BH adjusted pvalue)")

g0_gm

# Step 1: Get all related pairs (IBD > 0)
related_pairs <- to_plot1_0 %>% filter(value > 0)

# Step 2: Get all unique samples involved in related pairs
related_samples <- unique(c(related_pairs$sample.x, related_pairs$sample.y))

# Step 3: Get all unique samples in the full dataset
all_samples <- unique(c(to_plot1_0$sample.x, to_plot1_0$sample.y))

# Step 4: Compute proportion
proportion_related <- length(related_samples) / length(all_samples)

# Show result
proportion_related







# Extract IBD estimates from rnull > 0.125 run
ibd_est_rp5 <- dres_gm[, , "estimate"]
ibd_lower_rp5 <- dres_gm[, , "CI_lower"]
ibd_upper_rp5 <- dres_gm[, , "CI_upper"]

ibd_df_rp5_df  <- as.data.frame(ibd_est_rp5)
ibd_df_rp5_df  <- tibble::rownames_to_column(ibd_df_rp5_df, var = "sample.x")

ibd_df_rp5_lower_df  <- as.data.frame(ibd_lower_rp5)
ibd_df_rp5_lower_df  <- tibble::rownames_to_column(ibd_df_rp5_lower_df, var = "sample.x")

ibd_df_rp5_upper_df  <- as.data.frame(ibd_upper_rp5)
ibd_df_rp5_upper_df  <- tibble::rownames_to_column(ibd_df_rp5_upper_df, var = "sample.x")

#this is a DF with all ibd data in longform between non-matching pairs of samples
ibd_df_long_rp5 <-  ibd_df_rp5_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd"
  ) %>% filter(!is.na(ibd))
hist(ibd_df_long_rp5$ibd)
# CIs for rnull >0.125
CI_lower_df_long_r5 <- ibd_df_rp5_lower_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_lower")
CI_upper_df_long_r5 <- ibd_df_rp5_upper_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_upper")
# bind CIs into long df for rnull > 0.125
ibd_df_long_rp5 <- ibd_df_long_rp5 %>%
  left_join(CI_lower_df_long_r5) %>%
  left_join(CI_upper_df_long_r5)
pvals_r5 <- dres_gm[, , "p_value"]
pvals_r5 <- as.data.frame(pvals_r5)
pvals_r5 <- tibble::rownames_to_column(pvals_r5, var = "sample.x")


# pval 0.05
pvals_df_long_r5 <- pvals_r5 %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "p_value"
  ) %>% filter(!is.na(p_value)) %>%
  mutate(p_value = p_value/2) # divide by 2 for 1 sided test
# bind pvals into long df for rnull > 0.125
ibd_df_long_rp5 <- ibd_df_long_rp5 %>%
  left_join(pvals_df_long_r5)
# make p-value very large if r <0.125, set = 1
ibd_df_long_rp5 <- ibd_df_long_rp5 %>% mutate(p_value = ifelse(ibd <0.125, 1, p_value))
#here is where i do corrections for rnull > 0.5 dataframe
long_ibd_rp5 <- ibd_df_long_rp5 %>%
  mutate(p_BH = p.adjust(p_value, method = "BH"),
         p_BF = p.adjust(p_value, method = "bonferroni")) %>%
  mutate(sig_0.01 = ifelse(p_value < 0.01, 1, 0), sig_0.01_BH = ifelse(p_BH < 0.01, 1, 0),
         sig_0.01_BF = ifelse(p_BF < 0.01, 1, 0)) %>%
  mutate(sig_0.05 = ifelse(p_value < 0.05, 1, 0), sig_0.05_BH = ifelse(p_BH < 0.05, 1, 0),
         sig_0.05_BF = ifelse(p_BF < 0.05, 1, 0))

long_ibd_rp5 %<>% 
  filter(grepl("D0", sample.x)) %>% 
  filter(grepl("D0", sample.y))




threshold <- 0.5

# Get all sample names in the dataset
all_samples <- unique(c(long_ibd_rp5$sample.x, long_ibd_rp5$sample.y))

# Identify samples that have IBD >= threshold with at least one other sample
samples_with_ibd <- long_ibd_rp5 %>%
  filter(ibd >= threshold) %>%
  filter(p_BH <= 0.05) %>% 
  select(sample.x, sample.y) %>%
  tidyr::pivot_longer(cols = everything(), values_to = "sample") %>%
  distinct(sample) %>%
  pull(sample)

# Proportion of samples involved in such pairs
proportion <- length(samples_with_ibd) / length(all_samples)

print(length(all_samples))


print(length(samples_with_ibd))

proportion



```

## Dcifer - Mizan

```{r dcifer_Mizan, echo=FALSE}

mydsmp <- dcifer::formatDat(filtered_dat_mm, "sample_id", "locus", "allele")
coi   <- coi_mm$post_coi_med
afreq_mm <- dcifer::calcAfreq(mydsmp, coi)

dsmp <- dcifer::matchAfreq(mydsmp, afreq_mm)
dres_mm <- ibdDat(dsmp$dsmp, coi, dsmp$afreq, pval = TRUE, confint = TRUE, rnull = 0.125, alpha = 0.05, nr = 1e3)

## Output dcifer results
readr::write_rds(dres_mm, "../data/Dcifer/Results/dcifer_result_mm.rds")

# === Relatedness matrix ===
dmat <- dres_mm[, , "estimate"]
dmat[upper.tri(dmat)] <- t(dmat)[upper.tri(t(dmat))]

# === Adjust p-values ===
dres0 <- dres_mm[, , "p_value"]
pval_vec <- dres0[lower.tri(dres0)]                           # raw p-values
pval_adj <- p.adjust(pval_vec, method = "BH")                 # [MODIFIED]: BH-adjustment
adjusted_pmat <- matrix(NA, nrow = nrow(dres0), ncol = ncol(dres0),
                        dimnames = dimnames(dres0))           # [MODIFIED]
adjusted_pmat[lower.tri(adjusted_pmat)] <- pval_adj          # [MODIFIED]
adjusted_pmat[upper.tri(adjusted_pmat)] <- t(adjusted_pmat)[upper.tri(adjusted_pmat)]  # symmetric

# === Categorize adjusted p-values ===
dres_low <- -log10(adjusted_pmat)                            # [MODIFIED]
dres_low[dres_low > -log10(0.0001)] <- "P << 0.001"
dres_low[dres_low <= -log10(0.0001) & dres_low > -log10(0.001)] <- "P < 0.001"
dres_low[dres_low <= -log10(0.001) & dres_low > -log10(0.01)] <- "P < 0.01"
dres_low[dres_low <= -log10(0.01)] <- "Not Sig"
dres_low[upper.tri(dres_low)] <- t(dres_low)[upper.tri(dres_low)]

# === Matrix to long format ===
matrix_to_longformat <- function(mat) {
  as.data.frame(mat) |>
    rownames_to_column("sample.x") |>
    pivot_longer(-sample.x, names_to = "sample.y", values_to = "value")
}

to_plot1 <- matrix_to_longformat(dmat)

# === Clustering for ordering ===
hc <- hclust(as.dist(1 - dmat))
order <- hc$order
clustered_levels <- colnames(dmat)[order]

to_plot1 <- to_plot1 |>
  mutate(sample.x = fct_relevel(sample.x, clustered_levels),
         sample.y = fct_relevel(sample.y, clustered_levels))

to_plot2 <- matrix_to_longformat(dres_low)
to_plot2 <- to_plot2 |>
  mutate(sample.x = fct_relevel(sample.x, clustered_levels),
         sample.y = fct_relevel(sample.y, clustered_levels))

# === Define colors ===
colours_upper <- colorRampPalette(c("#FFFFFF", "#1520A6"))(100)
colours_lower <- c("P << 0.001" = "#C70039", "P < 0.001" = "#FF5733", "P < 0.01" = "#FFC300", "Not Sig" = "white")

# === Filter for D0 samples ===
to_plot1_0 <- to_plot1[grepl("D0", to_plot1$sample.x) & grepl("D0", to_plot1$sample.y), ]
to_plot2_0 <- to_plot2[grepl("D0", to_plot2$sample.x) & grepl("D0", to_plot2$sample.y), ]

sample_d0 <- unique(to_plot1_0$sample.x)
cluster_0 <- clustered_levels[clustered_levels %in% sample_d0]

to_plot1_0 <- to_plot1_0 |>
  mutate(sample.x = fct_relevel(sample.x, cluster_0),
         sample.y = fct_relevel(sample.y, cluster_0))

to_plot2_0 <- to_plot2_0 |>
  mutate(sample.x = fct_relevel(sample.x, cluster_0),
         sample.y = fct_relevel(sample.y, cluster_0))

pdf("../data/Dcifer/Results/figures/Mizan_relatedness_pvalue_corrected.pdf", width = 10, height = 10)

# === Heatmap ===
g0_mm <- ggplot(to_plot1_0, aes(x = sample.x, y = sample.y)) +
  geom_tile(data = subset(to_plot1_0, as.numeric(sample.x) > as.numeric(sample.y)), aes(fill = value)) +
  scale_fill_gradientn(colors = colours_upper, name = "Relatedness") +
  theme(axis.text.x = element_blank()) +
  theme(axis.text.y = element_blank()) +
  new_scale_fill() +
  geom_tile(data = subset(to_plot2_0, as.numeric(sample.x) < as.numeric(sample.y)),
            aes(fill = value)) +
  scale_fill_manual(values = colours_lower, name = "Significance", 
                    breaks = c("P << 0.001", "P < 0.001", "P < 0.01")) +
  scale_x_discrete(limits = cluster_0) +
  scale_y_discrete(limits = cluster_0) +
  ggtitle("Mizan (BH adjusted pvalue)")

g0_mm

# Step 1: Get all related pairs (IBD > 0)
related_pairs <- to_plot1_0 %>% filter(value > 0)

# Step 2: Get all unique samples involved in related pairs
related_samples <- unique(c(related_pairs$sample.x, related_pairs$sample.y))

# Step 3: Get all unique samples in the full dataset
all_samples <- unique(c(to_plot1_0$sample.x, to_plot1_0$sample.y))

# Step 4: Compute proportion
proportion_related <- length(related_samples) / length(all_samples)

# Show result
proportion_related







# Extract IBD estimates from rnull > 0.125 run
ibd_est_rp5 <- dres_mm[, , "estimate"]
ibd_lower_rp5 <- dres_mm[, , "CI_lower"]
ibd_upper_rp5 <- dres_mm[, , "CI_upper"]

ibd_df_rp5_df  <- as.data.frame(ibd_est_rp5)
ibd_df_rp5_df  <- tibble::rownames_to_column(ibd_df_rp5_df, var = "sample.x")

ibd_df_rp5_lower_df  <- as.data.frame(ibd_lower_rp5)
ibd_df_rp5_lower_df  <- tibble::rownames_to_column(ibd_df_rp5_lower_df, var = "sample.x")

ibd_df_rp5_upper_df  <- as.data.frame(ibd_upper_rp5)
ibd_df_rp5_upper_df  <- tibble::rownames_to_column(ibd_df_rp5_upper_df, var = "sample.x")

#this is a DF with all ibd data in longform between non-matching pairs of samples
ibd_df_long_rp5 <-  ibd_df_rp5_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd"
  ) %>% filter(!is.na(ibd))
hist(ibd_df_long_rp5$ibd)
# CIs for rnull >0.125
CI_lower_df_long_r5 <- ibd_df_rp5_lower_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_lower")
CI_upper_df_long_r5 <- ibd_df_rp5_upper_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_upper")
# bind CIs into long df for rnull > 0.125
ibd_df_long_rp5 <- ibd_df_long_rp5 %>%
  left_join(CI_lower_df_long_r5) %>%
  left_join(CI_upper_df_long_r5)
pvals_r5 <- dres_mm[, , "p_value"]
pvals_r5 <- as.data.frame(pvals_r5)
pvals_r5 <- tibble::rownames_to_column(pvals_r5, var = "sample.x")


# pval 0.05
pvals_df_long_r5 <- pvals_r5 %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "p_value"
  ) %>% filter(!is.na(p_value)) %>%
  mutate(p_value = p_value/2) # divide by 2 for 1 sided test
# bind pvals into long df for rnull > 0.125
ibd_df_long_rp5 <- ibd_df_long_rp5 %>%
  left_join(pvals_df_long_r5)
# make p-value very large if r <0.125, set = 1
ibd_df_long_rp5 <- ibd_df_long_rp5 %>% mutate(p_value = ifelse(ibd <0.125, 1, p_value))
#here is where i do corrections for rnull > 0.5 dataframe
long_ibd_rp5 <- ibd_df_long_rp5 %>%
  mutate(p_BH = p.adjust(p_value, method = "BH"),
         p_BF = p.adjust(p_value, method = "bonferroni")) %>%
  mutate(sig_0.01 = ifelse(p_value < 0.01, 1, 0), sig_0.01_BH = ifelse(p_BH < 0.01, 1, 0),
         sig_0.01_BF = ifelse(p_BF < 0.01, 1, 0)) %>%
  mutate(sig_0.05 = ifelse(p_value < 0.05, 1, 0), sig_0.05_BH = ifelse(p_BH < 0.05, 1, 0),
         sig_0.05_BF = ifelse(p_BF < 0.05, 1, 0))

long_ibd_rp5 %<>% 
  filter(grepl("D0", sample.x)) %>% 
  filter(grepl("D0", sample.y))



threshold <- 0.5

# Get all sample names in the dataset
all_samples <- unique(c(long_ibd_rp5$sample.x, long_ibd_rp5$sample.y))

# Identify samples that have IBD >= threshold with at least one other sample
samples_with_ibd <- long_ibd_rp5 %>%
  filter(ibd >= threshold) %>%
  filter(p_BH <= 0.05) %>% 
  select(sample.x, sample.y) %>%
  tidyr::pivot_longer(cols = everything(), values_to = "sample") %>%
  distinct(sample) %>%
  pull(sample)

# Proportion of samples involved in such pairs
proportion <- length(samples_with_ibd) / length(all_samples)

print(length(all_samples))


print(length(samples_with_ibd))


proportion






```

## Run Dcifer - all
Now with the averaged allele frequencies, we can run dcifer on all samples together

```{r dcifer_all, echo=FALSE}

# format data
mydsmp <- dcifer::formatDat(filtered_dat, "sample_id", "locus", "allele")
# Combined COI, sorted correctly
coi <- coi_summary$post_coi_med

afreq <- dcifer::calcAfreq(mydsmp, coi)

dsmp <- dcifer::matchAfreq(mydsmp, afreq)

# Do this also for rnull >0.125
# Run Dcifer on full dataset, with averaged AF
dres_rp <- ibdDat(dsmp$dsmp, coi, dsmp$afreq, pval = TRUE, confint = TRUE, rnull = 0.125, alpha = 0.01, nr = 1e3)
## Output dcifer results
readr::write_rds(dres_rp, "../data/Dcifer/Results/dcifer_result_125_all.rds")


```

### P-value corrections
Dcifer p-value with Benjamini-Hochberg of Bonferroni corrections, code from Jessica (UG_IRS)

```{r p correction, exercise=TRUE, echo=F}

ibd_est <- dres[, , "estimate"]
ibd_lower <- dres[, , "CI_lower"]
ibd_upper <- dres[, , "CI_upper"]

# do stupid stuff to make dataframe (there's probably a better way to to this)
# rnull = 0
ibd_df <- as.data.frame(ibd_est)
ibd_df <- tibble::rownames_to_column(ibd_df, var = "sample.x")

ibd_lower_df <- as.data.frame(ibd_lower)
ibd_lower_df <- tibble::rownames_to_column(ibd_lower_df, var = "sample.x")

ibd_upper_df <- as.data.frame(ibd_upper)
ibd_upper_df <- tibble::rownames_to_column(ibd_upper_df, var = "sample.x")

#this is a DF with all ibd data in longform between non-matching pairs of samples
ibd_df_long <- ibd_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd"
  ) %>% filter(!is.na(ibd)) 

hist(ibd_df_long$ibd)

# BIND IN CI VALUES TO EACH DF
CI_lower_df_long <- ibd_lower_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_lower")
CI_upper_df_long <- ibd_upper_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_upper")
# bind CIs into long df 
ibd_df_long <- ibd_df_long %>% 
  left_join(CI_lower_df_long) %>% 
  left_join(CI_upper_df_long)

#### BIND IN P VALUES TO EACH DF ###############
pvals <- dres[, , "p_value"]
pvals <- as.data.frame(pvals)
pvals <- tibble::rownames_to_column(pvals, var = "sample.x")

pvals_df_long <- pvals %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "p_value"
  ) %>% filter(!is.na(p_value)) 


# bind pvals into long df 
ibd_df_long <- ibd_df_long %>% 
  left_join(pvals_df_long)

# here is where i do corrections for rnull = 0 dataframe
long_ibd <- ibd_df_long %>% 
  mutate(p_BH = p.adjust(p_value, method = "BH"), 
         p_BF = p.adjust(p_value, method = "bonferroni")) %>% 
  mutate(sig_0.01 = ifelse(p_value < 0.01, 1, 0), sig_0.01_BH = ifelse(p_BH < 0.01, 1, 0), 
         sig_0.01_BF = ifelse(p_BF < 0.01, 1, 0)) %>% 
  mutate(sig_0.05 = ifelse(p_value < 0.05, 1, 0), sig_0.05_BH = ifelse(p_BH < 0.05, 1, 0),
         sig_0.05_BF= ifelse(p_BF < 0.05, 1, 0)) %>%
  mutate(same_site = ifelse(substring(sample.x, 1, 2) == substring(sample.y, 1, 2), 
                            "Same site", 
                            "Different site"))



#### NOW DO THE SAME FOR THE RNULL > 0.125 DATA ###############

# Extract IBD estimates from rnull > 0.125 run
ibd_est_rp5 <- dres_rp[, , "estimate"]
ibd_lower_rp5 <- dres_rp[, , "CI_lower"]
ibd_upper_rp5 <- dres_rp[, , "CI_upper"]

ibd_df_rp5_df  <- as.data.frame(ibd_est_rp5)
ibd_df_rp5_df  <- tibble::rownames_to_column(ibd_df_rp5_df, var = "sample.x")

ibd_df_rp5_lower_df  <- as.data.frame(ibd_lower_rp5)
ibd_df_rp5_lower_df  <- tibble::rownames_to_column(ibd_df_rp5_lower_df, var = "sample.x")

ibd_df_rp5_upper_df  <- as.data.frame(ibd_upper_rp5)
ibd_df_rp5_upper_df  <- tibble::rownames_to_column(ibd_df_rp5_upper_df, var = "sample.x")

#this is a DF with all ibd data in longform between non-matching pairs of samples
ibd_df_long_rp5 <-  ibd_df_rp5_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd"
  ) %>% filter(!is.na(ibd))
hist(ibd_df_long_rp5$ibd)
# CIs for rnull >0.125
CI_lower_df_long_r5 <- ibd_df_rp5_lower_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_lower")
CI_upper_df_long_r5 <- ibd_df_rp5_upper_df %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "ibd_upper")
# bind CIs into long df for rnull > 0.125
ibd_df_long_rp5 <- ibd_df_long_rp5 %>%
  left_join(CI_lower_df_long_r5) %>%
  left_join(CI_upper_df_long_r5)
pvals_r5 <- dres_rp[, , "p_value"]
pvals_r5 <- as.data.frame(pvals_r5)
pvals_r5 <- tibble::rownames_to_column(pvals_r5, var = "sample.x")
# pval 0.05
pvals_df_long_r5 <- pvals_r5 %>%
  pivot_longer(
    cols = -sample.x,
    names_to = "sample.y",
    values_to = "p_value"
  ) %>% filter(!is.na(p_value)) %>%
  mutate(p_value = p_value/2) # divide by 2 for 1 sided test
# bind pvals into long df for rnull > 0.125
ibd_df_long_rp5 <- ibd_df_long_rp5 %>%
  left_join(pvals_df_long_r5)
# make p-value very large if r <0.125, set = 1
ibd_df_long_rp5 <- ibd_df_long_rp5 %>% mutate(p_value = ifelse(ibd <0.125, 1, p_value))
#here is where i do corrections for rnull > 0.5 dataframe
long_ibd_rp5 <- ibd_df_long_rp5 %>%
  mutate(p_BH = p.adjust(p_value, method = "BH"),
         p_BF = p.adjust(p_value, method = "bonferroni")) %>%
  mutate(sig_0.01 = ifelse(p_value < 0.01, 1, 0), sig_0.01_BH = ifelse(p_BH < 0.01, 1, 0),
         sig_0.01_BF = ifelse(p_BF < 0.01, 1, 0)) %>%
  mutate(sig_0.05 = ifelse(p_value < 0.05, 1, 0), sig_0.05_BH = ifelse(p_BH < 0.05, 1, 0),
         sig_0.05_BF = ifelse(p_BF < 0.05, 1, 0)) %>%
  mutate(same_site = ifelse(substring(sample.x, 1, 2) == substring(sample.y, 1, 2), 
                            "Same site", 
                            "Different site"))

write_csv(long_ibd_rp5,"../data/Dcifer/Results/long_ibd_125.csv")

```

### Look at different ways of defining highly related pairs 
##### Standard DCIFER run, adjusting for multiple comparisons (bonferroni and BH)
##### Testing null that r>0.125, adjustment for multiple comparisons (bonferroni and BH)
##### Sig levels 0.05 and 0.01 examined for all methods 

```{r compare_methods, echo= FALSE, warning = FALSE, message = FALSE, fig.show = 'hide'}

################# BONFERRONI CORRECTION,  R NULL = 0 ##########

# for rnull = 0 ibd df (standard), using FWER (bonferroni)
check_sig_0.05_BF <- long_ibd %>% filter(sig_0.05_BF == 1)
check_sig_0.01_BF <- long_ibd %>% filter(sig_0.01_BF == 1)


hist(check_sig_0.05_BF$ibd, main = "Histogram of r values after BF correction, p<0.05") #40
print("Number of pairs with significant relatedness after correction after Bonferroni correction (p<0.05)")
table(check_sig_0.05_BF$sig_0.05_BF)

hist(check_sig_0.01_BF$ibd, main  = "Histogram of r values after BF correction, p<0.01")
print("Number of pairs with significant relatedness after correction after Bonferroni correction, (p<0.01)")
table(check_sig_0.01_BF$sig_0.01_BF)


################# BENJAMINI-HOCHBERG CORRECTION, RNULL = 0  ##########

#for rnull = 0 ibd df (standard), using FDR
check_sig_0.05 <- long_ibd %>% filter(sig_0.05_BH == 1)
check_sig_0.01 <- long_ibd %>% filter(sig_0.01_BH == 1)

hist(check_sig_0.05$ibd, main = "Histogram of r values after BH correction, p<0.05")
print("Number of pairs with significant relatedness after BH correction (p<0.05)")
table(check_sig_0.05$sig_0.05_BH)

hist(check_sig_0.01$ibd, main  = "Histogram of r values after BH correction, p<0.01")
print("Number of pairs with significant relatedness after correction after BH correction (p<0.01)")
table(check_sig_0.01$sig_0.01_BH)

#look at significance level BF 
check_BF <- long_ibd %>% 
 filter(p_value <= 0.05) %>% 
  select(p_value, p_BF) %>% 
  filter(p_BF <= 0.05) %>% summarize(p_value = max(p_value))
#bonferroni threshold for rnull = 0 df
p_val_BF <- check_BF$p_value

#look at significance level 
check_BH <- long_ibd %>% filter(p_value <= 0.05) %>% 
  select(p_value, p_BH) %>% 
  filter(p_BH <= 0.05) %>% 
  summarize(p_value = max(p_value))
# bonferroni threshold
p_val_BH <- check_BH$p_value


# Plot between site relatedness
long_ibd_between <- long_ibd %>%
  filter(same_site == "Different site")
g_between <- ggplot(data = long_ibd_between, aes(x = log10(p_value), y = ibd)) + 
  geom_point(color = "darkblue") + 
  geom_vline(aes(xintercept = log10(0.05), linetype = "0.05")) + 
  geom_vline(aes(xintercept = log10(p_val_BF), linetype = "Bonferroni")) + 
  geom_vline(aes(xintercept = log10(p_val_BH), linetype = "Benjamini-Hochberg")) + 
  scale_linetype_manual(name = "Adjustment Method",
                        values = c("0.05" = "solid", 
                                   "Bonferroni" = "dashed", 
                                   "Benjamini-Hochberg" = "dotdash"),
                        labels = c("0.05", "Benjamini-Hochberg", "Bonferroni")) +
  ggtitle("IBD between populations") +
  theme(legend.position = "none")
g_between

# Plot between site relatedness
long_ibd_within <- long_ibd %>%
  filter(same_site == "Same site") %>%
  mutate(site = case_when(
    grepl(pattern = "AA", sample.x) ~ "Asayita",
    grepl(pattern = "AC", sample.x) ~ "Gambella",
    grepl(pattern = "GM", sample.x) ~ "Gondar",
    grepl(pattern = "BA", sample.x) ~ "Assossa",
    grepl(pattern = "ARC", sample.x) ~ "Djibouti",
    grepl(pattern = "MM", sample.x) ~ "Mizan"))
g_within <- ggplot(data = long_ibd_within, aes(x = log10(p_value), y = ibd, color = site)) +
  geom_point() + 
  geom_vline(aes(xintercept = log10(0.05), linetype = "0.05")) + 
  geom_vline(aes(xintercept = log10(p_val_BF), linetype = "Bonferroni")) + 
  geom_vline(aes(xintercept = log10(p_val_BH), linetype = "Benjamini-Hochberg")) + 
  scale_linetype_manual(name = "Adjustment Method",
                        values = c("0.05" = "solid", 
                                   "Bonferroni" = "dashed", 
                                   "Benjamini-Hochberg" = "dotdash"),
                        labels = c("0.05", "Benjamini-Hochberg", "Bonferroni")) +
  ggtitle("IBD within populations")
g_within



################# BONFERRONI CORRECTION,  R NULL > 0.125 ##########

# for rnull = 0 ibd df (standard), using FWER (bonferroni)
check_sig_0.05_BF <- long_ibd_rp5 %>% filter(sig_0.05_BF == 1)
check_sig_0.01_BF <- long_ibd_rp5 %>% filter(sig_0.01_BF == 1)


hist(check_sig_0.05_BF$ibd, main = "Histogram of r values after BF correction, p<0.05") #40
print("Number of pairs with significant relatedness after correction after Bonferroni correction (p<0.05)")
table(check_sig_0.05_BF$sig_0.05_BF)

hist(check_sig_0.01_BF$ibd, main  = "Histogram of r values after BF correction, p<0.01")
print("Number of pairs with significant relatedness after correction after Bonferroni correction, (p<0.01)")
table(check_sig_0.01_BF$sig_0.01_BF)



########## #################  NOW FOR RNULL > 0.125 ################# #################

################# BENJAMINI-HOCHBERG CORRECTION ##########

#for rnull >0.125 ibd df (standard), using FDR
check_sig_0.05 <- long_ibd_rp5 %>% filter(sig_0.05_BH == 1)
check_sig_0.01 <- long_ibd_rp5 %>% filter(sig_0.01_BH == 1)

hist(check_sig_0.05$ibd, main = "Histogram of r values after BH correction, p<0.05")
print("Number of pairs with significant relatedness after BH correction (p<0.05)")
table(check_sig_0.05$sig_0.05_BH)

hist(check_sig_0.01$ibd, main  = "Histogram of r values after BH correction, p<0.01")
print("Number of pairs with significant relatedness after correction after BH correction (p<0.01)")
table(check_sig_0.01$sig_0.01_BH)

#look at significance level BF
check_BF <- long_ibd_rp5 %>%
 filter(p_value <= 0.05) %>%
  select(p_value, p_BF) %>%
  filter(p_BF <= 0.05) %>% summarize(p_value = max(p_value))
#bonferroni threshold
p_val_BF_0.125 <- check_BF$p_value

#look at significance level
check_BH <- long_ibd_rp5 %>% filter(p_value <= 0.05) %>%
  select(p_value, p_BH) %>%
  filter(p_BH <= 0.05) %>%
  summarize(p_value = max(p_value))

# bonferroni threshold
p_val_BH_0.125 <- check_BH$p_value


# Plot between pop. relatedness
long_ibd_rp5_between <- long_ibd_rp5 %>%
  filter(same_site == "Different site")
g.125_between <- ggplot(data = long_ibd_rp5_between, aes(x = log10(p_value), y = ibd)) +
  geom_point(color = "darkblue") +
  geom_vline(aes(xintercept = log10(0.05), linetype = "0.05")) +
  geom_vline(aes(xintercept = log10(p_val_BF), linetype = "Bonferroni")) +
  geom_vline(aes(xintercept = log10(p_val_BH), linetype = "Benjamini-Hochberg")) +
  scale_linetype_manual(name = "Adjustment Method",
                        values = c("0.05" = "solid",
                                   "Bonferroni" = "dashed",
                                   "Benjamini-Hochberg" = "dotdash"),
                        labels = c("0.05", "Benjamini-Hochberg", "Bonferroni")) +
  ggtitle("IBD between populations (rnull > 0.125)") +
  theme(legend.position = "none")
g.125_between

# Plot between pop. relatedness
long_ibd_rp5_within <- long_ibd_rp5 %>%
  filter(same_site == "Same site") %>%
  mutate(site = case_when(
    grepl(pattern = "AA", sample.x) ~ "Asayita",
    grepl(pattern = "AC", sample.x) ~ "Gambella",
    grepl(pattern = "GM", sample.x) ~ "Gondar",
    grepl(pattern = "BA", sample.x) ~ "Assossa",
    grepl(pattern = "ARC", sample.x) ~ "Djibouti",
    grepl(pattern = "MM", sample.x) ~ "Mizan"))
g.125_within <- ggplot(data = long_ibd_rp5_within, aes(x = log10(p_value), y = ibd, color = site)) +
  geom_point() + 
  geom_vline(aes(xintercept = log10(0.05), linetype = "0.05")) + 
  geom_vline(aes(xintercept = log10(p_val_BF), linetype = "Bonferroni")) + 
  geom_vline(aes(xintercept = log10(p_val_BH), linetype = "Benjamini-Hochberg")) + 
  scale_linetype_manual(name = "Adjustment Method",
                        values = c("0.05" = "solid", 
                                   "Bonferroni" = "dashed", 
                                   "Benjamini-Hochberg" = "dotdash"),
                        labels = c("0.05", "Benjamini-Hochberg", "Bonferroni")) +
  ggtitle("IBD within populations (rnull > 0.125)")
g.125_within


g_p.thres <- ggarrange(g_between, g_within, g.125_between, g.125_within,
                       ncol = 2, nrow = 2)
g_p.thres

```



```{r import, exercise=TRUE, echo=F}

long_ibd_rp5 <- long_ibd_rp5 %>% 
  filter(grepl("D0", sample.x)) %>% 
  filter(grepl("D0", sample.y)) %>% 
  mutate(site.x = case_when(
    grepl("AA", sample.x) ~ "Asayita",
    grepl("AC", sample.x) ~ "Abobo",
    grepl("BA", sample.x) ~ "Asosa",
    grepl("GM", sample.x) ~ "Maksegnit",
    grepl("MM", sample.x) ~ "Mizan",
  )) %>% 
  mutate(site.y = case_when(
    grepl("AA", sample.y) ~ "Asayita",
    grepl("AC", sample.y) ~ "Abobo",
    grepl("BA", sample.y) ~ "Asosa",
    grepl("GM", sample.y) ~ "Maksegnit",
    grepl("MM", sample.y) ~ "Mizan",
  ))


check_BH <- long_ibd_rp5 %>% filter(p_value <= 0.05) %>%
  select(p_value, p_BH) %>%
  filter(p_BH <= 0.05) %>%
  summarize(p_value = max(p_value))


# threshold
p_val_BH_0.125 <- check_BH$p_value



threshold <- 0.5

# All samples with their sites
sample_sites <- long_ibd_rp5 %>%
  select(sample = sample.x, site = site.x) %>%
  bind_rows(long_ibd_rp5 %>% select(sample = sample.y, site = site.y)) %>%
  distinct()

# For each sample, identify all sites it is related to (via IBD≥threshold)
cross_links <- long_ibd_rp5 %>%
  filter(ibd >= threshold, site.x != site.y) %>%
  select(sample = sample.x, site = site.x, related_site = site.y) %>%
  bind_rows(
    long_ibd_rp5 %>%
      filter(ibd >= threshold, site.x != site.y) %>%
      select(sample = sample.y, site = site.y, related_site = site.x)
  ) %>%
  distinct(sample, site, related_site)

# Count for each site how many samples link to each other site
site_to_site_counts <- cross_links %>%
  group_by(site, related_site) %>%
  summarise(count = n_distinct(sample), .groups = "drop")

# Total samples per site
site_totals <- sample_sites %>%
  count(site, name = "total_samples")

# Merge & format
formatted <- site_to_site_counts %>%
  left_join(site_totals, by = "site") %>%
  mutate(
    formatted = sprintf("%d/%d (%.1f%%)", count, total_samples, 100 * count / total_samples)
  ) %>%
  select(site, related_site, formatted) %>%
  pivot_wider(names_from = related_site, values_from = formatted, values_fill = "0/total (0%)")

formatted



```


### Global clustering
Dcifer cluster plot, using BH correction

```{r network, exercise=TRUE, echo=F}

# Should only be looking at D0s
long_ibd_rp5_d0 <- long_ibd_rp5 %>%
  filter(grepl("D0", sample.x) & grepl("D0", sample.y))

# Define the function for plotting the network
plot_network <- function(long_ibd_rp5, relatedness_threshold = 0.5, pvalue_threshold = 0.05, 
                         scaling_factor = 1.0, jitter_amount = 0.1, 
                         show_isolates = TRUE, edge_color_by = same_site, vertex_size = 5, seed = 100) { 
  
  # Set seed for reproducibility
  set.seed(seed)
  
  # Filter edges based on relatedness (ibd) and p-value thresholds
  filtered_edges <- long_ibd %>%
    filter(ibd >= relatedness_threshold, p_BH < pvalue_threshold)

  # Create a data frame of unique sample names for vertices
  unique_samples <- unique(c(filtered_edges$sample.x, filtered_edges$sample.y))
  
  # Create an igraph object from the filtered edges
  g <- graph_from_data_frame(filtered_edges, directed = FALSE, vertices = data.frame(name = unique_samples))

  # Identify isolates (samples that are not in the edges)
  samples_in_edges <- unique(c(filtered_edges$sample.x, filtered_edges$sample.y))
  all_samples <- unique(c(long_ibd$sample.x, long_ibd$sample.y))
  isolates <- setdiff(all_samples, samples_in_edges)

  # Add isolates as vertices in the graph if show_isolates is TRUE
  if (show_isolates) {
    g_full <- g + vertices(isolates)
  } else {
    g_full <- g  # Only use the main graph without isolates
  }

  # # Merge metadata to get site information for coloring vertices
  vertex_data <- data.frame(name = V(g_full)$name) %>%
    mutate(site = case_when(
      grepl(pattern = "AA", name) ~ "Asayita",
      grepl(pattern = "AC", name) ~ "Gambella",
      grepl(pattern = "GM", name) ~ "Gondar",
      grepl(pattern = "BA", name) ~ "Assossa",
      grepl(pattern = "ARC", name) ~ "Djibouti",
      grepl(pattern = "MM", name) ~ "Mizan"))
    
  #   left_join(metadata_cluster, by = "name")  # Updated to use metadata_cluster

  # Check for NA in site and set default color if necessary
  if ("site" %in% colnames(vertex_data)) {
    # Set vertex colors based on site
    unique_populations <- unique(vertex_data$site)
    colors <- c("Asayita" = "darkred", "Gambella" = "yellow3",
                "Assossa" = "darkorange", "Gondar" = "blue", "Mizan" = "darkviolet")

    # Map colors to sites
    color_map <- setNames(colors, unique_populations)

    # Assign colors to vertices based on site
    V(g_full)$color <- color_map[as.character(vertex_data$site)]
    
    # Assign default color for any samples without a site
    V(g_full)$color[is.na(V(g_full)$color)] <- "lightgray"  # Default color for unmatched samples
  } else {
    V(g_full)$color <- "gray"  # Default color if site is not found
  }

  # Set vertex size
  V(g_full)$size <- vertex_size

  # Color edges based on the specified category (now "cat")
  if (edge_color_by %in% colnames(filtered_edges)) {
    edge_categories <- filtered_edges %>%
      select(sample.x, sample.y, !!sym(edge_color_by)) %>%
      distinct()

    # Ensure there are no NA values in the edge categories
    edge_categories <- edge_categories[!is.na(edge_categories[[edge_color_by]]), ]

    if (nrow(edge_categories) > 0) {
      # Manually set edge colors based on "same_site"
      E(g_full)$color <- ifelse(edge_categories[[edge_color_by]] == "Same site", "darkgoldenrod", 
                                ifelse(edge_categories[[edge_color_by]] == "Different site", "grey", NA))
      
      # Set a default color for any unmatched edges
      E(g_full)$color[is.na(E(g_full)$color)] <- "black"  # Use black for NA values
    } else {
      E(g_full)$color <- "black"  # If no edge categories, default to black
    }
  } else {
    E(g_full)$color <- "black"  # Default edge color if category not found
  }

  # Set edge widths to ensure visibility
  edge_widths <- pmax(E(g_full)$ibd * 2, 0.2)

  # Use layout_nicely() to automatically select an appropriate layout
  layout <- layout_nicely(g_full)

  # Jitter the layout coordinates to avoid overlap
  jitter_matrix <- matrix(rnorm(nrow(layout) * 2, mean = 0, sd = jitter_amount), ncol = 2)
  layout <- layout + jitter_matrix

  # Plot the graph
  plot(g_full, 
       layout = layout, 
       vertex.label = NA,  # Remove labels
       vertex.size = V(g_full)$size,  # Use modifiable vertex sizes
       edge.width = edge_widths,  # Use adjusted edge widths
       edge.color = E(g_full)$color,  # Use colors for edges
       main = paste("Relatedness Threshold:", 
                    relatedness_threshold))

  # Add a legend for the vertex colors (site)
  legend("bottomright", legend = names(color_map), col = color_map, pch = 19, pt.cex = 1.5, title = "site")
  # Add a legend for the edge colors (Category)
  legend("topright", legend = c("Same site", "Different site"), col = c("darkgoldenrod", "grey"), lty = 1, title = edge_color_by)
  
}


pdf("../data/Dcifer/Results/figures/global_relatedness_0.5_jitter.pdf", width = 12, height = 10)
# Plot clusters
plot_network(long_ibd_rp5_d0, relatedness_threshold = 0.5, pvalue_threshold = p_val_BH, scaling_factor = 1, jitter_amount = 0, vertex_size = 5, show_isolates = FALSE, edge_color_by = "same_site", seed = 300)

dev.off()


```


### Clustering by HRP deletions
Determine whether HRP-deleted samples are also connected by IBD, for each site

```{r network_hrp, exercise=TRUE, echo=F}

# Import CNV report
cnv_metadata <- read_csv("../data/Dcifer/Input/cnv_for_relatedness.csv")

# filter samples
cnv_samples <- cnv_metadata$SampleID
long_ibd_cnv <- long_ibd_rp5  %>%
  filter(sample.x %in% cnv_samples & sample.y %in% cnv_samples) %>% 
  filter(grepl("D0", sample.x) & grepl("D0", sample.y))

plot_network_cnv_site <- function(long_ibd_cnv, cnv_metadata, relatedness_threshold = 0.5, pvalue_threshold = 0.05, 
                                  show_isolates = TRUE, seed = 100, site_filter) {  

  # Set seed for reproducibility
  set.seed(seed)

  # Mapping of site abbreviations to full names
  site_name_mapping <- c("AA" = "Asayita", 
                         "AC" = "Abobo", 
                         "GM" = "Gondar", 
                         "BA" = "Assossa",
                         "MM" = "Mizan")

  # Full site name based on site_filter (abbreviation)
  full_site_name <- site_name_mapping[site_filter]
  
  # Filter the long_ibd_cnv data frame for relevant samples starting with the site filter
  filtered_samples <- long_ibd_cnv %>%
    filter(grepl(paste("^", site_filter, sep = ""), sample.x) & grepl(paste("^", site_filter, sep = ""), sample.y))

  # Filter edges based on relatedness (ibd) and p-value thresholds
  filtered_edges <- filtered_samples %>%
    filter(ibd >= relatedness_threshold, p_BH < pvalue_threshold)

  # Create a data frame of unique sample names for vertices
  unique_samples <- unique(c(filtered_edges$sample.x, filtered_edges$sample.y))

  # Create an igraph object from the filtered edges
  g <- graph_from_data_frame(filtered_edges, directed = FALSE, vertices = data.frame(name = unique_samples))

  # Identify isolates (samples that are not in the edges)
  samples_in_edges <- unique(c(filtered_edges$sample.x, filtered_edges$sample.y))
  all_samples <- unique(c(long_ibd_cnv$sample.x, long_ibd_cnv$sample.y))
  isolates <- setdiff(all_samples, samples_in_edges)

  # Add isolates as vertices in the graph if show_isolates is TRUE
  if (show_isolates) {
    g_full <- g + vertices(isolates)
  } else {
    g_full <- g  # Only use the main graph without isolates
  }

  # Merge site information and HRP deletion status into vertex_data
  vertex_data <- data.frame(name = V(g_full)$name) %>%
    mutate(site = case_when(
      grepl(pattern = "^AA", name) ~ "Asayita",
      grepl(pattern = "^AC", name) ~ "Gambella",
      grepl(pattern = "^GM", name) ~ "Gondar",
      grepl(pattern = "^BA", name) ~ "Assossa",
      grepl(pattern = "^MM", name) ~ "Mizan")) %>%
    left_join(cnv_metadata, by = c("name" = "SampleID"))  # Join cnv_metadata for HRP deletion status

  
  # Apply site color
  V(g_full)$color <- ifelse(vertex_data$HRP_stat == "HRP2_only_deleted",
                           "red", ifelse(vertex_data$HRP_stat == "HRP3_only_deleted", "blue", ifelse(vertex_data$HRP_stat == "HRP2/3_deleted", "purple", ifelse(vertex_data$HRP_stat == "no_deletion", "grey", "black"))))
  

  # Set all edge colors to black
  E(g_full)$color <- "black"

  # Set edge widths to ensure visibility
  edge_widths <- pmax(E(g_full)$ibd * 2, 0.2)

  # Use layout_nicely() to automatically select an appropriate layout without jitter
  layout <- layout_nicely(g_full)

  # Create the plot
  plot(g_full, 
       layout = layout, 
       vertex.label = NA,  # Remove labels
       vertex.size = V(g_full)$size,  # Use conditional vertex sizes
       edge.width = edge_widths,  # Use adjusted edge widths
       edge.color = E(g_full)$color,  # All edges set to black
       vertex.shape = V(g_full)$shape,  # Use conditional vertex shapes
       main = full_site_name,
       cex.main = 8) 
}


# Define the sites
site_list <- c("AA", "AC", "GM", "BA", "MM")  # Abbreviations for the sites


pdf("../data/Dcifer/Results/figures/HRP_deletion_relatedness_0.9D0.pdf", width = 10, height = 10)

# Set up the plotting grid
par(mfrow = c(1, 5))  # Set 2 rows and 3 columns for grid layout

# Loop over each site and call the function
for (site in site_list) { 
  # Print current site for debugging
  print(paste("Plotting for site:", site))
  
  # Run the function for each site, using the site abbreviation as the filter
  plot_network_cnv_site(long_ibd_cnv = long_ibd_cnv,  # Ensure you're using the correct dataset
                        cnv_metadata = cnv_metadata, 
                        relatedness_threshold = 0.9, 
                        pvalue_threshold = p_val_BH_0.125,
                        show_isolates = FALSE, 
                        seed = 200, 
                        site_filter = site)  # Pass the abbreviation to filter
}

dev.off()


```

### Clustering by K13 mutation

```{r network_hrp, exercise=TRUE, echo=F}

# Import CNV report
k13_metadata <- read_csv("../data/Dcifer/Input/k13_for_relatedness.csv")

# filter samples
cnv_samples <- k13_metadata$SampleID
long_ibd_cnv <- long_ibd_rp5  %>%
  filter(sample.x %in% cnv_samples & sample.y %in% cnv_samples) %>% 
  filter(grepl("D0", sample.x) & grepl("D0", sample.y))

plot_network_cnv_site <- function(long_ibd_cnv, k13_metadata, relatedness_threshold = 0.5, pvalue_threshold = 0.05, 
                                  show_isolates = TRUE, seed = 100, site_filter) {  

  # Set seed for reproducibility
  set.seed(seed)

  # Mapping of site abbreviations to full names
  site_name_mapping <- c("AA" = "Asayita", 
                         "AC" = "Abobo", 
                         "GM" = "Gondar", 
                         "BA" = "Assossa",
                         "MM" = "Mizan")

  # Full site name based on site_filter (abbreviation)
  full_site_name <- site_name_mapping[site_filter]
  
  # Filter the long_ibd_cnv data frame for relevant samples starting with the site filter
  filtered_samples <- long_ibd_cnv %>%
    filter(grepl(paste("^", site_filter, sep = ""), sample.x) & grepl(paste("^", site_filter, sep = ""), sample.y))

  # Filter edges based on relatedness (ibd) and p-value thresholds
  filtered_edges <- filtered_samples %>%
    filter(ibd >= relatedness_threshold, p_BH < pvalue_threshold)

  # Create a data frame of unique sample names for vertices
  unique_samples <- unique(c(filtered_edges$sample.x, filtered_edges$sample.y))

  # Create an igraph object from the filtered edges
  g <- graph_from_data_frame(filtered_edges, directed = FALSE, vertices = data.frame(name = unique_samples))

  # Identify isolates (samples that are not in the edges)
  samples_in_edges <- unique(c(filtered_edges$sample.x, filtered_edges$sample.y))
  all_samples <- unique(c(long_ibd_cnv$sample.x, long_ibd_cnv$sample.y))
  isolates <- setdiff(all_samples, samples_in_edges)

  # Add isolates as vertices in the graph if show_isolates is TRUE
  if (show_isolates) {
    g_full <- g + vertices(isolates)
  } else {
    g_full <- g  # Only use the main graph without isolates
  }

  # Merge site information and HRP deletion status into vertex_data
  vertex_data <- data.frame(name = V(g_full)$name) %>%
    mutate(site = case_when(
      grepl(pattern = "^AA", name) ~ "Asayita",
      grepl(pattern = "^AC", name) ~ "Gambella",
      grepl(pattern = "^GM", name) ~ "Gondar",
      grepl(pattern = "^BA", name) ~ "Assossa",
      grepl(pattern = "^MM", name) ~ "Mizan")) %>%
    left_join(k13_metadata, by = c("name" = "SampleID"))  # Join k13_metadata for HRP deletion status

  V(g_full)$color <- ifelse(vertex_data$R622_k13 == "I",
                           "#0077b6", ifelse(vertex_data$R622_k13 == "R", "#03045e", ifelse(vertex_data$R622_k13 == "R/I", "#90e0ef", ifelse(vertex_data$R622_k13 == "-", "grey", "black"))))

  # Set all edge colors to black
  E(g_full)$color <- "black"

  # Set edge widths to ensure visibility
  edge_widths <- pmax(E(g_full)$ibd * 2, 0.2)

  # Use layout_nicely() to automatically select an appropriate layout without jitter
  layout <- layout_nicely(g_full)

  # Create the plot
  plot(g_full, 
       layout = layout, 
       vertex.label = NA,  # Remove labels
       vertex.size = V(g_full)$size,  # Use conditional vertex sizes
       edge.width = edge_widths,  # Use adjusted edge widths
       edge.color = E(g_full)$color,  # All edges set to black
       vertex.shape = V(g_full)$shape,  # Use conditional vertex shapes
       main = full_site_name,
       cex.main = 8) 
}


# Define the sites
site_list <- c("AA", "AC", "GM", "BA", "MM")  # Abbreviations for the sites

pdf("../data/Dcifer/Results/figures/k13_relatedness_0.9D0.pdf", width = 10, height = 10)

# Set up the plotting grid
par(mfrow = c(1, 5))  # Set 2 rows and 3 columns for grid layout

# Loop over each site and call the function
for (site in site_list) { 
  # Print current site for debugging
  print(paste("Plotting for site:", site))
  
  # Run the function for each site, using the site abbreviation as the filter
  plot_network_cnv_site(long_ibd_cnv = long_ibd_cnv,  # Ensure you're using the correct dataset
                        k13_metadata = k13_metadata, 
                        relatedness_threshold = 0.9, 
                        pvalue_threshold = p_val_BH_0.125,
                        show_isolates = FALSE, 
                        seed = 200, 
                        site_filter = site)  # Pass the abbreviation to filter
}

# Close the JPEG device after plotting
dev.off()


```

### Clustering by crt mutation

```{r network_hrp, exercise=TRUE, echo=F}

# Import CNV report
crt_metadata <- read_csv("../data/Dcifer/Input/crt_for_relatedness.csv")

# filter samples
cnv_samples <- crt_metadata$SampleID
long_ibd_cnv <- long_ibd_rp5  %>%
  filter(sample.x %in% cnv_samples & sample.y %in% cnv_samples) %>% 
  filter(grepl("D0", sample.x) & grepl("D0", sample.y))

plot_network_cnv_site <- function(long_ibd_cnv, crt_metadata, relatedness_threshold = 0.5, pvalue_threshold = 0.05, 
                                  show_isolates = TRUE, seed = 100, site_filter) {  

  # Set seed for reproducibility
  set.seed(seed)

  # Mapping of site abbreviations to full names
  site_name_mapping <- c("AA" = "Asayita", 
                         "AC" = "Abobo", 
                         "GM" = "Gondar", 
                         "BA" = "Assossa",
                         "MM" = "Mizan")

  # Full site name based on site_filter (abbreviation)
  full_site_name <- site_name_mapping[site_filter]
  
  # Filter the long_ibd_cnv data frame for relevant samples starting with the site filter
  filtered_samples <- long_ibd_cnv %>%
    filter(grepl(paste("^", site_filter, sep = ""), sample.x) & grepl(paste("^", site_filter, sep = ""), sample.y))

  # Filter edges based on relatedness (ibd) and p-value thresholds
  filtered_edges <- filtered_samples %>%
    filter(ibd >= relatedness_threshold, p_BH < pvalue_threshold)

  # Create a data frame of unique sample names for vertices
  unique_samples <- unique(c(filtered_edges$sample.x, filtered_edges$sample.y))

  # Create an igraph object from the filtered edges
  g <- graph_from_data_frame(filtered_edges, directed = FALSE, vertices = data.frame(name = unique_samples))

  # Identify isolates (samples that are not in the edges)
  samples_in_edges <- unique(c(filtered_edges$sample.x, filtered_edges$sample.y))
  all_samples <- unique(c(long_ibd_cnv$sample.x, long_ibd_cnv$sample.y))
  isolates <- setdiff(all_samples, samples_in_edges)

  # Add isolates as vertices in the graph if show_isolates is TRUE
  if (show_isolates) {
    g_full <- g + vertices(isolates)
  } else {
    g_full <- g  # Only use the main graph without isolates
  }

  # Merge site information and HRP deletion status into vertex_data
  vertex_data <- data.frame(name = V(g_full)$name) %>%
    mutate(site = case_when(
      grepl(pattern = "^AA", name) ~ "Asayita",
      grepl(pattern = "^AC", name) ~ "Gambella",
      grepl(pattern = "^GM", name) ~ "Gondar",
      grepl(pattern = "^BA", name) ~ "Assossa",
      grepl(pattern = "^MM", name) ~ "Mizan")) %>%
    left_join(crt_metadata, by = c("name" = "SampleID"))  # Join crt_metadata for HRP deletion status

  V(g_full)$color <- ifelse(vertex_data$Pfcrt == "CVIET",
                           "#dabfff", ifelse(vertex_data$Pfcrt == "CVMNK", "#2c2a4a", ifelse(vertex_data$Pfcrt == "CVMNK|CVIET", "#d67ad1", ifelse(vertex_data$Pfcrt == "-", "grey", "black"))))

  # Set all edge colors to black
  E(g_full)$color <- "black"

  # Set edge widths to ensure visibility
  edge_widths <- pmax(E(g_full)$ibd * 2, 0.2)

  # Use layout_nicely() to automatically select an appropriate layout without jitter
  layout <- layout_nicely(g_full)

  # Create the plot
  plot(g_full, 
       layout = layout, 
       vertex.label = NA,  # Remove labels
       vertex.size = V(g_full)$size,  # Use conditional vertex sizes
       edge.width = edge_widths,  # Use adjusted edge widths
       edge.color = E(g_full)$color,  # All edges set to black
       vertex.shape = V(g_full)$shape,  # Use conditional vertex shapes
       main = full_site_name,
       cex.main = 8) 
}


# Define the sites
site_list <- c("AA", "AC", "GM", "BA", "MM")  # Abbreviations for the sites

pdf("../data/Dcifer/Results/figures/CRT_relatedness_0.9D0.pdf", width = 10, height = 10)

# Set up the plotting grid
par(mfrow = c(1, 5))  # Set 2 rows and 3 columns for grid layout

# Loop over each site and call the function
for (site in site_list) { 
  # Print current site for debugging
  print(paste("Plotting for site:", site))
  
  # Run the function for each site, using the site abbreviation as the filter
  plot_network_cnv_site(long_ibd_cnv = long_ibd_cnv,  # Ensure you're using the correct dataset
                        crt_metadata = crt_metadata, 
                        relatedness_threshold = 0.9, 
                        pvalue_threshold = p_val_BH_0.125,
                        show_isolates = FALSE, 
                        seed = 200, 
                        site_filter = site)  # Pass the abbreviation to filter
}

dev.off()


```




